### 2022-01-17 RDT-equivalence.rmd
Initial commit

---
title: "RDT - Gold Standard Equivalence"
author: "Callum Arnold"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: FALSE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.dim = c(10, 8), fig.align = "center")
```

# Initial Look {.tabset .tabset-pills}

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(hrbrthemes)
library(kableExtra)
library(lubridate)
library(glue)
library(paletteer)
```

```{r}
theme_set(theme_minimal())
```

```{r}

```

### 2022-01-18 RDT-equivalence.rmd
Create shiny app and explore effects of sens and spec on test results
Only just started the shiny app, but have contour plots for number of test positive, and PPV, and table describing the values by two different types of tests (Gold Standard and Rapid)

@@
 library(lubridate)
+library(purrr)
 library(glue)
-library(paletteer)
+library(MetBrewer)
 ```
@@
 ```{r}
+sens_range <- seq(0, 1, 0.05)
+spec_range <- seq(0, 1, 0.05)
+data <- tidyr::crossing(sens_range, spec_range)
+```
 
+```{r}
+prev <- 0.05
+tot <- 100
+
+dis_pos <- tot * prev
+dis_neg <- tot * (1 - prev)
 ```
+
+```{r}
+test <- map2_dfr(
+  .x = data$sens_range,
+  .y = data$spec_range,
+  .f = function(.x, .y){
+    true_pos <- dis_pos * .x
+    true_neg <- dis_neg * .y
+    false_pos <- dis_neg * (1 - .y)
+    false_neg <- dis_pos * (1 - .x)
+
+    test_pos <- true_pos + false_pos
+    test_neg <- true_neg + false_neg
+
+    ppv <- true_pos / test_pos
+    npv <- true_neg / test_neg
+
+    tibble(
+      sens = .x,
+      spec = .y,
+      true_pos = true_pos,
+      true_neg = true_neg,
+      false_pos = false_pos,
+      false_neg = false_neg,
+      test_pos = test_pos,
+      test_neg = test_neg,
+      ppv = ppv,
+      npv = npv
+    )
+  }
+)
+```
+
+```{r}
+ggplot(test, aes(x = sens, y = spec, z = test_pos)) +
+  geom_contour_filled() +
+  scale_fill_manual(values = met.brewer("OKeeffe2", 10)) +
+  labs(
+    title = "Number of Positive Tests by Sensitivity and Specificity",
+    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}, Disease Positives = {dis_pos}"),
+    x = "Sensitivity",
+    y = "Specificity",
+    fill = "Number of Positive Tests"
+  ) +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+ggplot(test, aes(x = sens, y = spec, z = ppv)) +
+  geom_contour_filled() +
+  scale_fill_manual(values = met.brewer("Hokusai2", 10)) +
+  labs(
+    title = "PPV by Sensitivity and Specificity",
+    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}, Disease Positives = {dis_pos}"),
+    x = "Sensitivity",
+    y = "Specificity",
+    fill = "PPV"
+  ) +
+  theme(legend.position = "bottom")
+```

### 2022-01-18 app.R

@@ -1,0 +1,255 @@
+library(shiny)
+library(tidyverse)
+library(rlang)
+library(MetBrewer)
+
+
+# Define UI for dataset viewer app ----
+ui <- fluidPage(
+
+  # App title ----
+  titlePanel("RDT - Gold Standard Equivalence"),
+
+  # Sidebar layout with a input and output definitions ----
+  sidebarLayout(
+
+    # Sidebar panel for inputs ----
+    sidebarPanel(
+
+      # Input: Selector for tested cases ----
+      numericInput(
+        inputId = "num_tested",
+        label = "Choose the number of individuals tested:",
+        value = 100,
+        min = 0,
+        max = 100
+        ),
+
+      # Input: Selector for Prevalence ----
+      numericInput(
+        inputId = "prev",
+        label = "Choose the underlying prevalence (%) of measles:",
+        value = 5,
+        min = 0,
+        max = 100
+        ),
+
+      # Input: Selector for GS ----
+      numericInput(
+        inputId = "gs_sens",
+        label = "Choose the sensitivity (%) of the Gold Standard test:",
+        value = 80,
+        min = 0,
+        max = 100
+        ),
+      numericInput(
+        inputId = "gs_spec",
+        label = "Choose the specificity (%) of the Gold Standard test:",
+        value = 80,
+        min = 0,
+        max = 100
+        ),
+
+      # Input: Selector for RDT ----
+      numericInput(
+        inputId = "rdt_sens",
+        label = "Choose the sensitivity (%) of the Rapid Diagnostic test:",
+        value = 80,
+        min = 0,
+        max = 100
+        ),
+      numericInput(
+        inputId = "rdt_spec",
+        label = "Choose the specificity (%) of the Rapid Diagnostic test:",
+        value = 80,
+        min = 0,
+        max = 100
+        ),
+
+      
+    ),
+
+    # Main panel for displaying outputs ----
+    mainPanel(
+      
+      tabsetPanel(
+        # Output: Plot of results ----
+        tabPanel(
+          "Test Positive Plot",
+          plotOutput("test_pos_plot")
+        ),
+        tabPanel(
+          "PPV Plot",
+          plotOutput("ppv_plot")
+        ),
+
[…]

### 2022-01-20 RDT-equivalence.rmd
Update Rmd to look at uncertainty in true cases
Simulate true cases drawn from binomial distribution. Will need to think about how to incorporate uncertainty from testing

@@
 
-# Initial Look {.tabset .tabset-pills}
+# Disease Positive -> Test Positive  {.tabset .tabset-pills}
 
@@
 library(MetBrewer)
+library(future)
+library(furrr)
 ```
@@
 
+## Deterministic
+
 ```{r}
@@
 ```{r}
-test <- map2_dfr(
-  .x = data$sens_range,
-  .y = data$spec_range,
-  .f = function(.x, .y){
-    true_pos <- dis_pos * .x
-    true_neg <- dis_neg * .y
-    false_pos <- dis_neg * (1 - .y)
-    false_neg <- dis_pos * (1 - .x)
+test <- pmap_dfr(
+    .l = list(sens = data$sens_range, spec = data$spec_range),
+    .f = function(sens, spec){
+      true_pos <- dis_pos * sens
+      true_neg <- dis_neg * spec
+      false_pos <- dis_neg * (1 - spec)
+      false_neg <- dis_pos * (1 - sens)
 
@@
       tibble(
-      sens = .x,
-      spec = .y,
+        sens = sens,
+        spec = spec,
         true_pos = true_pos,
@@
 ```
+
+## Uncertainty
+
+```{r}
+plan(multisession)
+```
+
+```{r}
+set.seed(1234)
+
+stoch_dis_pos_neg <- tibble(
+  tot = tot,
+  pos = rbinom(n = 1e2, size = tot, prob = dis_pos/100)
+  ) %>%
+  mutate(neg = tot - pos)
+```
+
+```{r}
+ggplot(stoch_dis_pos_neg, aes(x = pos)) +
+  geom_bar() +
+  labs(
+    title = "Distribution of Disease Positives",
+    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
+  )
+```
+
+```{r}
+stoch_test <- future_pmap_dfr(
+  .l = list(
+    iter = 1:nrow(stoch_dis_pos_neg),
+    dis_pos = stoch_dis_pos_neg$pos,
+    dis_neg = stoch_dis_pos_neg$neg
+    ),
+  .f = function(iter, dis_pos, dis_neg){
+    future_pmap_dfr(
+      .l = list(
+        sens_spec_comb = 1:nrow(data),
+        sens = data$sens_range,
+        spec = data$spec_range
+        ),
+      .f = function(sens_spec_comb, sens, spec){
+        true_pos <- dis_pos * sens
+        true_neg <- dis_neg * spec
+        false_pos <- dis_neg * (1 - spec)
+        false_neg <- dis_pos * (1 - sens)
+
+        test_pos <- true_pos + false_pos
+        test_neg <- true_neg + false_neg
+
+        ppv <- true_pos / test_pos
+        npv <- true_neg / test_neg
+
+        tibble(
+          iter = iter,
+          sens_spec_comb = sens_spec_comb,
+          sens = sens,
+          spec = spec,
+          dis_pos = dis_pos,
+          dis_neg = dis_neg,
+          true_pos = true_pos,
+          true_neg = true_neg,
+          false_pos = false_pos,
+          false_neg = false_neg,
+          test_pos = test_pos,
+          test_neg = test_neg,
+          ppv = ppv,
+          npv = npv
+        )
+      },
+      .options = furrr_options(seed = 12345)
+    )
+  },
+  .options = furrr_options(seed = 12345)
+)
+```
+
+```{r}
+stoch_test %>%
+  filter(sens_spec_comb == 1) %>%
+  ggplot() +
+  geom_bar(aes(x = dis_pos), fill = "black") +
+  geom_bar(aes(x = test_pos), fill = "blue") +
+  labs(
+    title = "Example Distributions of Disease and Test Positives",
+    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}"),
+    x = "Number of Positives",
+    caption = "Blue = Test Positives, Black = Disease Positives"
+  )
+```
+
+```{r}
+stoch_test2 <- stoch_test %>%
+  mutate(
+    test_dis_pos_diff = test_pos - dis_pos
+  )
+```
+
+```{r}
+stoch_test2 %>%
+  filter(sens_spec_comb == 1) %>%
+  ggplot(aes(x = test_dis_pos_diff)) +
+  geom_bar() +
+  labs(
+    title = "Example Difference in Test Positives and Disease Positives",
+    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}"),
+    x = "Difference in Test Positives and Disease Positives",
+    caption = glue("Using sensitivity = {filter(stoch_test2, sens_spec_comb == 1)$sens}, specificity = {filter(stoch_test2, sens_spec_comb == 1)$spec}")
+  )
+```
+
+```{r}
+stoch_test2 %>%
+  filter(sens_spec_comb == 100) %>%
+  ggplot(aes(x = test_dis_pos_diff)) +
+  geom_bar() +
+  labs(
+    title = "Example Difference in Test Positives and Disease Positives",
+    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}"),
+    x = "Difference in Test Positives and Disease Positives",
+    caption = glue("Using sensitivity = {filter(stoch_test2, sens_spec_comb == 100)$sens}, specificity = {filter(stoch_test2, sens_spec_comb == 100)$spec}")
+  )
+```
+
+```{r}
+stoch_test3 <- future_map_dfr(
+  .x = unique(stoch_test2$sens_spec_comb),
+  .f = function(.x){
+    stoch_test2 %>%
+      filter(sens_spec_comb == .x) %>%
+      summarize(
+        sens_spec_comb = .x,
+        sens = unique(sens),
+        spec = unique(spec),
+        mean_test_dis_pos_diff = mean(test_dis_pos_diff),
+        sd_test_dis_pos_diff = sd(test_dis_pos_diff),
+        med_test_dis_pos_diff = median(test_dis_pos_diff),
+        min_test_dis_pos_diff = min(test_dis_pos_diff),
+        max_test_dis_pos_diff = max(test_dis_pos_diff),
+        quantile_10 = quantile(test_dis_pos_diff, .10),
+        quantile_25 = quantile(test_dis_pos_diff, .25),
+        quantile_50 = quantile(test_dis_pos_diff, .50),
+        quantile_75 = quantile(test_dis_pos_diff, .75),
+        quantile_90 = quantile(test_dis_pos_diff, .90)
+        )
+  },
+  .options = furrr_options(seed = 12345)
+)
+```
+
+```{r}
+stoch_test3 %>%
+  ggplot() +
+  geom_contour_filled(aes(x = sens, y = spec, z = min_test_dis_pos_diff)) +
+  scale_fill_manual(values = met.brewer("Robert", 11)) +
+  labs(
+    title = "Minimum Difference in Test Positives and Disease Positives",
+    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
+  )
+```
+
+```{r}
+stoch_test3 %>%
+  ggplot() +
+  geom_contour_filled(aes(x = sens, y = spec, z = quantile_10)) +
+  scale_fill_manual(values = met.brewer("Robert", 11)) +
+  labs(
+    title = "10% Quantile Difference in Test Positives and Disease Positives",
+    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
+  )
+```
+
+```{r}
+stoch_test3 %>%
+  ggplot() +
+  geom_contour_filled(aes(x = sens, y = spec, z = mean_test_dis_pos_diff)) +
+  scale_fill_manual(values = met.brewer("OKeeffe2", 11)) +
+  labs(
+    title = "Mean Difference in Test Positives and Disease Positives",
+    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
+  )
+```
+
+```{r}
+stoch_test3 %>%
+  ggplot() +
+  geom_contour_filled(aes(x = sens, y = spec, z = quantile_90)) +
+  scale_fill_manual(values = met.brewer("Peru2", 11)) +
+  labs(
+    title = "90% Quantile Difference in Test Positives and Disease Positives",
+    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
+  )
+```
+
+```{r}
+stoch_test3 %>%
+  ggplot() +
+  geom_contour_filled(aes(x = sens, y = spec, z = max_test_dis_pos_diff)) +
+  scale_fill_manual(values = met.brewer("Peru2", 11)) +
+  labs(
+    title = "Maximum Difference in Test Positives and Disease Positives",
+    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
+  )
+```

### 2022-01-25 RDT-equivalence.rmd
Update app and RMD to look at NegBin (num of tests required to detect)

@@
 ```
+
+# Test Positive (Discounting Testing Rate)
+
+Here, we're going to assume that the reference test is 100% sensitive and 100% specific. We're also going to discount the testing rate and the prevalence. The questions we're going to ask is:
+
+1. If there are x disease positives, what is the distribution of test positives, given a certain sensitivity and specificity?
+2. Given x disease positives, and the above distribution of test positives, how many additional test positives would be required such that y% of the distribution of positives was $\ge$ x?
+  - **Is this just the negative binomial distribution?**
+
+We'll start with the example of 3 disease positives, and a test of 80% sensitivity and 80% specificity.
+
+```{r}
+sens <- 0.8
+spec <- 0.8
+
+dis_pos <- 3
+```
+
+```{r}
+test_pos <- tibble(positives = rbinom(1e5, dis_pos + 4, sens))
+```
+
+```{r}
+ggplot(test_pos) +
+  geom_bar(aes(x = positives))
+```
+
+```{r}
+test_pos %>%
+  summarize(
+    mean = mean(positives),
+    sd = sd(positives),
+    min = min(positives),
+    quantile_10 = quantile(positives, .10),
+    quantile_25 = quantile(positives, .25),
+    quantile_50 = quantile(positives, .50),
+    quantile_75 = quantile(positives, .75),
+    quantile_90 = quantile(positives, .90),
+    max = max(positives)
+  )
+```
+
+We can use the negative binomial to examine the number of tests required to achieve a certain level of disease positives.
+In the example below, we'll use the same sensitivity (80%) and disease positives (3).
+
+```{r}
+quant <- 0.95
+
+total_test_dist <- tibble(
+  failures = seq(0, qnbinom(p = quant, size = 3, prob = 0.4), 1),
+  total = failures + dis_pos,
+  dens = dnbinom(failures, 3, prob = 0.4),
+  cum_prob = pnbinom(failures, 3, prob = 0.4)
+  ) 
+```
+
+```{r}
+
+qnbinom(p = quant, size = 3, prob = 0.8)
+```
+
+```{r}
+total_test_dist %>%
+  ggplot(aes(x = total, y = dens, fill = dens)) +
+  geom_col()+
+  scale_x_continuous(
+    breaks = seq(min(total_test_dist$total), max(total_test_dist$total), 1)
+    ) +
+  geom_vline(xintercept = dis_pos + qnbinom(p = quant, size = 3, prob = 0.4)) +
+  scale_fill_gradientn(colors = rev(met.brewer("Greek")))
+```
+
+```{r}
+total_test_dist %>%
+  ggplot() +
+  geom_col(aes(x = total, y = cum_prob, fill = total)) +
+  scale_x_continuous(
+    breaks = seq(min(total_test_dist$total), max(total_test_dist$total), 1)
+    ) +
+  geom_hline(yintercept = quant) +
+  geom_vline(xintercept = dis_pos + qnbinom(p = quant, size = 3, prob = 0.8)) +
+  scale_fill_gradientn(colors = met.brewer("Hokusai2"))
+```

### 2022-01-25 app.R

@@ -4,6 +4,72 @@ library(rlang)
 library(MetBrewer)
 
 
+selection_tabs <- tabsetPanel(
+  id = "selection_tabs",
+  type = "hidden",
+  tabPanel(
+    "Total",
+    # Input: Selector for tested cases ----
+    sliderInput(
+      inputId = "num_tested",
+      label = "Choose the number of individuals tested:",
+      value = 100,
+      min = 0,
+      max = 100
+      ),
+    # Input: Selector for Prior Probability ----
+    sliderInput(
+      inputId = "prior_prob",
+      label = "Choose the prior probability (%) of a positive measles test:",
+      value = 60,
+      min = 0,
+      max = 100
+      ),
+    # Input: Selector for GS ----
+    sliderInput(
+      inputId = "gs_spec",
+      label = "Choose the specificity (%) of the Gold Standard test:",
+      value = 97.3,
+      min = 0,
+      max = 100
+      ),
+    sliderInput(
+      inputId = "gs_sens",
+      label = "Choose the sensitivity (%) of the Gold Standard test:",
+      value = 98.1,
+      min = 0,
+      max = 100
+      ),
+    # Input: Selector for RDT ----
+    sliderInput(
+      inputId = "rdt_spec",
+      label = "Choose the specificity (%) of the Rapid Diagnostic test:",
+      value = 96.2,
+      min = 0,
+      max = 100
+      )
+  ),
+  tabPanel(
+    "Positive", 
+    # Input: Selector for positive cases ----
+    sliderInput(
+      inputId = "pos_tested",
+      label = "Choose the number of positive cases identified by the Gold Standard:",
+      value = 3,
+      min = 0,
+      max = 20
+      ),
+    # Input: Selector for Quantile ----
+    sliderInput(
+      inputId = "percentile_pos",
+      label = "Choose the percentile of test distribution to return the desired number of true positive cases:",
+      value = 50,
+      min = 0,
+      max = 100
+      )
+  )
+)
+
 # Define UI for dataset viewer app ----
 ui <- fluidPage(
 
@@ -15,53 +81,27 @@ ui <- fluidPage(
 
     # Sidebar panel for inputs ----
     sidebarPanel(
+      # Input: Select whether to input number of gold standard positives or
+      # total individuals tested
+      radioButtons(
+        inputId = "pos_or_total",
+        label = "Do you want to input the number of gold standard positives or the total number of individuals tested?",
+        selected = "Positive",
[…]

### 2023-02-24 notebooks/R/RDT-equivalence.rmd
Set up DrWatson Julia project and Quarto project, and move R code to subdirs

--- 2022-01-25 RDT-equivalence.rmd
@@
   geom_contour_filled(aes(x = sens, y = spec, z = min_test_dis_pos_diff)) +
-  scale_fill_manual(values = met.brewer("Robert", 11)) +
+  scale_fill_manual(values = met.brewer("Nattier", 11)) +
   labs(
@@
   geom_contour_filled(aes(x = sens, y = spec, z = quantile_10)) +
-  scale_fill_manual(values = met.brewer("Robert", 11)) +
+  scale_fill_manual(values = met.brewer("Nattier", 11)) +
   labs(

### 2023-02-24 scripts/intro.jl

@@ -1,0 +1,19 @@
+using DrWatson
+@quickactivate "OutbreakDetection"
+
+# Here you may include files from the source directory
+include(srcdir("dummy_src_file.jl"))
+
+println(
+"""
+Currently active project is: $(projectname())
+
+Path of active project: $(projectdir())
+
+Have fun with your new project!
+
+You can help us improve DrWatson by opening
+issues on GitHub, submitting feature requests,
+or even opening your own Pull Requests!
+"""
+)

### 2023-02-24 src/R/app.R

@@ -1,0 +1,492 @@
+library(shiny)
+library(tidyverse)
+library(rlang)
+library(paletteer)
+
+
+selection_tabs <- tabsetPanel(
+  id = "selection_tabs",
+  type = "hidden",
+  tabPanel(
+    "Total",
+    # Input: Selector for tested cases ----
+    sliderInput(
+      inputId = "num_tested",
+      label = "Choose the number of individuals tested:",
+      value = 100,
+      min = 0,
+      max = 100
+      ),
+    # Input: Selector for Prior Probability ----
+    sliderInput(
+      inputId = "prior_prob",
+      label = "Choose the prior probability (%) of a positive measles test:",
+      value = 60,
+      min = 0,
+      max = 100
+      ),
+    # Input: Selector for GS ----
+    sliderInput(
+      inputId = "gs_spec",
+      label = "Choose the specificity (%) of the Gold Standard test:",
+      value = 97.3,
+      min = 0,
+      max = 100
+      ),
+    sliderInput(
+      inputId = "gs_sens",
+      label = "Choose the sensitivity (%) of the Gold Standard test:",
+      value = 98.1,
+      min = 0,
+      max = 100
+      ),
+    # Input: Selector for RDT ----
+    sliderInput(
+      inputId = "rdt_spec",
+      label = "Choose the specificity (%) of the Rapid Diagnostic test:",
+      value = 96.2,
+      min = 0,
+      max = 100
+      )
+  ),
+  tabPanel(
+    "Positive", 
+    # Input: Selector for positive cases ----
+    sliderInput(
+      inputId = "pos_tested",
+      label = "Choose the number of positive cases identified by the Gold Standard:",
+      value = 3,
+      min = 0,
+      max = 20
+      ),
+    # Input: Selector for Quantile ----
+    sliderInput(
+      inputId = "percentile_pos",
+      label = "Choose the percentile of test distribution to return the desired number of true positive cases:",
+      value = 50,
+      min = 0,
+      max = 100
+      )
+  )
+)
+
+# Define UI for dataset viewer app ----
+ui <- fluidPage(
+
+  # App title ----
+  titlePanel("RDT - Gold Standard Equivalence"),
+
+  # Sidebar layout with a input and output definitions ----
+  sidebarLayout(
+
+    # Sidebar panel for inputs ----
+    sidebarPanel(
[…]

### 2023-02-24 test/runtests.jl

@@ -1,0 +1,17 @@
+using DrWatson, Test
+@quickactivate "OutbreakDetection"
+
+# Here you include files using `srcdir`
+# include(srcdir("file.jl"))
+
+# Run test suite
+println("Starting tests")
+ti = time()
+
+@testset "OutbreakDetection tests" begin
+    @test 1 == 1
+end
+
+ti = time() - ti
+println("\nTest took total time of:")
+println(round(ti/60, digits = 3), " minutes")

### 2023-02-24 notebooks/Julia/simulation-detection.qmd
Initial stochastic SIR model

---
title: "Outbreak Threshold Detection"
author: "Callum Arnold"
date: today
title-block-banner: true
format:
  html:
    theme: flatly
    toc: true
    toc-depth: 4
    code-copy: true
    code-fold: true
    code-overflow: wrap
    code-tools: true
    code-link: true
    anchor-sections: true
    fig-width: 10
    fig-height: 8
execute:
  cache: true
jupyter: julia-1.8
---

```{julia}
using DrWatson
quickactivate(@__DIR__, "OutbreakDetection")
```

```{julia}
using ModelingToolkit, JumpProcesses, Random, Statistics
using DataFrames, CairoMakie

CairoMakie.activate!()
```

```{julia}
@parameters β γ
@variables t S(t) I(t) R(t)
N = S + I + R
rate₁   = β * S * I/N
affect₁ = [S ~ S - 1, I ~ I + 1]
rate₂   = γ*I
affect₂ = [I ~ I - 1, R ~ R + 1]
j₁      = ConstantRateJump(rate₁,affect₁)
j₂      = ConstantRateJump(rate₂,affect₂)
@named js      = JumpSystem([j₁, j₂], t, [S, I, R], [β, γ])
```

```{julia}
u₀ = [S => 990, I => 10, R => 0];
p = [β => 0.5, γ => 0.25];
```

```{julia}
tmax = 250.0
tspan = (0.0, tmax)
δt = 1.0;
Random.seed!(1234);
```

```{julia}
function run_sir_sim(; js, u₀, tspan, p)
    prob = DiscreteProblem(js, u₀, tspan, p)
    sir_jprob = JumpProblem(js, prob, Direct());
    sol_jump = solve(sir_jprob, SSAStepper());
    return sol_jump(0:δt:tmax);
end
```

```{julia}
function create_sim_array(; out_jump)
    array_jump[1:3, :] = Array(out_jump)
    array_jump[4, :] .= out_jump.t

    return array_jump
end
```

```{julia}
sir_sim_plot = Figure()
ax = Axis(sir_sim_plot[1, 1],
    xlabel = "Time",
    ylabel = "Number"
)

nsims = 100

array_jump = zeros(Float64, 4, Int64(tmax) + 1)
all_sims_array = zeros(Float64, 4, Int64(tmax) + 1, nsims)
sim_means = zeros(Float64, 3, Int(tmax) + 1)

for i in 1:nsims
    out_jump = run_sir_sim(js = js, u₀ = u₀, tspan = tspan, p = p)
    array_jump = create_sim_array(out_jump = out_jump)

    all_sims_array[:, :, i] = array_jump

    lines!(ax, array_jump[4, :], array_jump[1, :], linewidth = 4, color = ("dodgerblue4", 0.05))
    lines!(ax, array_jump[4, :], array_jump[2, :], linewidth = 4, color = ("firebrick3", 0.05))
    lines!(ax, array_jump[4, :], array_jump[3, :], linewidth = 4, color = ("chocolate2", 0.05))

end

for sim in 1:Int(tmax)+1, outcome in 1:3
    sim_means[outcome, sim] = mean(all_sims_array[outcome, sim, :])
end

lines!(ax, array_jump[4, :], sim_means[1, :], linewidth = 4, color = "dodgerblue4")
lines!(ax, array_jump[4, :], sim_means[2, :], linewidth = 4, color = "firebrick3")
lines!(ax, array_jump[4, :], sim_means[3, :], linewidth = 4, color = "chocolate2")

sir_sim_plot
```

### 2023-02-25 notebooks/Julia/simulation-detection.qmd
Update stochastic model to use tau-leaping Gillespie Needed to change the core functions to be able to use tau-leaping as modelingtoolkit doesn't seem to handle leaping.

@@
 ```{julia}
-using ModelingToolkit, JumpProcesses, Random, Statistics
-using DataFrames, CairoMakie
+using JumpProcesses, StochasticDiffEq, Random, Statistics, LinearAlgebra
+using DataFrames, CairoMakie, ProgressMeter
 
@@
 ```{julia}
-@parameters β γ
-@variables t S(t) I(t) R(t)
-N = S + I + R
-rate₁   = β * S * I/N
-affect₁ = [S ~ S - 1, I ~ I + 1]
-rate₂   = γ*I
-affect₂ = [I ~ I - 1, R ~ R + 1]
-j₁      = ConstantRateJump(rate₁,affect₁)
-j₂      = ConstantRateJump(rate₂,affect₂)
-@named js      = JumpSystem([j₁, j₂], t, [S, I, R], [β, γ])
-```
-
-```{julia}
-u₀ = [S => 990, I => 10, R => 0];
-p = [β => 0.5, γ => 0.25];
+β = 0.1 / 1000.0
+ν = 0.01;
+p = (β, ν)
 ```
@@
 tmax = 250.0
+δt = 0.001
 tspan = (0.0, tmax)
-δt = 1.0;
+tlength = length(tspan[1]:δt:tspan[2])
 Random.seed!(1234);
@@
 ```{julia}
-function run_sir_sim(; js, u₀, tspan, p)
-    prob = DiscreteProblem(js, u₀, tspan, p)
-    sir_jprob = JumpProblem(js, prob, Direct());
-    sol_jump = solve(sir_jprob, SSAStepper());
+function rate(out, u, p, t)
+    out[1] = p[1] * u[1] * u[2]   # β * S * I
+    out[2] = p[2] * u[2]          # ν * I
+    nothing
+end
+```
+
+```{julia}
+c = zeros(3, 2)
+# S + I --> I
+c[1, 1] = -1    # S -> S - 1
+c[2, 1] = 1     # I -> I + 1
+
+# I --> R
+c[2, 2] = -1    # I -> I - 1
+c[3, 2] = 1     # R -> R + 1
+```
+
+```{julia}
+function change(du, u, p, t, counts, mark)
+    mul!(du, c, counts)
+    nothing
+end
+```
+
+```{julia}
+rj = RegularJump(rate, change, 2)
+
+u₀ = [1000.0, 50.0, 0.0]
+prob = DiscreteProblem(u₀, tspan, p)
+jump_prob = JumpProblem(prob, Direct(), rj)
+
+sol = solve(jump_prob, TauLeaping(); dt = δt)
+```
+
+```{julia}
+sir_array = zeros(Float64, 4, tlength) 
+sir_array[1:3, :] = Array(sol)
+sir_array[4, :] = sol.t
+```
+
+```{julia}
+f = Figure()
+ax = Axis(
+    f[1, 1],
+    xlabel = "Time",
+    ylabel = "Number"
+    )
+
+lines!(ax, sir_array[4, :], sir_array[1, :])
+lines!(ax, sir_array[4, :], sir_array[2, :])
+lines!(ax, sir_array[4, :], sir_array[3, :])
+
+f
+```
+
+```{julia}
+function run_sir_sim(; u₀, tspan, p)
+    prob = DiscreteProblem(u₀, tspan, p)
+    jump_prob = JumpProblem(prob, Direct(), rj)
+    sol_jump = solve(jump_prob, TauLeaping(); dt = δt)
     return sol_jump(0:δt:tmax);
@@
 function create_sim_array(; out_jump)
-    array_jump[1:3, :] = Array(out_jump)
-    array_jump[4, :] .= out_jump.t
+    test_array[1:3, :] = Array(out_jump)
+    test_array[4, :] .= out_jump.t
 
-    return array_jump
+    return test_array
 end
@@
 
-array_jump = zeros(Float64, 4, Int64(tmax) + 1)
-all_sims_array = zeros(Float64, 4, Int64(tmax) + 1, nsims)
-sim_means = zeros(Float64, 3, Int(tmax) + 1)
+sir_array = zeros(Float64, 4, tlength)
+all_sims_array = zeros(Float64, 4, tlength, nsims)
+sim_means = zeros(Float64, 3, tlength)
 
-for i in 1:nsims
-    out_jump = run_sir_sim(js = js, u₀ = u₀, tspan = tspan, p = p)
-    array_jump = create_sim_array(out_jump = out_jump)
+@showprogress for i in 1:nsims
+    out_jump = run_sir_sim(u₀ = u₀, tspan = tspan, p = p)
+    sir_array = create_sim_array(out_jump = out_jump)
 
-    all_sims_array[:, :, i] = array_jump
+    all_sims_array[:, :, i] = sir_array
 
-    lines!(ax, array_jump[4, :], array_jump[1, :], linewidth = 4, color = ("dodgerblue4", 0.05))
-    lines!(ax, array_jump[4, :], array_jump[2, :], linewidth = 4, color = ("firebrick3", 0.05))
-    lines!(ax, array_jump[4, :], array_jump[3, :], linewidth = 4, color = ("chocolate2", 0.05))
+    lines!(ax, sir_array[4, :], sir_array[1, :], linewidth = 4, color = ("dodgerblue4", 0.01))
+    lines!(ax, sir_array[4, :], sir_array[2, :], linewidth = 4, color = ("firebrick3", 0.01))
+    lines!(ax, sir_array[4, :], sir_array[3, :], linewidth = 4, color = ("chocolate2", 0.01))
 
@@
 
-for sim in 1:Int(tmax)+1, outcome in 1:3
+@showprogress for sim in 1:tlength, outcome in 1:3
     sim_means[outcome, sim] = mean(all_sims_array[outcome, sim, :])
@@
 
-lines!(ax, array_jump[4, :], sim_means[1, :], linewidth = 4, color = "dodgerblue4")
-lines!(ax, array_jump[4, :], sim_means[2, :], linewidth = 4, color = "firebrick3")
-lines!(ax, array_jump[4, :], sim_means[3, :], linewidth = 4, color = "chocolate2")
+lines!(ax, sir_array[4, :], sim_means[1, :], linewidth = 3, color = "dodgerblue4")
+lines!(ax, sir_array[4, :], sim_means[2, :], linewidth = 3, color = "firebrick3")
+lines!(ax, sir_array[4, :], sim_means[3, :], linewidth = 3, color = "chocolate2")
 

### 2023-03-01 notebooks/Julia/simulation-detection.qmd
Update initial test SIR to add births and deaths Needed to switch to specifying the rates and affects directly as MTK implementation wouldn't solve with tau-leaping. Might be worth re-visiting as would be useful to use tau-leaping when running multiple simulations.
Updated code implementation to explicitly use frequency-dependent
transmission model (note `rate1` includes `/N` term).

@@
 ```{julia}
-using JumpProcesses, StochasticDiffEq, Random, Statistics, LinearAlgebra
-using DataFrames, CairoMakie, ProgressMeter
+using JumpProcesses, StochasticDiffEq, DifferentialEquations, LinearAlgebra
+using Catalyst, Random, Statistics, DataFrames, CairoMakie, ProgressMeter
 
@@
 ```{julia}
-β = 0.1 / 1000.0
-ν = 0.01;
-p = (β, ν)
+X = 450
+Y = 50
+Z = 0
+N = X + Y + Z
+u₀ = [X, Y, Z, N]
 ```
@@
 ```{julia}
-tmax = 250.0
+R₀ = 12
+γ = 1/8
+μ = 1 / (365 * 62.5)
+β = R₀ * (γ + μ)
+p = (; β, γ, μ)
+```
+
+```{julia}
+tmax = 100 * 365
 δt = 0.001
@@
 ```{julia}
-function rate(out, u, p, t)
-    out[1] = p[1] * u[1] * u[2]   # β * S * I
-    out[2] = p[2] * u[2]          # ν * I
+rate1(u, p, t) = p.β * u[1] * u[2] / u[4]  # β*X*Y / N
+
+function affect1!(integrator)
+    integrator.u[1] -= 1         # X -> X - 1
+    integrator.u[2] += 1         # Y -> Y + 1
     nothing
 end
+
+jump = ConstantRateJump(rate1, affect1!)
 ```
@@
 ```{julia}
-c = zeros(3, 2)
-# S + I --> I
-c[1, 1] = -1    # S -> S - 1
-c[2, 1] = 1     # I -> I + 1
+rate2(u, p, t) = p.γ * u[2]         # γ*Y
 
-# I --> R
-c[2, 2] = -1    # I -> I - 1
-c[3, 2] = 1     # R -> R + 1
+function affect2!(integrator)
+    integrator.u[2] -= 1        # Y -> Y - 1
+    integrator.u[3] += 1        # Z -> Z + 1
+    nothing
+end
+
+jump2 = ConstantRateJump(rate2, affect2!)
 ```
@@
 ```{julia}
-function change(du, u, p, t, counts, mark)
-    mul!(du, c, counts)
+birth_rate(u, p, t) = p.μ * u[4]    # μ * N
+
+function birth_affect!(integrator)
+    integrator.u[1] += 1
+    integrator.u[4] += 1
     nothing
 end
+
+birth_jump = ConstantRateJump(birth_rate, birth_affect!)
 ```
@@
 ```{julia}
-rj = RegularJump(rate, change, 2)
+death_rate(u, p, t) = p.μ * u[4]    # μ * N
 
-u₀ = [1000.0, 50.0, 0.0]
-prob = DiscreteProblem(u₀, tspan, p)
-jump_prob = JumpProblem(prob, Direct(), rj)
+function death_affect!(integrator)
+    map!(
+        i -> ifelse(integrator[i] > 0, integrator.u[i] - 1, integrator.u[i]),
+        integrator.u,
+        eachindex(integrator.u)
+    )
 
-sol = solve(jump_prob, TauLeaping(); dt = δt)
+end
+
+death_jump = ConstantRateJump(death_rate, death_affect!)
 ```
@@
 ```{julia}
-sir_array = zeros(Float64, 4, tlength) 
-sir_array[1:3, :] = Array(sol)
-sir_array[4, :] = sol.t
+prob = DiscreteProblem(u₀, tspan, p)
+jump_prob = JumpProblem(prob, Direct(), jump, jump2, birth_jump, death_jump)
+sol = solve(jump_prob, SSAStepper())
+
+sol_array = Array(sol)
 ```
-
 ```{julia}
@@
 
-lines!(ax, sir_array[4, :], sir_array[1, :])
-lines!(ax, sir_array[4, :], sir_array[2, :])
-lines!(ax, sir_array[4, :], sir_array[3, :])
+lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "X(t)")
+lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "Y(t)")
+lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "Z(t)")
+lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "N(t)")
+
+axislegend(ax)
 

### 2023-03-03 notebooks/Julia/simulation-detection.qmd
Create separate jumps for each compartment and calculate N, rather than being separate jump Wanted each compartment to be able to be born and die separately, and calculating N ensures that the total number is always correct (rather than potentially drifting from the sum of the compartments), which is important when values depend on N.

@@
 ```{julia}
-X = 450
-Y = 50
-Z = 0
-N = X + Y + Z
-u₀ = [X, Y, Z, N]
+X = 800
+Y = 20
+Z = 800
+u₀ = [X, Y, Z]
 ```
@@
 ```{julia}
-R₀ = 12
+R₀ = 16
 γ = 1/8
@@
 ```{julia}
-tmax = 100 * 365
+tmax = 250
 δt = 0.001
@@
 ```{julia}
-rate1(u, p, t) = p.β * u[1] * u[2] / u[4]  # β*X*Y / N
+infection_rate(u, p, t) = p.β * u[1] * u[2] / (u[1] + u[2] + u[3])      # β*X*Y / N
 
-function affect1!(integrator)
+function infection_affect!(integrator)
     integrator.u[1] -= 1         # X -> X - 1
@@
 
-jump = ConstantRateJump(rate1, affect1!)
+infection_jump = ConstantRateJump(infection_rate, infection_affect!)
 ```
@@
 ```{julia}
-rate2(u, p, t) = p.γ * u[2]         # γ*Y
+recovery_rate(u, p, t) = p.γ * u[2]     # γ*Y
 
-function affect2!(integrator)
+function recovery_affect!(integrator)
     integrator.u[2] -= 1        # Y -> Y - 1
@@
 
-jump2 = ConstantRateJump(rate2, affect2!)
+recovery_jump = ConstantRateJump(recovery_rate, recovery_affect!)
 ```
@@
 ```{julia}
-birth_rate(u, p, t) = p.μ * u[4]    # μ * N
+birth_rate(u, p, t) = p.μ * (u[1] + u[2] + u[3])        # μ * N
 
@@
     integrator.u[1] += 1
-    integrator.u[4] += 1
+    # integrator.u[4] += 1
     nothing
@@
 ```{julia}
-death_rate(u, p, t) = p.μ * u[4]    # μ * N
+X_death_rate(u, p, t) = p.μ * u[1]       # μ * X
 
-function death_affect!(integrator)
-    map!(
-        i -> ifelse(integrator[i] > 0, integrator.u[i] - 1, integrator.u[i]),
-        integrator.u,
-        eachindex(integrator.u)
-    )
-
+function X_death_affect!(integrator)
+    integrator.u[1] -= 1
+    nothing
 end
 
-death_jump = ConstantRateJump(death_rate, death_affect!)
+X_death_jump = ConstantRateJump(X_death_rate, X_death_affect!)
+```
+
+```{julia}
+Y_death_rate(u, p, t) = p.μ * u[2]       # μ * Y
+
+function Y_death_affect!(integrator)
+    integrator.u[2] -= 1
+    nothing
+end
+
+Y_death_jump = ConstantRateJump(Y_death_rate, Y_death_affect!)
+```
+
+```{julia}
+Z_death_rate(u, p, t) = p.μ * u[3]       # μ * R
+
+function Z_death_affect!(integrator)
+    integrator.u[3] -= 1
+    nothing
+end
+
+Z_death_jump = ConstantRateJump(Z_death_rate, Z_death_affect!)
+```
+
+```{julia}
+jumps = [infection_jump, recovery_jump, birth_jump, X_death_jump, Y_death_jump, Z_death_jump]
 ```
@@
 prob = DiscreteProblem(u₀, tspan, p)
-jump_prob = JumpProblem(prob, Direct(), jump, jump2, birth_jump, death_jump)
+jump_prob = JumpProblem(prob, Diredt(), jumps...)
 sol = solve(jump_prob, SSAStepper())
 
-sol_array = Array(sol)
+sol_array = zeros(4, length(sol))
+sol_array[1:3, :] = Array(sol)
+sol_array[4, :] = sum(sol_array[1:3, :], dims = 1);
 ```

### 2023-03-03 notebooks/Julia/simulation-detection.qmd
Add Tau-leaping Add RegularJump tau-leaping implementation. Needed to include N explicitly in the states because calculating rates was causing an issue using u[3] i.e. Z state. Need to look into why this is an issue, but for now, tau-leaping produces the expected result

@@
 β = R₀ * (γ + μ)
-p = (; β, γ, μ)
+p = (β, γ, μ)
 ```
@@
 ```{julia}
-tmax = 250
+tmax = 250.0
 δt = 0.001
@@
 ```{julia}
-infection_rate(u, p, t) = p.β * u[1] * u[2] / (u[1] + u[2] + u[3])      # β*X*Y / N
+function rate(out, u, p, t)
+    out[1] = p[1] * u[1] * u[2] / u[4]      # β * X * Y / N
+    out[2] = p[2] * u[2]                    # ν * Y
+    out[3] = p[3] * u[4]                    # μ * N
+    out[4] = p[3] * u[1]                    # μ * X
+    out[5] = p[3] * u[2]                    # μ * Y
+    out[6] = p[3] * u[3]                    # μ * Z
+    nothing
+end
+```
+
+```{julia}
+c = zeros(4, 6)
+# X + Y --> Y
+c[1, 1] = -1        # X -> X - 1
+c[2, 1] = 1         # Y -> Y + 1
+
+# Y --> Z
+c[2, 2] = -1        # Y -> Y - 1
+c[3, 2] = 1         # Z -> Z + 1
+
+# X --> 2X
+c[1, 3] = 1         # X -> X + 1
+c[4, 3] = 1         # N -> N + 1
+
+# X --> 0X
+c[1, 4] = -1        # X -> X - 1
+c[4, 4] = -1        # N -> N - 1
+
+# Y --> 0Y
+c[2, 5] = -1        # Y -> Y - 1
+c[4, 5] = -1        # N -> N - 1
+
+# Z --> 0Z
+c[3, 6] = -1        # Z -> Z - 1
+c[4, 6] = -1        # N -> N - 1
+
+function change!(du, u, p, t, counts, mark)
+    mul!(du, c, counts)
+    nothing
+end
+```
+
+```{julia}
+rj = RegularJump(rate, change!, 6)
+```
+
+```{julia}
+u₀ = [800, 20, 800, NaN]
+u₀[4] = sum(u₀[1:3])
+R₀ = 16
+γ = 1/8
+μ = 20 / (365 * 62.5)
+β = R₀ * (γ + μ)
+p = (β, γ, μ)
+prob = DiscreteProblem(u₀, tspan, p)
+jump_prob = JumpProblem(prob, Direct(), rj)
+sol = solve(jump_prob, TauLeaping(); dt = δt)
+sol_array = zeros(5, length(sol))
+sol_array[1:4, :] = Array(sol)
+sol_array[5, :] = sum(sol_array[1:3, :], dims = 1);
+
+sol_array[4, :] == sol_array[5, :]
+```
+
+```{julia}
+f = Figure()
+ax = Axis(
+    f[1, 1],
+    xlabel = "Time",
+    ylabel = "Number"
+)
+
+lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "X(t)")
+lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "Y(t)")
+lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "Z(t)")
+lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "N(t) jump")
+lines!(ax, sol.t, sol_array[5, :], linewidth = 3, label = "N(t) calc")
+
+axislegend(ax)
+
+f
+```
+
+```{julia}
+infection_rate(u, p, t) = p[1] * u[1] * u[2] / (u[1] + u[2] + u[3])      # β*X*Y / N
 
@@
 ```{julia}
-recovery_rate(u, p, t) = p.γ * u[2]     # γ*Y
+recovery_rate(u, p, t) = p[2] * u[2]     # γ*Y
 
@@
 ```{julia}
-birth_rate(u, p, t) = p.μ * (u[1] + u[2] + u[3])        # μ * N
+birth_rate(u, p, t) = p[3] * (u[1] + u[2] + u[3])        # μ * N
 
@@
 ```{julia}
-X_death_rate(u, p, t) = p.μ * u[1]       # μ * X
+X_death_rate(u, p, t) = p[3] * u[1]       # μ * X
 
@@
 ```{julia}
-Y_death_rate(u, p, t) = p.μ * u[2]       # μ * Y
+Y_death_rate(u, p, t) = p[3] * u[2]       # μ * Y
 
@@
 ```{julia}
-Z_death_rate(u, p, t) = p.μ * u[3]       # μ * R
+Z_death_rate(u, p, t) = p[3] * u[3]       # μ * R
 
@@
 prob = DiscreteProblem(u₀, tspan, p)
-jump_prob = JumpProblem(prob, Diredt(), jumps...)
+jump_prob = JumpProblem(prob, Direct(), jumps...)
 sol = solve(jump_prob, SSAStepper())
@@
 sol_array = zeros(4, length(sol))
-sol_array[1:3, :] = Array(sol)
+sol_array[1:4, :] = Array(sol)
 sol_array[4, :] = sum(sol_array[1:3, :], dims = 1);

### 2023-03-03 notebooks/Julia/simulation-detection.qmd
Remove old ConstantJump simulation code as using Tau-Leaping

@@
 CairoMakie.activate!()
-```
-
-```{julia}
-X = 800
-Y = 20
-Z = 800
-u₀ = [X, Y, Z]
-```
-
-```{julia}
-R₀ = 16
-γ = 1/8
-μ = 1 / (365 * 62.5)
-β = R₀ * (γ + μ)
-p = (β, γ, μ)
-```
-
-```{julia}
-tmax = 250.0
-δt = 0.001
-tspan = (0.0, tmax)
-tlength = length(tspan[1]:δt:tspan[2])
-Random.seed!(1234);
 ```
@@
 ```{julia}
+tmax = 250.0
+δt = 0.001
+tspan = (0.0, tmax)
+tlength = length(tspan[1]:δt:tspan[2])
+Random.seed!(1234);
+```
+
+```{julia}
+R₀ = 16
+γ = 1/8
+μ = 1 / (365 * 62.5)
+β = R₀ * (γ + μ)
+p = (β, γ, μ)
+
 u₀ = [800, 20, 800, NaN]
 u₀[4] = sum(u₀[1:3])
-R₀ = 16
-γ = 1/8
-μ = 20 / (365 * 62.5)
-β = R₀ * (γ + μ)
-p = (β, γ, μ)
+```
+
+```{julia}
 prob = DiscreteProblem(u₀, tspan, p)
@@
 lines!(ax, sol.t, sol_array[5, :], linewidth = 3, label = "N(t) calc")
-
-axislegend(ax)
-
-f
-```
-
-```{julia}
-infection_rate(u, p, t) = p[1] * u[1] * u[2] / (u[1] + u[2] + u[3])      # β*X*Y / N
-
-function infection_affect!(integrator)
-    integrator.u[1] -= 1         # X -> X - 1
-    integrator.u[2] += 1         # Y -> Y + 1
-    nothing
-end
-
-infection_jump = ConstantRateJump(infection_rate, infection_affect!)
-```
-
-```{julia}
-recovery_rate(u, p, t) = p[2] * u[2]     # γ*Y
-
-function recovery_affect!(integrator)
-    integrator.u[2] -= 1        # Y -> Y - 1
-    integrator.u[3] += 1        # Z -> Z + 1
-    nothing
-end
-
-recovery_jump = ConstantRateJump(recovery_rate, recovery_affect!)
-```
-
-```{julia}
-birth_rate(u, p, t) = p[3] * (u[1] + u[2] + u[3])        # μ * N
-
-function birth_affect!(integrator)
-    integrator.u[1] += 1
-    # integrator.u[4] += 1
-    nothing
-end
-
-birth_jump = ConstantRateJump(birth_rate, birth_affect!)
-```
-
-```{julia}
-X_death_rate(u, p, t) = p[3] * u[1]       # μ * X
-
-function X_death_affect!(integrator)
-    integrator.u[1] -= 1
-    nothing
-end
-
-X_death_jump = ConstantRateJump(X_death_rate, X_death_affect!)
-```
-
-```{julia}
-Y_death_rate(u, p, t) = p[3] * u[2]       # μ * Y
-
-function Y_death_affect!(integrator)
-    integrator.u[2] -= 1
-    nothing
-end
-
-Y_death_jump = ConstantRateJump(Y_death_rate, Y_death_affect!)
-```
-
-```{julia}
-Z_death_rate(u, p, t) = p[3] * u[3]       # μ * R
-
-function Z_death_affect!(integrator)
-    integrator.u[3] -= 1
-    nothing
-end
-
-Z_death_jump = ConstantRateJump(Z_death_rate, Z_death_affect!)
-```
-
-```{julia}
-jumps = [infection_jump, recovery_jump, birth_jump, X_death_jump, Y_death_jump, Z_death_jump]
-```
-
-```{julia}
-prob = DiscreteProblem(u₀, tspan, p)
-jump_prob = JumpProblem(prob, Direct(), jumps...)
-sol = solve(jump_prob, SSAStepper())
-
-sol_array = zeros(4, length(sol))
-sol_array[1:4, :] = Array(sol)
-sol_array[4, :] = sum(sol_array[1:3, :], dims = 1);
-```
-```{julia}
-f = Figure()
-ax = Axis(
-    f[1, 1],
-    xlabel = "Time",
-    ylabel = "Number"
-)
-
-lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "X(t)")
-lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "Y(t)")
-lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "Z(t)")
-lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "N(t)")
 

### 2023-03-06 notebooks/Julia/simulation-detection.qmd
Update infection class notation Change from Keeling and Rohani notation of X/Y/Z to S/I/R with lowercase to denote proportions, as find it easier to add categories with relevant labels - will be useful with latent, vaccinated classes etc

@@
 function rate(out, u, p, t)
-    out[1] = p[1] * u[1] * u[2] / u[4]      # β * X * Y / N
-    out[2] = p[2] * u[2]                    # ν * Y
-    out[3] = p[3] * u[4]                    # μ * N
-    out[4] = p[3] * u[1]                    # μ * X
-    out[5] = p[3] * u[2]                    # μ * Y
-    out[6] = p[3] * u[3]                    # μ * Z
+    out[1] = p[1] * u[1] * u[2] / (u[1] + u[2] + u[3])           # β * S * I
+    out[2] = p[2] * u[2]                      # ν * I
+    # out[3] = p[3] * u[1]                    # μ * N
+    # out[4] = p[3] * u[1]                    # μ * S
+    # out[5] = p[3] * u[2]                    # μ * I
+    # out[6] = p[3] * u[3]                    # μ * R
     nothing
@@
 ```{julia}
-c = zeros(4, 6)
-# X + Y --> Y
-c[1, 1] = -1        # X -> X - 1
-c[2, 1] = 1         # Y -> Y + 1
+c = zeros(3, 2)
+# S + I --> I
+c[1, 1] = -1        # S -> S - 1
+c[2, 1] = 1         # I -> I + 1
 
-# Y --> Z
-c[2, 2] = -1        # Y -> Y - 1
-c[3, 2] = 1         # Z -> Z + 1
+# I --> R
+c[2, 2] = -1        # I -> I - 1
+c[3, 2] = 1         # R -> R + 1
 
-# X --> 2X
-c[1, 3] = 1         # X -> X + 1
-c[4, 3] = 1         # N -> N + 1
+# S --> 2X
+# c[1, 3] = 1         # S -> S + 1
 
-# X --> 0X
-c[1, 4] = -1        # X -> X - 1
-c[4, 4] = -1        # N -> N - 1
+# # S --> 0X
+# c[1, 4] = -1        # S -> S - 1
 
-# Y --> 0Y
-c[2, 5] = -1        # Y -> Y - 1
-c[4, 5] = -1        # N -> N - 1
+# # I --> 0Y
+# c[2, 5] = -1        # I -> I - 1
 
-# Z --> 0Z
-c[3, 6] = -1        # Z -> Z - 1
-c[4, 6] = -1        # N -> N - 1
+# # R --> 0Z
+# c[3, 6] = -1        # R -> R - 1
 
@@
 ```{julia}
-rj = RegularJump(rate, change!, 6)
+rj = RegularJump(rate, change!, 2)
 ```
@@
 ```{julia}
-tmax = 250.0
+tmax = 100.0
 δt = 0.001
@@
 ```{julia}
-R₀ = 16
-γ = 1/8
-μ = 1 / (365 * 62.5)
-β = R₀ * (γ + μ)
+# R₀ = 2
+# γ = 1/8
+μ = 0.001
+# β = R₀ * (γ + μ)
+# ζ = 0.001
+# p = (β, γ, μ, ζ)
+
+u₀ = [1000.0, 50.0, 0.0]
+
+β = 0.1/1000.0
+γ = 0.01
+μ = 0.001
 p = (β, γ, μ)
 
-u₀ = [800, 20, 800, NaN]
-u₀[4] = sum(u₀[1:3])
 ```
@@
 sol = solve(jump_prob, TauLeaping(); dt = δt)
-sol_array = zeros(5, length(sol))
-sol_array[1:4, :] = Array(sol)
-sol_array[5, :] = sum(sol_array[1:3, :], dims = 1);
-
-sol_array[4, :] == sol_array[5, :]
+sol_array = zeros(4, length(sol))
+sol_array[1:3, :] = Array(sol)
+sol_array[4, :] = sum(sol_array[1:3, :], dims = 1)
 ```
@@
 
-lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "X(t)")
-lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "Y(t)")
-lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "Z(t)")
-lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "N(t) jump")
-lines!(ax, sol.t, sol_array[5, :], linewidth = 3, label = "N(t) calc")
+lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "S(t)")
+lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "I(t)")
+lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "R(t)")
+lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "N(t)")
 

### 2023-03-06 notebooks/Julia/simulation-detection.qmd
Add new jump simulation code with SDE problem Re-wrote the code for the Tau-leaping jump problem with births/deaths to try and work out some bugs. Chris Rackauckus suggested that it should use SDEProblem() rather than DiscreteProblem(), so had to add some functions to include Brownian/Wiener motion (as in a Jump-diffusion model).
Currently the jumps can go below 0, so need to use some kind of
switching to implement Direct Gillespie in these instances.

@@
 ```{julia}
+function rate(out, u, p, t)
+    out[1] = p[1] * u[1] * u[2] / (u[1] + u[2] + u[3] + u[4])         # β * S * I / N
+    out[2] = p[2] * u[2]                        # σ * E
+    out[3] = p[3] * u[3]                        # γ * I
+    out[4] = p[4] * (u[1] + u[2] + u[3] + u[4])        # μ * N
+    out[5] = p[4] * u[1]                        # μ * S
+    out[6] = p[4] * u[2]                        # μ * E
+    out[7] = p[4] * u[3]                        # μ * I
+    out[8] = p[4] * u[4]                        # μ * R
+    nothing
+end
+
+c = zeros(4, 8)
+# S + I --> E
+c[1, 1] = -1        # S -> S - 1
+c[2, 1] = 1         # E -> E + 1
+
+# E --> I
+c[2, 2] = -1        # E -> E - 1
+c[3, 2] = 1         # I -> I + 1
+
+# I --> R
+c[3, 3] = -1        # I -> I - 1
+c[4, 3] = 1         # R -> R + 1
+
+# S --> 2S
+c[1, 4] = 1         # S -> S + 1
+
+# S --> 0S
+c[1, 5] = -1        # S -> S - 1
+
+# E --> 0E
+c[2, 6] = -1        # S -> S - 1
+
+# I --> 0I
+c[3, 7] = -1        # I -> I - 1
+
+# R --> 0R
+c[4, 8] = -1        # R -> R - 1
+
+function change!(du, u, p, t, counts, mark)
+    mul!(du, c, counts)
+    nothing
+end
+
+rj = RegularJump(rate, change!, 8)
+
+tmax = 2.0
+δt = 0.1/365
+tspan = (0.0, tmax)
+tlength = length(tspan[1]:δt:tspan[2])
+Random.seed!(1234);
+
+u₀ = [999.0, 10.0, 0.0, 0.0]
+
+μ = 1
+β = 1000
+σ = 365/8
+γ = 365/5
+p = (β, σ, γ, μ)
+
+forcing!(du,u,p,t) = (du .= 0)
+noise!(du,u,p,t) = (du .= 0)
+W = WienerProcess(0.0,0.0,0.0)
+
+prob = SDEProblem(forcing!, noise!, u₀, tspan, p, noise = W)
+jump_prob = JumpProblem(prob, Direct(), rj)
+sol = solve(jump_prob, EM(), dt = δt, maxiters = 1e5)
+sol_array = zeros(5, length(sol))
+sol_array[1:4, :] = Array(sol)
+sol_array[5, :] = sum(sol_array[1:4, :], dims = 1)
+
+fig = Figure()
+ax = Axis(
+    fig[1, 1],
+    xlabel = "Time",
+    ylabel = "Number"
+)
+
+lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "S(t)")
+lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "E(t)")
+lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "I(t)")
+lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "R(t)")
+lines!(ax, sol.t, sol_array[5, :], linewidth = 3, label = "N(t)")
+
+axislegend(ax)
+
+fig
+
+min(sol_array[1:5, :]...)
+# min(sol_array[4, :]...)
+```
+
+```{julia}
 function run_sir_sim(; u₀, tspan, p)

### 2023-03-06 notebooks/Julia/simulation-detection.qmd
Convert back to Gillespie Direct Using BenchmarkTools indicated that Gillespie Direct was quicker than my implementation of Tau-Leaping or SDEProblem, and solves the main issue I was having of jumps going below 0. Look into performance again if necessary

--- 2023-03-06 notebooks/Julia/simulation-detection.qmd
@@
 using Catalyst, Random, Statistics, DataFrames, CairoMakie, ProgressMeter
+using BenchmarkTools
 
@@
 ```{julia}
-function rate(out, u, p, t)
-    out[1] = p[1] * u[1] * u[2] / (u[1] + u[2] + u[3])           # β * S * I
-    out[2] = p[2] * u[2]                      # ν * I
-    # out[3] = p[3] * u[1]                    # μ * N
-    # out[4] = p[3] * u[1]                    # μ * S
-    # out[5] = p[3] * u[2]                    # μ * I
-    # out[6] = p[3] * u[3]                    # μ * R
-    nothing
-end
+tmax = 2.0
+δt = 0.1/365
+tspan = (0.0, tmax)
+tlength = length(tspan[1]:δt:tspan[2])
+Random.seed!(1234);
+
+u₀ = [999.0, 10.0, 0.0, 0.0]
+
+μ = 1
+β = 1000
+σ = 365/8
+γ = 365/5
+p = (β, σ, γ, μ)
 ```
@@
 ```{julia}
-c = zeros(3, 2)
-# S + I --> I
-c[1, 1] = -1        # S -> S - 1
-c[2, 1] = 1         # I -> I + 1
-
-# I --> R
-c[2, 2] = -1        # I -> I - 1
-c[3, 2] = 1         # R -> R + 1
-
-# S --> 2X
-# c[1, 3] = 1         # S -> S + 1
-
-# # S --> 0X
-# c[1, 4] = -1        # S -> S - 1
-
-# # I --> 0Y
-# c[2, 5] = -1        # I -> I - 1
-
-# # R --> 0Z
-# c[3, 6] = -1        # R -> R - 1
-
-function change!(du, u, p, t, counts, mark)
-    mul!(du, c, counts)
+expose_rate(u, p, t) = p[1] * u[1] * u[2] / (u[1] + u[2] + u[3] + u[4])  # β * S * I / N
+function expose_affect!(integrator)
+    integrator.u[1] -= 1         # S -> S - 1
+    integrator.u[2] += 1         # E -> E + 1
     nothing
 end
-```
+expose_jump = ConstantRateJump(expose_rate, expose_affect!)
 
-```{julia}
-rj = RegularJump(rate, change!, 2)
-```
+infec_rate(u, p, t) = p[2] * u[2]   # σ * E
+function infec_affect!(integrator)
+    integrator.u[2] -= 1         # E -> E - 1
+    integrator.u[3] += 1         # I -> I + 1
+    nothing
+end
+infec_jump = ConstantRateJump(infec_rate, infec_affect!)
 
-```{julia}
-tmax = 100.0
-δt = 0.001
-tspan = (0.0, tmax)
-tlength = length(tspan[1]:δt:tspan[2])
-Random.seed!(1234);
-```
+recov_rate(u, p, t) = p[3] * u[3]   # γ * I
+function recov_affect!(integrator)
+    integrator.u[3] -= 1         # I -> I - 1
+    integrator.u[4] += 1         # R -> R + 1
+    nothing
+end
+recov_jump = ConstantRateJump(recov_rate, recov_affect!)
 
-```{julia}
-# R₀ = 2
-# γ = 1/8
-μ = 0.001
-# β = R₀ * (γ + μ)
-# ζ = 0.001
-# p = (β, γ, μ, ζ)
+birth_rate(u, p, t) = p[4] * (u[1] + u[2] + u[3] + u[4]) # μ * N
+function birth_affect!(integrator)
+    integrator.u[1] += 1         # S -> S + 1
+    nothing
+end
+birth_jump = ConstantRateJump(birth_rate, birth_affect!)
 
-u₀ = [1000.0, 50.0, 0.0]
+S_death_rate(u, p, t) = p[4] * u[1] # μ * S
+function S_death_affect!(integrator)
+    integrator.u[1] -= 1        # S -> S - 1
+end
+S_death_jump = ConstantRateJump(S_death_rate, S_death_affect!)
 
-β = 0.1/1000.0
-γ = 0.01
-μ = 0.001
-p = (β, γ, μ)
+E_death_rate(u, p, t) = p[4] * u[2] # μ * E
+function E_death_affect!(integrator)
+    integrator.u[2] -= 1        # E -> E -1
+end
+E_death_jump = ConstantRateJump(E_death_rate, E_death_affect!)
 
+I_death_rate(u, p, t) = p[4] * u[3] # μ * I
+function I_death_affect!(integrator)
+    integrator.u[3] -= 1        # I -> I - 1
+end
+I_death_jump = ConstantRateJump(I_death_rate, I_death_affect!)
+
+R_death_rate(u, p, t) = p[4] * u[4] # μ * R
+function R_death_affect!(integrator)
+    integrator.u[4] -= 1        # R -> R - 1 
+end
+R_death_jump = ConstantRateJump(R_death_rate, R_death_affect!)
+
+gillespie_jumps = [
+    expose_jump, infec_jump, recov_jump, birth_jump, S_death_jump, E_death_jump,
+    I_death_jump, R_death_jump
+]
 ```
@@
 prob = DiscreteProblem(u₀, tspan, p)
-jump_prob = JumpProblem(prob, Direct(), rj)
-sol = solve(jump_prob, TauLeaping(); dt = δt)
-sol_array = zeros(4, length(sol))
-sol_array[1:3, :] = Array(sol)
-sol_array[4, :] = sum(sol_array[1:3, :], dims = 1)
+jump_prob = JumpProblem(prob, Direct(), gillespie_jumps...)
+sol = solve(jump_prob, SSAStepper())
+sol_array = zeros(5, length(sol))
+sol_array[1:4, :] = Array(sol)
+sol_array[5, :] = sum(sol_array[1:4, :], dims = 1)
 ```
@@
 ```{julia}
-f = Figure()
+fig = Figure()
 ax = Axis(
-    f[1, 1],
+    fig[1, 1],
     xlabel = "Time",
@@
 lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "S(t)")
-lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "I(t)")
-lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "R(t)")
-lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "N(t)")
+lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "E(t)")
+lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "I(t)")
+lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "R(t)")
+lines!(ax, sol.t, sol_array[5, :], linewidth = 3, label = "N(t)")
 
@@
 
-f
+fig
 ```

### 2023-03-06 notebooks/Julia/simulation-detection.qmd
Update simulation functions to account for E and N compartments and update to median Update calculation of average to use median to specify dims, as is likely quicker than my loop implementation

@@
     prob = DiscreteProblem(u₀, tspan, p)
-    jump_prob = JumpProblem(prob, Direct(), rj)
-    sol_jump = solve(jump_prob, TauLeaping(); dt = δt)
-    return sol_jump(0:δt:tmax);
+    jump_prob = JumpProblem(prob, Direct(), gillespie_jumps..., save_positions = (false, false))
+    jump_sol = solve(jump_prob, SSAStepper(), saveat = δt)
+    return jump_sol;
 end
@@
 ```{julia}
-function create_sim_array(; out_jump)
-    test_array[1:3, :] = Array(out_jump)
-    test_array[4, :] .= out_jump.t
+out_jump = run_sir_sim(u₀ = u₀, tspan = tspan, p = p)
+Array(out_jump)
+```
 
-    return test_array
+```{julia}
+function create_sim_array!(; out_jump)
+    sir_array[1:4, :] = Array(out_jump)
+    sir_array[5, :] = sum(sir_array[1:4, :], dims = 1) 
+    sir_array[6, :] .= out_jump.t
+
+    return nothing
 end
@@
 
-sir_array = zeros(Float64, 4, tlength)
-all_sims_array = zeros(Float64, 4, tlength, nsims)
-sim_means = zeros(Float64, 3, tlength)
+sir_array = zeros(6, tlength + 1)
+
+all_sims_array = zeros(Float64, 6, tlength + 1, nsims)
+sim_means = zeros(Float64, 5, tlength + 1)
 
@@
     out_jump = run_sir_sim(u₀ = u₀, tspan = tspan, p = p)
-    sir_array = create_sim_array(out_jump = out_jump)
+    create_sim_array!(out_jump = out_jump)
 
@@
 
-    lines!(ax, sir_array[4, :], sir_array[1, :], linewidth = 4, color = ("dodgerblue4", 0.01))
-    lines!(ax, sir_array[4, :], sir_array[2, :], linewidth = 4, color = ("firebrick3", 0.01))
-    lines!(ax, sir_array[4, :], sir_array[3, :], linewidth = 4, color = ("chocolate2", 0.01))
+    lines!(ax, sir_array[6, :], sir_array[1, :], linewidth = 4, color = ("dodgerblue4", 0.02))
+    lines!(ax, sir_array[6, :], sir_array[2, :], linewidth = 4, color = ("firebrick3", 0.02))
+    lines!(ax, sir_array[6, :], sir_array[3, :], linewidth = 4, color = ("chocolate2", 0.02)) 
+    lines!(ax, sir_array[6, :], sir_array[4, :], linewidth = 4, color = ("green", 0.02)) 
+    lines!(ax, sir_array[6, :], sir_array[5, :], linewidth = 4, color = ("purple", 0.02))
 
@@
 
-@showprogress for sim in 1:tlength, outcome in 1:3
-    sim_means[outcome, sim] = mean(all_sims_array[outcome, sim, :])
+@showprogress for outcome in 1:5
+    sim_means[outcome, :] = median(all_sims_array[outcome, :, :], dims = 2)
 end
 
-lines!(ax, sir_array[4, :], sim_means[1, :], linewidth = 3, color = "dodgerblue4")
-lines!(ax, sir_array[4, :], sim_means[2, :], linewidth = 3, color = "firebrick3")
-lines!(ax, sir_array[4, :], sim_means[3, :], linewidth = 3, color = "chocolate2")
+lines!(ax, sir_array[6, :], sim_means[1, :], linewidth = 3, color = "dodgerblue4", label = "S(t)")
+lines!(ax, sir_array[6, :], sim_means[2, :], linewidth = 3, color = "firebrick3", label = "E(t)")
+lines!(ax, sir_array[6, :], sim_means[3, :], linewidth = 3, color = "chocolate2", label = "I(t)")
+lines!(ax, sir_array[6, :], sim_means[4, :], linewidth = 3, color = "green", label = "R(t)")
+lines!(ax, sir_array[6, :], sim_means[5, :], linewidth = 3, color = "purple", label = "N(t)")
+
+axislegend(ax)
 

### 2023-03-07 notebooks/Julia/simulation-detection.qmd
Try using AlgebraOfGraphics for single SIR plot Wanted to try using AOG as uses ggplot2 style syntax. Seems like it's pretty doable for a simple plot, but would require I create a large dataframe for the multiple simulations, rather than using an array. Given Julia's col-major, I'm not sure what implications this has for speed, but I'm guessing this doesn't have that big an effect.

@@
 using JumpProcesses, StochasticDiffEq, DifferentialEquations, LinearAlgebra
-using Catalyst, Random, Statistics, DataFrames, CairoMakie, ProgressMeter
-using BenchmarkTools
+using Catalyst, Random, Statistics, DataFrames, CairoMakie, AlgebraOfGraphics
+using BenchmarkTools, ProgressMeter
 
 CairoMakie.activate!()
+set_aog_theme!()
 ```
@@
 sol = solve(jump_prob, SSAStepper())
-sol_array = zeros(5, length(sol))
+sol_array = zeros(6, length(sol))
 sol_array[1:4, :] = Array(sol)
 sol_array[5, :] = sum(sol_array[1:4, :], dims = 1)
+sol_array[6, :] = sol.t;
 ```
@@
 ```{julia}
-fig = Figure()
-ax = Axis(
-    fig[1, 1],
-    xlabel = "Time",
-    ylabel = "Number"
-)
+sol_df = DataFrame(Tables.table(sol_array'))
 
-lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "S(t)")
-lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "E(t)")
-lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "I(t)")
-lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "R(t)")
-lines!(ax, sol.t, sol_array[5, :], linewidth = 3, label = "N(t)")
+rename!(sol_df, :Column1 => :S, :Column2 => :E, :Column3 => :I, :Column4 => :R, :Column5 => :N, :Column6 => :time)
 
-axislegend(ax)
+stacked_df = stack(sol_df, [:S, :E, :I, :R, :N])
+```
 
-fig
+```{julia}
+colors = ["dodgerblue4", "firebrick3", "chocolate2", "green", "purple"]
+sir_plot = data(stacked_df) *
+    mapping(
+        :time => "Time (Years)", :value => "Number",
+        color = :variable => sorter("S", "E", "I", "R", "N") => "State"
+    ) *
+    visual(Lines, linewidth = 4)
+draw(sir_plot; palettes = (; color = colors))
 ```

### 2023-03-09 notebooks/Julia/simulation-detection.qmd
Update to Julia 1.9 release candidate 1

@@
   cache: true
-jupyter: julia-1.8
+jupyter: julia-1.9
 ---

### 2023-03-09 notebooks/Julia/simulation-detection.qmd
Start to add time-varying beta (sine-wave scaled) Need to look into what's going on as sometimes the population increases exponentially, but the VariableRateJump should only affect the S -> E process, not the births.

@@
 using JumpProcesses, StochasticDiffEq, DifferentialEquations, LinearAlgebra
-using Catalyst, Random, Statistics, DataFrames, CairoMakie, AlgebraOfGraphics
+using Catalyst, Random, Statistics, DataFrames, DataFramesMeta
+using CairoMakie, AlgebraOfGraphics
 using BenchmarkTools, ProgressMeter
@@
 ```{julia}
-tmax = 2.0
+tmax = 10.0
 δt = 0.1/365
@@
 γ = 365/5
-p = (β, σ, γ, μ)
+ε = 100
+p = (β, σ, γ, μ, ε)
 ```
@@
 ```{julia}
-expose_rate(u, p, t) = p[1] * u[1] * u[2] / (u[1] + u[2] + u[3] + u[4])  # β * S * I / N
+x = 0:0.1:10
+# 0.5 .* (sin.(2π .* x) .+ 1)
+fig = lines(x, 0.5 .* (sin.(pi .* x ./ 2) .+ 1))
+vlines!(collect(0:1:10); color = (:red, 0.5), linestyle = :dash)
+fig
+```
+
+```{julia}
+# 0.5 .* (sin.(2π .* x) .+ 1)
+expose_rate(u, p, t) = p[1] * 0.5 * (sin(2π * t) + 1) * u[1] * u[2]/ (u[1] + u[2] + u[3] + u[4])  # β * ϕ * S * I / N
 function expose_affect!(integrator)
@@
 end
-expose_jump = ConstantRateJump(expose_rate, expose_affect!)
+
+# We require that rate1(u,p,s) <= urate(u,p,s)
+# for t <= s <= t + rateinterval(u,p,t)
+rateinterval(u, p, t) = typemax(t)
+urate(u, p, t) = p[1] * u[1] * u[2]/ (u[1] + u[2] + u[3] + u[4])
+
+# Optionally, we can give a lower bound over the same interval.
+# This may boost computational performance.
+lrate(u, p, t) = p[1] * u[1] * u[2]/ (u[1] + u[2] + u[3] + u[4])
+
+# now we construct the bounded VariableRateJump
+expose_jump = VariableRateJump(expose_rate, expose_affect!; lrate, urate, rateinterval)
+
+# import_expose_rate(u, p, t) = p[5]
+# function import_expose_affect!(integrator)
+#     integrator.u[3] += 1         # I -> I + 1
+#     nothing
+# end
+# import_expose_jump = ConstantRateJump(import_expose_rate, import_expose_affect!)
 
@@
 ```{julia}
+# Constant jumps are ordered ahead of VariableJumps, but otherwise retain the ordering
+dep_graph = [
+    [1], [2], [3], [4], [5], [6], [7],
+    [1, 2, 3, 4, 5, 6, 7, 8] # Exposure jump
+    ]
+
 prob = DiscreteProblem(u₀, tspan, p)
-jump_prob = JumpProblem(prob, Direct(), gillespie_jumps...)
+jump_prob = JumpProblem(prob, Coevolve(), gillespie_jumps...; dep_graph)
 sol = solve(jump_prob, SSAStepper())
 sol_array = zeros(6, length(sol))
-sol_array[1:4, :] = Array(sol)
+sol_array[1:4, :] = Array(sol)[1:4, :]
 sol_array[5, :] = sum(sol_array[1:4, :], dims = 1)
-sol_array[6, :] = sol.t;
+sol_array[6, :] = sol.t
 ```
@@
     visual(Lines, linewidth = 4)
-draw(sir_plot; palettes = (; color = colors))
+
+draw(
+    sir_plot;
+    palettes = (; color = colors),
+    axis = (; limits = ((0.0, 2.0), nothing))
+    )
 ```
@@
     prob = DiscreteProblem(u₀, tspan, p)
-    jump_prob = JumpProblem(prob, Direct(), gillespie_jumps..., save_positions = (false, false))
+    jump_prob = JumpProblem(
+        prob, Coevolve(), gillespie_jumps..., save_positions = (false, false);
+        dep_graph
+        )
     jump_sol = solve(jump_prob, SSAStepper(), saveat = δt)
@@
 
-axislegend(ax)
+Legend(sir_sim_plot[1, 2], ax, "State")
 

### 2023-03-09 notebooks/Julia/simulation-detection.qmd
Update dependency graph - doesn't work as produces too many jumps and exploding pop Look into using unbounded VariableJump that doesn't require specifying the dependency graph

@@
 x = 0:0.1:10
-# 0.5 .* (sin.(2π .* x) .+ 1)
-fig = lines(x, 0.5 .* (sin.(pi .* x ./ 2) .+ 1))
+fig = lines(x, 0.5 .* (sin.(2pi .* x) .+ 1))
 vlines!(collect(0:1:10); color = (:red, 0.5), linestyle = :dash)
@@
 # for t <= s <= t + rateinterval(u,p,t)
-rateinterval(u, p, t) = typemax(t)
 urate(u, p, t) = p[1] * u[1] * u[2]/ (u[1] + u[2] + u[3] + u[4]) 
-
 # Optionally, we can give a lower bound over the same interval.
 # This may boost computational performance.
-lrate(u, p, t) = p[1] * u[1] * u[2]/ (u[1] + u[2] + u[3] + u[4])
+lrate(u, p, t) = 0
+
+# rateinterval needs to be between lrate <= ri <= urate
+rateinterval(u, p, t) = 1 / (2 * urate(u, p, t))
 
@@
 dep_graph = [
-    [1], [2], [3], [4], [5], [6], [7],
-    [1, 2, 3, 4, 5, 6, 7, 8] # Exposure jump
+    [1, 8],    # infec depends on states changed by itself and exposure jump
+    [2],    # recov depends on states changed by itself
+    [3],    # birth depends on states changed by itself
+    [4, 8],    # S_death depends on states changed by itself and exposure jump
+    [5, 8],    # E_death depends on states changed by itself and exposure jump
+    [6],    # I_death depends on states changed by itself
+    [7],    # R_death depends on states changed by itself and exposure jump
+    [1, 3, 4, 5, 8]    # Exposure jump depends on states changed by itself, birth, S_death, E_death, and infection jumps
     ]
@@
 end
-```
-
-```{julia}
-out_jump = run_sir_sim(u₀ = u₀, tspan = tspan, p = p)
-Array(out_jump)
 ```

### 2023-03-09 notebooks/Julia/Gillespie-basic.jl
Add basic gillespie and ODE SIR models for debugging purposes Translate Tiffany's code to calculate beta into Julia. Currently works, but it looks like it only works for a 1x1 contact matrix i.e. when there is only one group

@@ -1,0 +1,93 @@
+"""
+This is a basic Gillespie simulation of an SIR model.
+It uses the JumpProcesses.jl package and the Direct Gillespie method.
+All jumps are constant rate jumps and are manually defined using JumpProcesses, not using the ModelingToolkit or Catalyst interfaces.
+"""
+
+using DrWatson
+@quickactivate "OutbreakDetection"
+
+using JumpProcesses, Statistics, DataFrames, DataFramesMeta, LinearAlgebra
+using CairoMakie, AlgebraOfGraphics
+
+CairoMakie.activate!()
+set_aog_theme!()
+
+u₀ = [999, 1, 0]
+tspan = (0.0, 250.0)
+
+function calculatebeta(R₀, 𝐂, γ, μ, totalPop)
+    # compute the eignevalues of -F*V^(-1)
+    n1, n2 = size(𝐂)
+
+    # create F (transmission matrix) and V (transition matrix)
+    𝐅 = zeros(n1, n1)
+    𝐕 = Diagonal(repeat([-(ν + μ)], n1))
+    N = sum(totalPop)
+
+    for jvals in 1:n1
+        𝐅[jvals, :] = (𝐂[jvals, :]) * totalPop[jvals]
+    end
+
+    𝐅𝐕⁻¹ = -𝐅 * inv(𝐕)
+
+    myEig = eigvals(𝐅𝐕⁻¹)
+    largestEig = maximum(myEig)
+
+    # @assert largestEig.imag == 0.0 "largest eigenvalue is not real"
+    β = R₀ / largestEig
+    return β
+end
+
+R₀ = 2
+ν = 1/8
+μ = 0
+
+
+calculatebeta(R₀, ones(1, 1), ν, μ, sum(u₀)) == R₀ * ν / sum(u₀)
+β = calculatebeta(R₀, ones(1, 1), ν, μ, sum(u₀))
+p = (β, ν)
+
+infec_rate(u, p, t) = p[1] * u[1] * u[2]  # β*S*I
+function infec_affect!(integrator)
+    integrator.u[1] -= 1         # S -> S - 1
+    integrator.u[2] += 1         # I -> I + 1
+    nothing
+end
+infec_jump = ConstantRateJump(infec_rate, infec_affect!)
+
+recov_rate(u, p, t) = p[2] * u[2]         # ν*I
+function recov_affect!(integrator)
+    integrator.u[2] -= 1        # I -> I - 1
+    integrator.u[3] += 1        # R -> R + 1
+    nothing
+end
[…]

### 2023-03-09 notebooks/Julia/ode.jl

@@ -1,0 +1,57 @@
+"""
+This script just performs a simple SIR model using an ODE solver.
+"""
+
+using DrWatson
+@quickactivate "OutbreakDetection"
+
+using DifferentialEquations, Statistics, DataFrames, DataFramesMeta, LinearAlgebra
+using CairoMakie, AlgebraOfGraphics
+
+CairoMakie.activate!()
+set_aog_theme!()
+
+u₀ = [999.0, 1.0, 0.0]
+tspan = (0.0, 250.0)
+
+function sir!(du, u, p, t)
+    (β, γ) = p
+    (S, I, R) = u
+    du[1] = -β * S * I
+    du[2] = β * S * I - γ * I
+    du[3] = γ * I
+    nothing
+end
+
+
+R₀ = 2
+γ = 1/8
+β = R₀ * γ / sum(u₀)
+
+p = [β, γ]
+
+prob = ODEProblem(sir!, u₀, tspan, p)
+sol = solve(prob)
+
+
+sol_df = @chain Tables.table(sol') begin
+    DataFrame(_)
+    rename!(:Column1 => :S, :Column2 => :I, :Column3 => :R)
+    @rtransform! :N = :S + :I + :R
+    hcat(_, DataFrame(time = sol.t))
+    stack(_, [:S, :I, :R, :N], variable_name = :State, value_name = :Number)
+end
+
+colors = ["dodgerblue4", "firebrick3", "chocolate2", "purple"]
+sir_plot = data(sol_df) *
+    mapping(
+        :time => "Time (days)", :Number,
+        color = :State => sorter("S", "I", "R", "N")
+    ) *
+    visual(Lines, linewidth = 4)
+
+draw(
+    sir_plot;
+    palettes = (; color = colors),
+    # axis = (; limits = ((0.0, 2.0), nothing))
+    )

### 2023-08-16 notebooks/Julia/simulation-detection.qmd
Add basic gillespie and ODE SIR models for debugging purposes Translate Tiffany's code to calculate beta into Julia. Currently works, but it looks like it only works for a 1x1 contact matrix i.e. when there is only one group
Rename beta vars to use text rather than symbols
Want to be clearer about meaning and some editors etc don't render unicode characters

@@
 μ = 1
-β = 1000
+beta = 1000
 σ = 365/8
@@
 ε = 100
-p = (β, σ, γ, μ, ε)
+p = (beta, σ, γ, μ, ε)
 ```
@@
 ```{julia}
-x = 0:0.1:10
+x = 0:0.01:2
 fig = lines(x, 0.5 .* (sin.(2pi .* x) .+ 1))
-vlines!(collect(0:1:10); color = (:red, 0.5), linestyle = :dash)
+vlines!(collect(0:1:2); color = (:red, 0.5), linestyle = :dash)
 fig
@@
 # 0.5 .* (sin.(2π .* x) .+ 1)
-expose_rate(u, p, t) = p[1] * 0.5 * (sin(2π * t) + 1) * u[1] * u[2]/ (u[1] + u[2] + u[3] + u[4])  # β * ϕ * S * I / N
+expose_rate(u, p, t) = p[1] * 0.5 * (sin(2π * t) + 1) * u[1] * u[2]/ (u[1] + u[2] + u[3] + u[4])  # beta * ϕ * S * I / N
 function expose_affect!(integrator)

### 2023-08-16 scripts/ensemble-sim.jl

@@ -15,7 +15,7 @@ u₀_prop_map = [
 ]
 dt_vec = [1.0]
 tmax_vec = [365.0 * 100]
-β_force_vec = collect(0.0:0.1:0.4)
+beta_force_vec = collect(0.0:0.1:0.4)
 μ_min = 5
 μ_max = 20
 μ_step = 5.0
@@ -43,7 +43,7 @@ base_param_dict = @dict(
     nsims = nsims_vec,
     dt = dt_vec,
     tmax = tmax_vec,
-    β_force = β_force_vec,
+    beta_force = beta_force_vec,
     births_per_k = μ_vec,
     seed = seed,
 )

### 2023-08-16 scripts/single-sim.jl

@@ -28,18 +28,18 @@ R₀ = 10.0
 σ = 1 / latent_per
 γ = 1 / dur_inf
 μ = 1 / (62.5 * 365)
-# β₀ is the average transmission rate
-β₀ = calculate_beta(R₀, γ, μ, 1, N)
-# Adjust the scale of the seasonal variation in infectivity i.e. β₁ scales the amplitude of cosine function
-β₁ = 0.2
+# beta_mean is the average transmission rate
+beta_mean = calculate_beta(R₀, γ, μ, 1, N)
+# Adjust the scale of the seasonal variation in infectivity i.e. beta_force scales the amplitude of cosine function
+beta_force = 0.2
 ε = (1.06 * μ * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
-p = (β₀, β₁, σ, γ, μ, ε, R₀)
+p = (beta_mean, beta_force, σ, γ, μ, ε, R₀)
 
 Random.seed!(1234)
 
 #%%
-seir_array, change_array, jump_array, β_arr = seir_mod(
-    u₀, p, trange; retβarr = true, type = "stoch"
+seir_array, change_array, jump_array, beta_arr = seir_mod(
+    u₀, p, trange; retbetaarr = true, type = "stoch"
 );
 
 seir_df = create_sir_df(

### 2023-08-16 scripts/single-sim_bifurcation.jl

@@ -29,7 +29,7 @@ prog = Progress(n_μs)
     change = @view bifurc_μ_change_arr[:, :, k]
     jump = @view bifurc_μ_jump_arr[:, :, k]
 
-    params = (β₀, β₁, σ, γ, μ_run, ε, R₀)
+    params = (beta_mean, beta_force, σ, γ, μ_run, ε, R₀)
     seir_mod!(seir, change, jump,
         u₀, params, trange; dt = τ, type = "det",
     )
@@ -69,81 +69,81 @@ end
 bifurc_μ_fig
 
 #%%
-β₁_min = 0.0
-β₁_max = 1.0
-β₁_step = 0.02
-n_β₁s = length(β₁_min:β₁_step:β₁_max)
-β₁_vec = zeros(Float64, n_β₁s)
-β₁_vec .= collect(β₁_min:β₁_step:β₁_max)
+beta_force_min = 0.0
+beta_force_max = 1.0
+beta_force_step = 0.02
+n_beta_forces = length(beta_force_min:beta_force_step:beta_force_max)
+beta_force_vec = zeros(Float64, n_beta_forces)
+beta_force_vec .= collect(beta_force_min:beta_force_step:beta_force_max)
 
-bifurc_β₁_seir_arr = zeros(Float64, size(u₀, 1), tlength, n_β₁s);
-bifurc_β₁_change_arr = zeros(Float64, size(u₀, 1), tlength, n_β₁s);
-bifurc_β₁_jump_arr = zeros(Float64, 9, tlength, n_β₁s);
+bifurc_beta_force_seir_arr = zeros(Float64, size(u₀, 1), tlength, n_beta_forces);
+bifurc_beta_force_change_arr = zeros(Float64, size(u₀, 1), tlength, n_beta_forces);
+bifurc_beta_force_jump_arr = zeros(Float64, 9, tlength, n_beta_forces);
 
 μ = 50 / (1000 * 365)
 ε = 1.06 * μ * (R₀ - 1) / sqrt(N)
 
-@showprogress for (k, β₁) in pairs(β₁_vec)
-    seir = @view bifurc_β₁_seir_arr[:, :, k]
-    change = @view bifurc_β₁_change_arr[:, :, k]
-    jump = @view bifurc_β₁_jump_arr[:, :, k]
+@showprogress for (k, beta_force) in pairs(beta_force_vec)
+    seir = @view bifurc_beta_force_seir_arr[:, :, k]
+    change = @view bifurc_beta_force_change_arr[:, :, k]
[…]

### 2023-08-16 scripts/single-sim_plots.jl

@@ -66,10 +66,10 @@ end
 end
 
 #%%
-@chain DataFrame(Tables.table(β_arr)) begin
+@chain DataFrame(Tables.table(beta_arr)) begin
     hcat(trange, _)
-    rename!([:time, :β_t])
-    stack(_, Not("time"); variable_name = :β, value_name = :Number)
+    rename!([:time, :beta_t])
+    stack(_, Not("time"); variable_name = :beta, value_name = :Number)
     data(_) *
     mapping(
         :time => (t -> t / 365) => "Time (years)", :Number;

### 2023-08-16 src/Julia/SEIR-model.jl

@@ -12,28 +12,28 @@ using Distributions
 using Random
 
 """
-    calculate_beta_amp(β_mean, β_force, t)
+    calculate_beta_amp(beta_mean, beta_force, t)
 
-Calculate the amplitude of the transmission rate β as a function of time.
-`β_mean` is the mean transmission rate, `β_force` is the amplitude of the cosine function.
+Calculate the amplitude of the transmission rate beta as a function of time.
+`beta_mean` is the mean transmission rate, `beta_force` is the amplitude of the cosine function.
 """
-function calculate_beta_amp(β_mean, β_force, t)
-    return β_mean * (1 + β_force * cos(2pi * t / 365))
+function calculate_beta_amp(beta_mean, beta_force, t)
+    return beta_mean * (1 + beta_force * cos(2pi * t / 365))
 end
 
 """
-    seir_mod(u, p, trange; retβarr = false, type = "stoch")
+    seir_mod(u, p, trange; retbetaarr = false, type = "stoch")
 
 A Tau-leaping SEIR model with commuter style importers.
-the model can return an array of transmission rates if `retβarr = true`, and can be set to be deterministic or stochastic.
+the model can return an array of transmission rates if `retbetaarr = true`, and can be set to be deterministic or stochastic.
 
 ```julia-repl
-seir_array, change_array, jump_array, β_arr = seir_mod(
-    u₀, p, trange; retβarr = true, type = "stoch"
+seir_array, change_array, jump_array, beta_arr = seir_mod(
+    u₀, p, trange; retbetaarr = true, type = "stoch"
 );
 ```
 """
-function seir_mod(u, p, trange; retβarr = false, type = "stoch", seed = 1234)
+function seir_mod(u, p, trange; retbetaarr = false, type = "stoch", seed = 1234)
     tlength = length(trange)
     dt = step(trange)
 
@@ -43,7 +43,7 @@ function seir_mod(u, p, trange; retβarr = false, type = "stoch", seed = 1234)
 
     jump_arr = zeros(Float64, 9, tlength)
 
-    if retβarr == true
+    if retbetaarr == true
         beta_arr = zeros(Float64, tlength)
         seir_mod!(
             state_arr, change_arr, jump_arr, beta_arr, u, p, trange; dt = dt,
@@ -92,11 +92,11 @@ function seir_mod!(
[…]

### 2023-08-16 src/Julia/cleaning-functions.jl

@@ -73,10 +73,10 @@ end
 function create_sir_beta_dfs(sol, states = [:S, :I, :R])
     state_df = create_sir_df(sol, states)
 
-    beta_df = select(state_df, [:time, :β])
+    beta_df = select(state_df, [:time, :beta])
     unique!(beta_df)
 
-    select!(state_df, Not(:β))
+    select!(state_df, Not(:beta))
 
     return state_df, beta_df
 end
@@ -129,9 +129,9 @@ function create_sir_all_sims_array(ensemble_sol::EnsembleSolution, nsims::Int)
 end
 
 function create_sir_all_sims_array!(
-    ensemble_sol::EnsembleSolution, nsims::Int, β::Bool
+    ensemble_sol::EnsembleSolution, nsims::Int, beta::Bool
 )
-    if β == false
+    if beta == false
         return create_sir_all_sims_array!(ensemble_sol, nsims)
     end
 

### 2023-08-16 src/Julia/ensemble-functions.jl

@@ -23,7 +23,7 @@ function run_ensemble_jump_prob(params_dict; prog = prog)
                 "r_$(p[:u₀_prop][:r])",
                 "nsims_$(p[:nsims])",
                 "births_per_k_$(p[:births_per_k])",
-                "beta_force_$(p[:β_force])",
+                "beta_force_$(p[:beta_force])",
                 "tmax_$(p[:tmax])",
                 "deltat_$(p[:dt])",
             );
@@ -32,7 +32,7 @@ function run_ensemble_jump_prob(params_dict; prog = prog)
                 p;
                 allowedtypes = (Symbol, Dict, String, Real),
                 accesses = [
-                    :N, :u₀_prop, :nsims, :tmax, :dt, :births_per_k, :β_force
+                    :N, :u₀_prop, :nsims, :tmax, :dt, :births_per_k, :beta_force
                 ],
                 expand = ["u₀_prop"],
                 sort = false,
@@ -47,7 +47,7 @@ end
     run_jump_prob(param_dict)
 """
 function run_jump_prob(param_dict)
-    @unpack N, u₀_prop, transmission_p, time_p, nsims, dt, β_force,
+    @unpack N, u₀_prop, transmission_p, time_p, nsims, dt, beta_force,
     births_per_k, seed = param_dict
     @unpack s, e, i, r = u₀_prop
     @unpack R₀, σ, γ = transmission_p
@@ -57,10 +57,10 @@ function run_jump_prob(param_dict)
     u0_dict = Dict(zip([:S, :E, :I, :R, :N], u₀))
 
     μ = births_per_k / (1000 * 365)
-    β₀ = calculate_beta(R₀, γ, μ, 1, N)
+    beta_mean = calculate_beta(R₀, γ, μ, 1, N)
     ε = (1.06 * μ * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
 
-    p = (β₀, β_force, σ, γ, μ, ε, R₀)
+    p = (beta_mean, beta_force, σ, γ, μ, ε, R₀)
 
     ensemble_seir_arr = zeros(Int64, size(u₀, 1), tlength, nsims)
     ensemble_change_arr = zeros(Int64, size(u₀, 1), tlength, nsims)
@@ -89,7 +89,7 @@ function summarize_ensemble_jump_prob(params_dict; prog = prog)
                 "r_$(p[:u₀_prop][:r])",
[…]

### 2023-08-16 src/Julia/transmission-functions.jl

@@ -9,7 +9,7 @@ using FLoops
 """
     calculate_beta(R₀, γ, μ, C, pop_matrix)
 
-Calculate the value β for a given set of parameters and contact matrix.
+Calculate the value beta for a given set of parameters and contact matrix.
 
 ```jldoctest
 julia> calculate_beta(2.0, 1 / 8, 0.0, ones(1, 1), [1000])
@@ -33,9 +33,9 @@ function calculate_beta(
 
     FV⁻¹ = F * inv(V)
     eigenvals, eigenvectors = eigen(FV⁻¹)
-    β = R₀ / maximum(real(eigenvals))
+    beta = R₀ / maximum(real(eigenvals))
 
-    return β
+    return beta
 end
 
 function calculate_beta(R₀, γ, μ, C, pop_matrix)
@@ -64,7 +64,7 @@ function calculate_beta(
 
     F = C .* pop_matrix
     # F = substitute(Jac, Dict(γ => 0.0, μ => 0.0))
-    V = substitute(Jac, Dict(β => 0.0))
+    V = substitute(Jac, Dict(beta => 0.0))
     FV⁻¹ = F * -inv(V)
     eigenvals =
         convert.(
@@ -95,7 +95,7 @@ function calculate_beta(
 end
 
 """
-    calculateR0(β, γ, μ, C, pop_matrix)
+    calculateR0(beta, γ, μ, C, pop_matrix)
 
 Calculate the basic reproduction number R₀ for a given set of parameters and contact matrix.
 
@@ -111,7 +111,7 @@ julia> calculateR0(0.00025, 1 / 8, 0.0, ones(1, 1), [1000])
 * * *
 """
 function calculateR0(
-    β::T, γ::T, μ::T, C::Array{T}, pop_matrix::Array{T}
+    beta::T, γ::T, μ::T, C::Array{T}, pop_matrix::Array{T}
 ) where {T<:AbstractFloat}
     size(C, 1) == size(C, 2) ? nothing : error("C must be square")
     if size(C, 1) == size(pop_matrix, 1)
@@ -120,7 +120,7 @@ function calculateR0(
         error("C and pop_matrix must have the same number of rows")
     end
 
-    B = β * C
+    B = beta * C
 
     F = B .* pop_matrix
     V = Diagonal(repeat([γ + μ], size(C, 1)))
@@ -133,9 +133,9 @@ function calculateR0(
     return R₀
 end
 
-function calculateR0(β, γ, μ, C, pop_matrix)
[…]

### 2023-08-16 notebooks/Julia/simulation-detection.qmd
Rename sigma vars for clarity

@@
 beta = 1000
-σ = 365/8
+sigma = 365/8
 γ = 365/5
 ε = 100
-p = (beta, σ, γ, μ, ε)
+p = (beta, sigma, γ, μ, ε)
 ```
@@
 
-infec_rate(u, p, t) = p[2] * u[2]   # σ * E
+infec_rate(u, p, t) = p[2] * u[2]   # sigma * E
 function infec_affect!(integrator)

### 2023-08-16 scripts/ensemble-sim.jl

@@ -27,9 +27,9 @@ seed = 1234
 latent_per_days = 8
 dur_inf_days = 5
 R₀ = 10.0
-σ = 1 / latent_per_days
+sigma = 1 / latent_per_days
 γ = 1 / dur_inf_days
-transmission_p = EnsembleTransmissionParameters(R₀, σ, γ)
+transmission_p = EnsembleTransmissionParameters(R₀, sigma, γ)
 tmin = 0.0
 tmax = 365.0 * 100
 tstep = 1.0

### 2023-08-16 scripts/single-sim.jl

@@ -25,7 +25,7 @@ singlesim_time_p = SimTimeParameters(0.0, 365.0 * 100, 1.0)
 latent_per = 8
 dur_inf = 5
 R₀ = 10.0
-σ = 1 / latent_per
+sigma = 1 / latent_per
 γ = 1 / dur_inf
 μ = 1 / (62.5 * 365)
 # beta_mean is the average transmission rate
@@ -33,7 +33,7 @@ beta_mean = calculate_beta(R₀, γ, μ, 1, N)
 # Adjust the scale of the seasonal variation in infectivity i.e. beta_force scales the amplitude of cosine function
 beta_force = 0.2
 ε = (1.06 * μ * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
-p = (beta_mean, beta_force, σ, γ, μ, ε, R₀)
+p = (beta_mean, beta_force, sigma, γ, μ, ε, R₀)
 
 Random.seed!(1234)
 

### 2023-08-16 scripts/single-sim_bifurcation.jl

@@ -29,7 +29,7 @@ prog = Progress(n_μs)
     change = @view bifurc_μ_change_arr[:, :, k]
     jump = @view bifurc_μ_jump_arr[:, :, k]
 
-    params = (beta_mean, beta_force, σ, γ, μ_run, ε, R₀)
+    params = (beta_mean, beta_force, sigma, γ, μ_run, ε, R₀)
     seir_mod!(seir, change, jump,
         u₀, params, trange; dt = τ, type = "det",
     )
@@ -88,7 +88,7 @@ bifurc_beta_force_jump_arr = zeros(Float64, 9, tlength, n_beta_forces);
     change = @view bifurc_beta_force_change_arr[:, :, k]
     jump = @view bifurc_beta_force_jump_arr[:, :, k]
 
-    local p = (beta_mean, beta_force, σ, γ, μ, ε, R₀)
+    local p = (beta_mean, beta_force, sigma, γ, μ, ε, R₀)
 
     seir_mod!(seir, change, jump, u₀, p, trange; dt = τ, type = "det")
 end
@@ -143,7 +143,7 @@ prog = Progress(length(μ_vec) * length(beta_force_vec))
     change = @view bifurc_μ_beta_force_change_arr[:, :, k, l]
     jump = @view bifurc_μ_beta_force_jump_arr[:, :, k, l]
 
-    params = (beta_mean, beta_force, σ, γ, μ, ε, R₀)
+    params = (beta_mean, beta_force, sigma, γ, μ, ε, R₀)
 
     seir_mod!(seir, change, jump, u₀, params, trange; dt = τ, type = "det")
     next!(prog)

### 2023-08-16 src/Julia/SEIR-model.jl

@@ -129,12 +129,12 @@ function seir_mod_loop!(
     N = state_arr[5, j - 1]
 
     # Unpack the parameters for easier use
-    beta_mean, beta_force, σ, γ, μ, ε, R₀ = p
+    beta_mean, beta_force, sigma, γ, μ, ε, R₀ = p
     beta_t = calculate_beta_amp(beta_mean, beta_force, t)
 
     # Calculate the rates of each event
     infec_rate = beta_t * S * I        # 1
-    latent_rate = σ * E             # 2
+    latent_rate = sigma * E             # 2
     recov_rate = γ * I              # 3
     birth_rate = μ * N              # 4
     S_death_rate = μ * S            # 5

### 2023-08-16 src/Julia/ensemble-functions.jl

@@ -50,7 +50,7 @@ function run_jump_prob(param_dict)
     @unpack N, u₀_prop, transmission_p, time_p, nsims, dt, beta_force,
     births_per_k, seed = param_dict
     @unpack s, e, i, r = u₀_prop
-    @unpack R₀, σ, γ = transmission_p
+    @unpack R₀, sigma, γ = transmission_p
     @unpack tstep, tlength, trange = time_p
 
     u₀ = convert.(Int64, [s * N, e * N, i * N, r * N, N])
@@ -60,7 +60,7 @@ function run_jump_prob(param_dict)
     beta_mean = calculate_beta(R₀, γ, μ, 1, N)
     ε = (1.06 * μ * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
 
-    p = (beta_mean, beta_force, σ, γ, μ, ε, R₀)
+    p = (beta_mean, beta_force, sigma, γ, μ, ε, R₀)
 
     ensemble_seir_arr = zeros(Int64, size(u₀, 1), tlength, nsims)
     ensemble_change_arr = zeros(Int64, size(u₀, 1), tlength, nsims)

### 2023-08-16 src/Julia/structs.jl

@@ -1,6 +1,6 @@
 struct EnsembleTransmissionParameters
     R₀::Float64
-    σ::Float64
+    sigma::Float64
     γ::Float64
 end
 

### 2023-08-16 notebooks/Julia/simulation-detection.qmd
Rename gamma vars for clarity

@@
 sigma = 365/8
-γ = 365/5
+gamma = 365/5
 ε = 100
-p = (beta, sigma, γ, μ, ε)
+p = (beta, sigma, gamma, μ, ε)
 ```
@@
 
-recov_rate(u, p, t) = p[3] * u[3]   # γ * I
+recov_rate(u, p, t) = p[3] * u[3]   # gamma * I
 function recov_affect!(integrator)

### 2023-08-16 scripts/ensemble-sim.jl

@@ -28,8 +28,8 @@ latent_per_days = 8
 dur_inf_days = 5
 R₀ = 10.0
 sigma = 1 / latent_per_days
-γ = 1 / dur_inf_days
-transmission_p = EnsembleTransmissionParameters(R₀, sigma, γ)
+gamma = 1 / dur_inf_days
+transmission_p = EnsembleTransmissionParameters(R₀, sigma, gamma)
 tmin = 0.0
 tmax = 365.0 * 100
 tstep = 1.0

### 2023-08-16 scripts/single-sim.jl

@@ -26,14 +26,14 @@ latent_per = 8
 dur_inf = 5
 R₀ = 10.0
 sigma = 1 / latent_per
-γ = 1 / dur_inf
+gamma = 1 / dur_inf
 μ = 1 / (62.5 * 365)
 # beta_mean is the average transmission rate
-beta_mean = calculate_beta(R₀, γ, μ, 1, N)
+beta_mean = calculate_beta(R₀, gamma, μ, 1, N)
 # Adjust the scale of the seasonal variation in infectivity i.e. beta_force scales the amplitude of cosine function
 beta_force = 0.2
 ε = (1.06 * μ * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
-p = (beta_mean, beta_force, sigma, γ, μ, ε, R₀)
+p = (beta_mean, beta_force, sigma, gamma, μ, ε, R₀)
 
 Random.seed!(1234)
 

### 2023-08-16 scripts/single-sim_bifurcation.jl

@@ -29,7 +29,7 @@ prog = Progress(n_μs)
     change = @view bifurc_μ_change_arr[:, :, k]
     jump = @view bifurc_μ_jump_arr[:, :, k]
 
-    params = (beta_mean, beta_force, sigma, γ, μ_run, ε, R₀)
+    params = (beta_mean, beta_force, sigma, gamma, μ_run, ε, R₀)
     seir_mod!(seir, change, jump,
         u₀, params, trange; dt = τ, type = "det",
     )
@@ -88,7 +88,7 @@ bifurc_beta_force_jump_arr = zeros(Float64, 9, tlength, n_beta_forces);
     change = @view bifurc_beta_force_change_arr[:, :, k]
     jump = @view bifurc_beta_force_jump_arr[:, :, k]
 
-    local p = (beta_mean, beta_force, sigma, γ, μ, ε, R₀)
+    local p = (beta_mean, beta_force, sigma, gamma, μ, ε, R₀)
 
     seir_mod!(seir, change, jump, u₀, p, trange; dt = τ, type = "det")
 end
@@ -143,7 +143,7 @@ prog = Progress(length(μ_vec) * length(beta_force_vec))
     change = @view bifurc_μ_beta_force_change_arr[:, :, k, l]
     jump = @view bifurc_μ_beta_force_jump_arr[:, :, k, l]
 
-    params = (beta_mean, beta_force, sigma, γ, μ, ε, R₀)
+    params = (beta_mean, beta_force, sigma, gamma, μ, ε, R₀)
 
     seir_mod!(seir, change, jump, u₀, params, trange; dt = τ, type = "det")
     next!(prog)

### 2023-08-16 src/Julia/SEIR-model.jl

@@ -129,13 +129,13 @@ function seir_mod_loop!(
     N = state_arr[5, j - 1]
 
     # Unpack the parameters for easier use
-    beta_mean, beta_force, sigma, γ, μ, ε, R₀ = p
+    beta_mean, beta_force, sigma, gamma, μ, ε, R₀ = p
     beta_t = calculate_beta_amp(beta_mean, beta_force, t)
 
     # Calculate the rates of each event
     infec_rate = beta_t * S * I        # 1
     latent_rate = sigma * E             # 2
-    recov_rate = γ * I              # 3
+    recov_rate = gamma * I              # 3
     birth_rate = μ * N              # 4
     S_death_rate = μ * S            # 5
     E_death_rate = μ * E            # 6

### 2023-08-16 src/Julia/ensemble-functions.jl

@@ -50,17 +50,17 @@ function run_jump_prob(param_dict)
     @unpack N, u₀_prop, transmission_p, time_p, nsims, dt, beta_force,
     births_per_k, seed = param_dict
     @unpack s, e, i, r = u₀_prop
-    @unpack R₀, sigma, γ = transmission_p
+    @unpack R₀, sigma, gamma = transmission_p
     @unpack tstep, tlength, trange = time_p
 
     u₀ = convert.(Int64, [s * N, e * N, i * N, r * N, N])
     u0_dict = Dict(zip([:S, :E, :I, :R, :N], u₀))
 
     μ = births_per_k / (1000 * 365)
-    beta_mean = calculate_beta(R₀, γ, μ, 1, N)
+    beta_mean = calculate_beta(R₀, gamma, μ, 1, N)
     ε = (1.06 * μ * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
 
-    p = (beta_mean, beta_force, sigma, γ, μ, ε, R₀)
+    p = (beta_mean, beta_force, sigma, gamma, μ, ε, R₀)
 
     ensemble_seir_arr = zeros(Int64, size(u₀, 1), tlength, nsims)
     ensemble_change_arr = zeros(Int64, size(u₀, 1), tlength, nsims)

### 2023-08-16 src/Julia/structs.jl

@@ -1,7 +1,7 @@
 struct EnsembleTransmissionParameters
     R₀::Float64
     sigma::Float64
-    γ::Float64
+    gamma::Float64
 end
 
 struct SimTimeParameters

### 2023-08-16 src/Julia/transmission-functions.jl

@@ -7,7 +7,7 @@ using FLoops
 
 
 """
-    calculate_beta(R₀, γ, μ, C, pop_matrix)
+    calculate_beta(R₀, gamma, μ, C, pop_matrix)
 
 Calculate the value beta for a given set of parameters and contact matrix.
 
@@ -19,7 +19,7 @@ julia> calculate_beta(2.0, 1 / 8, 0.0, ones(1, 1), [1000])
 """
 # TODO: Currently only works when the populations are the same size as each other, and doesn't account for an exposed state.
 function calculate_beta(
-    R₀::T, γ::T, μ::T, C::Array{T}, pop_matrix::Array{T}
+    R₀::T, gamma::T, μ::T, C::Array{T}, pop_matrix::Array{T}
 ) where {T<:AbstractFloat}
     size(C, 1) == size(C, 2) ? nothing : error("C must be square")
     if size(C, 1) == size(pop_matrix, 1)
@@ -29,7 +29,7 @@ function calculate_beta(
     end
 
     F = C .* pop_matrix
-    V = Diagonal(repeat([γ + μ], size(C, 1)))
+    V = Diagonal(repeat([gamma + μ], size(C, 1)))
 
     FV⁻¹ = F * inv(V)
     eigenvals, eigenvectors = eigen(FV⁻¹)
@@ -38,10 +38,10 @@ function calculate_beta(
     return beta
 end
 
-function calculate_beta(R₀, γ, μ, C, pop_matrix)
+function calculate_beta(R₀, gamma, μ, C, pop_matrix)
     return calculate_beta(
         convert(Float64, R₀),
-        convert(Float64, γ),
+        convert(Float64, gamma),
         convert(Float64, μ),
         convert(Array{Float64}, [C]),
         convert(Array{Float64}, [pop_matrix]),
@@ -63,7 +63,7 @@ function calculate_beta(
         (nac + 1):(nac + nic * nac)]
 
     F = C .* pop_matrix
-    # F = substitute(Jac, Dict(γ => 0.0, μ => 0.0))
+    # F = substitute(Jac, Dict(gamma => 0.0, μ => 0.0))
     V = substitute(Jac, Dict(beta => 0.0))
     FV⁻¹ = F * -inv(V)
     eigenvals =
@@ -95,7 +95,7 @@ function calculate_beta(
 end
 
 """
-    calculateR0(beta, γ, μ, C, pop_matrix)
[…]

### 2023-08-16 notebooks/Julia/simulation-detection.qmd
Rename mu vars for clarity

@@
 
-μ = 1
+mu = 1
 beta = 1000
@@
 ε = 100
-p = (beta, sigma, gamma, μ, ε)
+p = (beta, sigma, gamma, mu, ε)
 ```
@@
 
-birth_rate(u, p, t) = p[4] * (u[1] + u[2] + u[3] + u[4]) # μ * N
+birth_rate(u, p, t) = p[4] * (u[1] + u[2] + u[3] + u[4]) # mu * N
 function birth_affect!(integrator)
@@
 
-S_death_rate(u, p, t) = p[4] * u[1] # μ * S
+S_death_rate(u, p, t) = p[4] * u[1] # mu * S
 function S_death_affect!(integrator)
@@
 
-E_death_rate(u, p, t) = p[4] * u[2] # μ * E
+E_death_rate(u, p, t) = p[4] * u[2] # mu * E
 function E_death_affect!(integrator)
@@
 
-I_death_rate(u, p, t) = p[4] * u[3] # μ * I
+I_death_rate(u, p, t) = p[4] * u[3] # mu * I
 function I_death_affect!(integrator)
@@
 
-R_death_rate(u, p, t) = p[4] * u[4] # μ * R
+R_death_rate(u, p, t) = p[4] * u[4] # mu * R
 function R_death_affect!(integrator)

### 2023-08-16 scripts/ensemble-sim.jl

@@ -16,12 +16,12 @@ u₀_prop_map = [
 dt_vec = [1.0]
 tmax_vec = [365.0 * 100]
 beta_force_vec = collect(0.0:0.1:0.4)
-μ_min = 5
-μ_max = 20
-μ_step = 5.0
-n_μs = length(μ_min:μ_step:μ_max)
-μ_vec = zeros(Float64, n_μs)
-μ_vec = convert.(Int64, collect(μ_min:μ_step:μ_max))
+mu_min = 5
+mu_max = 20
+mu_step = 5.0
+n_mus = length(mu_min:mu_step:mu_max)
+mu_vec = zeros(Float64, n_mus)
+mu_vec = convert.(Int64, collect(mu_min:mu_step:mu_max))
 seed = 1234
 
 latent_per_days = 8
@@ -44,7 +44,7 @@ base_param_dict = @dict(
     dt = dt_vec,
     tmax = tmax_vec,
     beta_force = beta_force_vec,
-    births_per_k = μ_vec,
+    births_per_k = mu_vec,
     seed = seed,
 )
 

### 2023-08-16 scripts/single-sim.jl

@@ -27,13 +27,13 @@ dur_inf = 5
 R₀ = 10.0
 sigma = 1 / latent_per
 gamma = 1 / dur_inf
-μ = 1 / (62.5 * 365)
+mu = 1 / (62.5 * 365)
 # beta_mean is the average transmission rate
-beta_mean = calculate_beta(R₀, gamma, μ, 1, N)
+beta_mean = calculate_beta(R₀, gamma, mu, 1, N)
 # Adjust the scale of the seasonal variation in infectivity i.e. beta_force scales the amplitude of cosine function
 beta_force = 0.2
-ε = (1.06 * μ * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
-p = (beta_mean, beta_force, sigma, gamma, μ, ε, R₀)
+ε = (1.06 * mu * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
+p = (beta_mean, beta_force, sigma, gamma, mu, ε, R₀)
 
 Random.seed!(1234)
 

### 2023-08-16 scripts/single-sim_bifurcation.jl

@@ -8,28 +8,28 @@ using IterTools
 includet(scriptsdir("single-sim.jl"))
 
 #%%
-μ_min = 10
-μ_max = 100
-μ_step = 2.0
-n_μs = length(μ_min:μ_step:μ_max)
-μ_vec = zeros(Float64, n_μs)
-μ_vec .= collect(μ_min:μ_step:μ_max) ./ (1000 * 365)
+mu_min = 10
+mu_max = 100
+mu_step = 2.0
+n_mus = length(mu_min:mu_step:mu_max)
+mu_vec = zeros(Float64, n_mus)
+mu_vec .= collect(mu_min:mu_step:mu_max) ./ (1000 * 365)
 
-ε_vec = (1.06 .* μ_vec .* (R₀ - 1)) ./ sqrt(N)
+ε_vec = (1.06 .* mu_vec .* (R₀ - 1)) ./ sqrt(N)
 
-bifurc_μ_seir_arr = zeros(Float64, size(u₀, 1), tlength, n_μs);
-bifurc_μ_change_arr = zeros(Float64, size(u₀, 1), tlength, n_μs);
-bifurc_μ_jump_arr = zeros(Float64, 9, tlength, n_μs);
+bifurc_mu_seir_arr = zeros(Float64, size(u₀, 1), tlength, n_mus);
+bifurc_mu_change_arr = zeros(Float64, size(u₀, 1), tlength, n_mus);
+bifurc_mu_jump_arr = zeros(Float64, 9, tlength, n_mus);
 
-prog = Progress(n_μs)
-@floop for (k, μ_run) in pairs(μ_vec)
+prog = Progress(n_mus)
+@floop for (k, mu_run) in pairs(mu_vec)
     ε = ε_vec[k]
 
-    seir = @view bifurc_μ_seir_arr[:, :, k]
-    change = @view bifurc_μ_change_arr[:, :, k]
-    jump = @view bifurc_μ_jump_arr[:, :, k]
+    seir = @view bifurc_mu_seir_arr[:, :, k]
+    change = @view bifurc_mu_change_arr[:, :, k]
+    jump = @view bifurc_mu_jump_arr[:, :, k]
 
-    params = (beta_mean, beta_force, sigma, gamma, μ_run, ε, R₀)
+    params = (beta_mean, beta_force, sigma, gamma, mu_run, ε, R₀)
     seir_mod!(seir, change, jump,
[…]

### 2023-08-16 src/Julia/SEIR-model.jl

@@ -129,18 +129,18 @@ function seir_mod_loop!(
     N = state_arr[5, j - 1]
 
     # Unpack the parameters for easier use
-    beta_mean, beta_force, sigma, gamma, μ, ε, R₀ = p
+    beta_mean, beta_force, sigma, gamma, mu, ε, R₀ = p
     beta_t = calculate_beta_amp(beta_mean, beta_force, t)
 
     # Calculate the rates of each event
     infec_rate = beta_t * S * I        # 1
     latent_rate = sigma * E             # 2
     recov_rate = gamma * I              # 3
-    birth_rate = μ * N              # 4
-    S_death_rate = μ * S            # 5
-    E_death_rate = μ * E            # 6
-    I_death_rate = μ * I            # 7
-    R_death_rate = μ * R            # 8
+    birth_rate = mu * N              # 4
+    S_death_rate = mu * S            # 5
+    E_death_rate = mu * E            # 6
+    I_death_rate = mu * I            # 7
+    R_death_rate = mu * R            # 8
     import_rate = ε * N / R₀        # 9
 
     rates = [

### 2023-08-16 src/Julia/ensemble-functions.jl

@@ -56,11 +56,11 @@ function run_jump_prob(param_dict)
     u₀ = convert.(Int64, [s * N, e * N, i * N, r * N, N])
     u0_dict = Dict(zip([:S, :E, :I, :R, :N], u₀))
 
-    μ = births_per_k / (1000 * 365)
-    beta_mean = calculate_beta(R₀, gamma, μ, 1, N)
-    ε = (1.06 * μ * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
+    mu = births_per_k / (1000 * 365)
+    beta_mean = calculate_beta(R₀, gamma, mu, 1, N)
+    ε = (1.06 * mu * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
 
-    p = (beta_mean, beta_force, sigma, gamma, μ, ε, R₀)
+    p = (beta_mean, beta_force, sigma, gamma, mu, ε, R₀)
 
     ensemble_seir_arr = zeros(Int64, size(u₀, 1), tlength, nsims)
     ensemble_change_arr = zeros(Int64, size(u₀, 1), tlength, nsims)

### 2023-08-16 src/Julia/transmission-functions.jl

@@ -7,7 +7,7 @@ using FLoops
 
 
 """
-    calculate_beta(R₀, gamma, μ, C, pop_matrix)
+    calculate_beta(R₀, gamma, mu, C, pop_matrix)
 
 Calculate the value beta for a given set of parameters and contact matrix.
 
@@ -19,7 +19,7 @@ julia> calculate_beta(2.0, 1 / 8, 0.0, ones(1, 1), [1000])
 """
 # TODO: Currently only works when the populations are the same size as each other, and doesn't account for an exposed state.
 function calculate_beta(
-    R₀::T, gamma::T, μ::T, C::Array{T}, pop_matrix::Array{T}
+    R₀::T, gamma::T, mu::T, C::Array{T}, pop_matrix::Array{T}
 ) where {T<:AbstractFloat}
     size(C, 1) == size(C, 2) ? nothing : error("C must be square")
     if size(C, 1) == size(pop_matrix, 1)
@@ -29,7 +29,7 @@ function calculate_beta(
     end
 
     F = C .* pop_matrix
-    V = Diagonal(repeat([gamma + μ], size(C, 1)))
+    V = Diagonal(repeat([gamma + mu], size(C, 1)))
 
     FV⁻¹ = F * inv(V)
     eigenvals, eigenvectors = eigen(FV⁻¹)
@@ -38,11 +38,11 @@ function calculate_beta(
     return beta
 end
 
-function calculate_beta(R₀, gamma, μ, C, pop_matrix)
+function calculate_beta(R₀, gamma, mu, C, pop_matrix)
     return calculate_beta(
         convert(Float64, R₀),
         convert(Float64, gamma),
-        convert(Float64, μ),
+        convert(Float64, mu),
         convert(Array{Float64}, [C]),
         convert(Array{Float64}, [pop_matrix]),
     )
@@ -63,7 +63,7 @@ function calculate_beta(
         (nac + 1):(nac + nic * nac)]
 
     F = C .* pop_matrix
-    # F = substitute(Jac, Dict(gamma => 0.0, μ => 0.0))
+    # F = substitute(Jac, Dict(gamma => 0.0, mu => 0.0))
     V = substitute(Jac, Dict(beta => 0.0))
     FV⁻¹ = F * -inv(V)
     eigenvals =
@@ -95,7 +95,7 @@ function calculate_beta(
 end
 
 """
-    calculateR0(beta, gamma, μ, C, pop_matrix)
[…]

### 2023-08-16 scripts/single-sim.jl
Rename epsilon vars for clarity

@@ -32,8 +32,8 @@ mu = 1 / (62.5 * 365)
 beta_mean = calculate_beta(R₀, gamma, mu, 1, N)
 # Adjust the scale of the seasonal variation in infectivity i.e. beta_force scales the amplitude of cosine function
 beta_force = 0.2
-ε = (1.06 * mu * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
-p = (beta_mean, beta_force, sigma, gamma, mu, ε, R₀)
+epsilon = (1.06 * mu * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
+p = (beta_mean, beta_force, sigma, gamma, mu, epsilon, R₀)
 
 Random.seed!(1234)
 

### 2023-08-16 scripts/single-sim_bifurcation.jl

@@ -15,7 +15,7 @@ n_mus = length(mu_min:mu_step:mu_max)
 mu_vec = zeros(Float64, n_mus)
 mu_vec .= collect(mu_min:mu_step:mu_max) ./ (1000 * 365)
 
-ε_vec = (1.06 .* mu_vec .* (R₀ - 1)) ./ sqrt(N)
+epsilon_vec = (1.06 .* mu_vec .* (R₀ - 1)) ./ sqrt(N)
 
 bifurc_mu_seir_arr = zeros(Float64, size(u₀, 1), tlength, n_mus);
 bifurc_mu_change_arr = zeros(Float64, size(u₀, 1), tlength, n_mus);
@@ -23,13 +23,13 @@ bifurc_mu_jump_arr = zeros(Float64, 9, tlength, n_mus);
 
 prog = Progress(n_mus)
 @floop for (k, mu_run) in pairs(mu_vec)
-    ε = ε_vec[k]
+    epsilon = epsilon_vec[k]
 
     seir = @view bifurc_mu_seir_arr[:, :, k]
     change = @view bifurc_mu_change_arr[:, :, k]
     jump = @view bifurc_mu_jump_arr[:, :, k]
 
-    params = (beta_mean, beta_force, sigma, gamma, mu_run, ε, R₀)
+    params = (beta_mean, beta_force, sigma, gamma, mu_run, epsilon, R₀)
     seir_mod!(seir, change, jump,
         u₀, params, trange; dt = τ, type = "det",
     )
@@ -81,14 +81,14 @@ bifurc_beta_force_change_arr = zeros(Float64, size(u₀, 1), tlength, n_beta_for
 bifurc_beta_force_jump_arr = zeros(Float64, 9, tlength, n_beta_forces);
 
 mu = 50 / (1000 * 365)
-ε = 1.06 * mu * (R₀ - 1) / sqrt(N)
+epsilon = 1.06 * mu * (R₀ - 1) / sqrt(N)
 
 @showprogress for (k, beta_force) in pairs(beta_force_vec)
     seir = @view bifurc_beta_force_seir_arr[:, :, k]
     change = @view bifurc_beta_force_change_arr[:, :, k]
     jump = @view bifurc_beta_force_jump_arr[:, :, k]
 
-    local p = (beta_mean, beta_force, sigma, gamma, mu, ε, R₀)
+    local p = (beta_mean, beta_force, sigma, gamma, mu, epsilon, R₀)
 
[…]

### 2023-08-16 src/Julia/SEIR-model.jl

@@ -129,7 +129,7 @@ function seir_mod_loop!(
     N = state_arr[5, j - 1]
 
     # Unpack the parameters for easier use
-    beta_mean, beta_force, sigma, gamma, mu, ε, R₀ = p
+    beta_mean, beta_force, sigma, gamma, mu, epsilon, R₀ = p
     beta_t = calculate_beta_amp(beta_mean, beta_force, t)
 
     # Calculate the rates of each event
@@ -141,7 +141,7 @@ function seir_mod_loop!(
     E_death_rate = mu * E            # 6
     I_death_rate = mu * I            # 7
     R_death_rate = mu * R            # 8
-    import_rate = ε * N / R₀        # 9
+    import_rate = epsilon * N / R₀        # 9
 
     rates = [
         infec_rate, latent_rate, recov_rate, birth_rate, S_death_rate,

### 2023-08-16 src/Julia/ensemble-functions.jl

@@ -58,9 +58,9 @@ function run_jump_prob(param_dict)
 
     mu = births_per_k / (1000 * 365)
     beta_mean = calculate_beta(R₀, gamma, mu, 1, N)
-    ε = (1.06 * mu * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
+    epsilon = (1.06 * mu * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
 
-    p = (beta_mean, beta_force, sigma, gamma, mu, ε, R₀)
+    p = (beta_mean, beta_force, sigma, gamma, mu, epsilon, R₀)
 
     ensemble_seir_arr = zeros(Int64, size(u₀, 1), tlength, nsims)
     ensemble_change_arr = zeros(Int64, size(u₀, 1), tlength, nsims)

### 2023-08-16 scripts/single-sim.jl
Rename R_0 vars for clarity

@@ -24,16 +24,16 @@ singlesim_time_p = SimTimeParameters(0.0, 365.0 * 100, 1.0)
 
 latent_per = 8
 dur_inf = 5
-R₀ = 10.0
+R_0 = 10.0
 sigma = 1 / latent_per
 gamma = 1 / dur_inf
 mu = 1 / (62.5 * 365)
 # beta_mean is the average transmission rate
-beta_mean = calculate_beta(R₀, gamma, mu, 1, N)
+beta_mean = calculate_beta(R_0, gamma, mu, 1, N)
 # Adjust the scale of the seasonal variation in infectivity i.e. beta_force scales the amplitude of cosine function
 beta_force = 0.2
-epsilon = (1.06 * mu * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
-p = (beta_mean, beta_force, sigma, gamma, mu, epsilon, R₀)
+epsilon = (1.06 * mu * (R_0 - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
+p = (beta_mean, beta_force, sigma, gamma, mu, epsilon, R_0)
 
 Random.seed!(1234)
 

### 2023-08-16 scripts/single-sim_bifurcation.jl

@@ -15,7 +15,7 @@ n_mus = length(mu_min:mu_step:mu_max)
 mu_vec = zeros(Float64, n_mus)
 mu_vec .= collect(mu_min:mu_step:mu_max) ./ (1000 * 365)
 
-epsilon_vec = (1.06 .* mu_vec .* (R₀ - 1)) ./ sqrt(N)
+epsilon_vec = (1.06 .* mu_vec .* (R_0 - 1)) ./ sqrt(N)
 
 bifurc_mu_seir_arr = zeros(Float64, size(u₀, 1), tlength, n_mus);
 bifurc_mu_change_arr = zeros(Float64, size(u₀, 1), tlength, n_mus);
@@ -29,7 +29,7 @@ prog = Progress(n_mus)
     change = @view bifurc_mu_change_arr[:, :, k]
     jump = @view bifurc_mu_jump_arr[:, :, k]
 
-    params = (beta_mean, beta_force, sigma, gamma, mu_run, epsilon, R₀)
+    params = (beta_mean, beta_force, sigma, gamma, mu_run, epsilon, R_0)
     seir_mod!(seir, change, jump,
         u₀, params, trange; dt = τ, type = "det",
     )
@@ -81,14 +81,14 @@ bifurc_beta_force_change_arr = zeros(Float64, size(u₀, 1), tlength, n_beta_for
 bifurc_beta_force_jump_arr = zeros(Float64, 9, tlength, n_beta_forces);
 
 mu = 50 / (1000 * 365)
-epsilon = 1.06 * mu * (R₀ - 1) / sqrt(N)
+epsilon = 1.06 * mu * (R_0 - 1) / sqrt(N)
 
 @showprogress for (k, beta_force) in pairs(beta_force_vec)
     seir = @view bifurc_beta_force_seir_arr[:, :, k]
     change = @view bifurc_beta_force_change_arr[:, :, k]
     jump = @view bifurc_beta_force_jump_arr[:, :, k]
 
-    local p = (beta_mean, beta_force, sigma, gamma, mu, epsilon, R₀)
+    local p = (beta_mean, beta_force, sigma, gamma, mu, epsilon, R_0)
 
     seir_mod!(seir, change, jump, u₀, p, trange; dt = τ, type = "det")
 end
@@ -143,7 +143,7 @@ prog = Progress(length(mu_vec) * length(beta_force_vec))
     change = @view bifurc_mu_beta_force_change_arr[:, :, k, l]
[…]

### 2023-08-16 src/Julia/SEIR-model.jl

@@ -129,7 +129,7 @@ function seir_mod_loop!(
     N = state_arr[5, j - 1]
 
     # Unpack the parameters for easier use
-    beta_mean, beta_force, sigma, gamma, mu, epsilon, R₀ = p
+    beta_mean, beta_force, sigma, gamma, mu, epsilon, R_0 = p
     beta_t = calculate_beta_amp(beta_mean, beta_force, t)
 
     # Calculate the rates of each event
@@ -141,7 +141,7 @@ function seir_mod_loop!(
     E_death_rate = mu * E            # 6
     I_death_rate = mu * I            # 7
     R_death_rate = mu * R            # 8
-    import_rate = epsilon * N / R₀        # 9
+    import_rate = epsilon * N / R_0        # 9
 
     rates = [
         infec_rate, latent_rate, recov_rate, birth_rate, S_death_rate,

### 2023-08-16 src/Julia/ensemble-functions.jl

@@ -50,17 +50,17 @@ function run_jump_prob(param_dict)
     @unpack N, u₀_prop, transmission_p, time_p, nsims, dt, beta_force,
     births_per_k, seed = param_dict
     @unpack s, e, i, r = u₀_prop
-    @unpack R₀, sigma, gamma = transmission_p
+    @unpack R_0, sigma, gamma = transmission_p
     @unpack tstep, tlength, trange = time_p
 
     u₀ = convert.(Int64, [s * N, e * N, i * N, r * N, N])
     u0_dict = Dict(zip([:S, :E, :I, :R, :N], u₀))
 
     mu = births_per_k / (1000 * 365)
-    beta_mean = calculate_beta(R₀, gamma, mu, 1, N)
-    epsilon = (1.06 * mu * (R₀ - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
+    beta_mean = calculate_beta(R_0, gamma, mu, 1, N)
+    epsilon = (1.06 * mu * (R_0 - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
 
-    p = (beta_mean, beta_force, sigma, gamma, mu, epsilon, R₀)
+    p = (beta_mean, beta_force, sigma, gamma, mu, epsilon, R_0)
 
     ensemble_seir_arr = zeros(Int64, size(u₀, 1), tlength, nsims)
     ensemble_change_arr = zeros(Int64, size(u₀, 1), tlength, nsims)

### 2023-08-16 src/Julia/structs.jl

@@ -1,5 +1,5 @@
 struct EnsembleTransmissionParameters
-    R₀::Float64
+    R_0::Float64
     sigma::Float64
     gamma::Float64
 end

### 2023-08-16 src/Julia/transmission-functions.jl

@@ -7,7 +7,7 @@ using FLoops
 
 
 """
-    calculate_beta(R₀, gamma, mu, C, pop_matrix)
+    calculate_beta(R_0, gamma, mu, C, pop_matrix)
 
 Calculate the value beta for a given set of parameters and contact matrix.
 
@@ -19,7 +19,7 @@ julia> calculate_beta(2.0, 1 / 8, 0.0, ones(1, 1), [1000])
 """
 # TODO: Currently only works when the populations are the same size as each other, and doesn't account for an exposed state.
 function calculate_beta(
-    R₀::T, gamma::T, mu::T, C::Array{T}, pop_matrix::Array{T}
+    R_0::T, gamma::T, mu::T, C::Array{T}, pop_matrix::Array{T}
 ) where {T<:AbstractFloat}
     size(C, 1) == size(C, 2) ? nothing : error("C must be square")
     if size(C, 1) == size(pop_matrix, 1)
@@ -33,14 +33,14 @@ function calculate_beta(
 
     FV⁻¹ = F * inv(V)
     eigenvals, eigenvectors = eigen(FV⁻¹)
-    beta = R₀ / maximum(real(eigenvals))
+    beta = R_0 / maximum(real(eigenvals))
 
     return beta
 end
 
-function calculate_beta(R₀, gamma, mu, C, pop_matrix)
+function calculate_beta(R_0, gamma, mu, C, pop_matrix)
     return calculate_beta(
-        convert(Float64, R₀),
+        convert(Float64, R_0),
         convert(Float64, gamma),
         convert(Float64, mu),
         convert(Array{Float64}, [C]),
@@ -49,7 +49,7 @@ function calculate_beta(R₀, gamma, mu, C, pop_matrix)
 end
 
 function calculate_beta(
-    ode::S, nic::T, nac::T, R₀::U, param::Dict{Num,U}, C::Array{U},
+    ode::S, nic::T, nac::T, R_0::U, param::Dict{Num,U}, C::Array{U},
     pop_matrix::Array{U},
 ) where {S<:ODESystem,T<:Int,U<:AbstractFloat}
     size(C, 1) == size(C, 2) ? nothing : error("C must be square")
@@ -71,33 +71,33 @@ function calculate_beta(
             Float64,
             Symbolics.value.(eigvals(eigen(substitute(FV⁻¹, Dict(param...))))),
         )
[…]

### 2023-08-16 scripts/SEIR_Tau-leaping-seasonal-imports.jl
Rewrite numbers to use _ to delimit every thousand

@@ -35,7 +35,7 @@ includet(funsdir("SEIR-model.jl"))
 create_sir_quantiles_plot(
     ensemble_seir_summary; labels = state_labels, colors = seircolors,
     annual = true, caption = caption, δt = param_dict[:dt], xlims = (80, 100),
-    ylims = (0, 1000),
+    ylims = (0, 1_000),
 )
 
 #%%
@@ -144,7 +144,7 @@ map(
     ax -> xlims!(ax, (92, 94)),
     [above5ax_prev, above5ax_inc, above5ax_periodsum],
 )
-ylims!(above5ax_periodsum, (0, 10000))
+ylims!(above5ax_periodsum, (0, 10_000))
 ylims!(above5ax_inc, (0, 300))
 
 axislegend(

### 2023-08-16 scripts/ensemble-sim.jl

@@ -9,7 +9,7 @@ includet(funsdir("ensemble-functions.jl"))
 
 #%%
 N_vec = convert.(Int64, [5e5])
-nsims_vec = [1000]
+nsims_vec = [1_000]
 u₀_prop_map = [
     Dict(:s => 0.1, :e => 0.01, :i => 0.01, :r => 0.88)
 ]
@@ -72,9 +72,9 @@ summarize_ensemble_jump_prob(summ_param_dict; prog = prog)
 #%%
 ensemble_spec = EnsembleSpecification(
     ("seasonal-infectivity-import", "tau-leaping"),
-    500000,
+    500_000,
     0.88,
-    1000,
+    1_000,
     20,
     0.2,
     ensemble_time_p,

### 2023-08-16 scripts/ensemble-sim_plots.jl

@@ -11,6 +11,6 @@ includet(scriptsdir("ensemble-sim.jl"))
 create_sir_quantiles_plot(
     ensemble_seir_summary; labels = seir_state_labels, colors = seircolors,
     annual = true, caption = caption, δt = param_dict[:dt], xlims = (80, 100),
-    ylims = (0, 1000),
+    ylims = (0, 1_000),
 )
 

### 2023-08-16 scripts/single-sim.jl

@@ -14,7 +14,7 @@ includet(funsdir("SEIR-model.jl"))
 includet(funsdir("structs.jl"))
 
 #%%
-N = 5e5
+N = 500_000
 s = 0.1
 e = 0.01
 i = 0.01

### 2023-08-16 scripts/single-sim_bifurcation.jl

@@ -13,7 +13,7 @@ mu_max = 100
 mu_step = 2.0
 n_mus = length(mu_min:mu_step:mu_max)
 mu_vec = zeros(Float64, n_mus)
-mu_vec .= collect(mu_min:mu_step:mu_max) ./ (1000 * 365)
+mu_vec .= collect(mu_min:mu_step:mu_max) ./ (1_000 * 365)
 
 epsilon_vec = (1.06 .* mu_vec .* (R_0 - 1)) ./ sqrt(N)
 
@@ -53,7 +53,7 @@ bifurc_mu_seir_arr[2, (40 * 365):(40 * 365 + 364), 10]
 #%%
 bifurc_mu_fig = Figure()
 bifurc_mu_ax = Axis(
-    bifurc_mu_fig[1, 1]; xlabel = "mu (per 1000, per annum)", ylabel = "Max. I"
+    bifurc_mu_fig[1, 1]; xlabel = "mu (per 1_000, per annum)", ylabel = "Max. I"
 )
 
 for year in eachindex(years)
@@ -80,7 +80,7 @@ bifurc_beta_force_seir_arr = zeros(Float64, size(u₀, 1), tlength, n_beta_force
 bifurc_beta_force_change_arr = zeros(Float64, size(u₀, 1), tlength, n_beta_forces);
 bifurc_beta_force_jump_arr = zeros(Float64, 9, tlength, n_beta_forces);
 
-mu = 50 / (1000 * 365)
+mu = 50 / (1_000 * 365)
 epsilon = 1.06 * mu * (R_0 - 1) / sqrt(N)
 
 @showprogress for (k, beta_force) in pairs(beta_force_vec)
@@ -180,11 +180,11 @@ end
 #%%
 # Because Julia is column-major, we need to transpose the heatmap as it reads the data one column at a time, and in our original matrix, each column represents a different beta_force value.
 bifurc_mu_beta_force_fig, bifurc_mu_beta_force_ax, bifurc_mu_beta_force_hm = heatmap(
-    mu_vec .* (1000 * 365), beta_force_vec, bifurc_mu_beta_force_cycle_summary[:, :, 2]'
+    mu_vec .* (1_000 * 365), beta_force_vec, bifurc_mu_beta_force_cycle_summary[:, :, 2]'
 )
 Colorbar(bifurc_mu_beta_force_fig[:, end + 1], bifurc_mu_beta_force_hm)
 
[…]

### 2023-08-16 src/Julia/ensemble-functions.jl

@@ -56,7 +56,7 @@ function run_jump_prob(param_dict)
     u₀ = convert.(Int64, [s * N, e * N, i * N, r * N, N])
     u0_dict = Dict(zip([:S, :E, :I, :R, :N], u₀))
 
-    mu = births_per_k / (1000 * 365)
+    mu = births_per_k / (1_000 * 365)
     beta_mean = calculate_beta(R_0, gamma, mu, 1, N)
     epsilon = (1.06 * mu * (R_0 - 1)) / sqrt(N) # Commuter imports - see p210 Keeling & Rohani
 

### 2023-08-16 src/Julia/plotting-functions.jl

@@ -8,7 +8,7 @@ WGLMakie.activate!()
 #= CairoMakie.activate!(type = "pdf") =#
 set_aog_theme!()
 # Set depending on size of screen
-update_theme!(; resolution = (1300, 900))
+update_theme!(; resolution = (1_300, 900))
 
 seircolors = ["dodgerblue4", "green", "firebrick3", "chocolate2", "purple"]
 seir_state_labels = ["S", "E", "I", "R", "N"]

### 2023-08-16 src/Julia/transmission-functions.jl

@@ -12,7 +12,7 @@ using FLoops
 Calculate the value beta for a given set of parameters and contact matrix.
 
 ```jldoctest
-julia> calculate_beta(2.0, 1 / 8, 0.0, ones(1, 1), [1000])
+julia> calculate_beta(2.0, 1 / 8, 0.0, ones(1, 1), [1_000])
 0.00025
 ```
 
@@ -100,7 +100,7 @@ end
 Calculate the basic reproduction number R_0 for a given set of parameters and contact matrix.
 
 ```jldoctest
-julia> calculateR0(0.00025, 1 / 8, 0.0, ones(1, 1), [1000])
+julia> calculateR0(0.00025, 1 / 8, 0.0, ones(1, 1), [1_000])
 2.0
 ```
 

### 2023-08-16 notebooks/Julia/simulation-detection.qmd
Rename epsilon vars for clarity
Rewrite numbers to use _ to delimit every thousand
Rename init_states var for clarity

@@
 
-u₀ = [999.0, 10.0, 0.0, 0.0]
+init_states = [999.0, 10.0, 0.0, 0.0]
 
 mu = 1
-beta = 1000
+beta = 1_000
 sigma = 365/8
 gamma = 365/5
-ε = 100
-p = (beta, sigma, gamma, mu, ε)
+epsilon = 100
+p = (beta, sigma, gamma, mu, epsilon)
 ```
@@
 
-prob = DiscreteProblem(u₀, tspan, p)
+prob = DiscreteProblem(init_states, tspan, p)
 jump_prob = JumpProblem(prob, Coevolve(), gillespie_jumps...; dep_graph)
@@
 ```{julia}
-function run_sir_sim(; u₀, tspan, p)
-    prob = DiscreteProblem(u₀, tspan, p)
+function run_sir_sim(; init_states, tspan, p)
+    prob = DiscreteProblem(init_states, tspan, p)
     jump_prob = JumpProblem(
@@
 @showprogress for i in 1:nsims
-    out_jump = run_sir_sim(u₀ = u₀, tspan = tspan, p = p)
+    out_jump = run_sir_sim(init_states = init_states, tspan = tspan, p = p)
     create_sim_array!(out_jump = out_jump)

### 2023-08-16 scripts/SEIR_Tau-leaping-seasonal-imports.jl

@@ -175,9 +175,9 @@ sde_cb = DiscreteCallback(
 
 #%%
 # Noise should be incidence, not prevalence
-noise_u₀ = [10.0]
+init_noise = [10.0]
 tspan = (tlower, tmax)
-noise_prob = SDEProblem(background_ode!, background_noise!, noise_u₀, tspan, p)
+noise_prob = SDEProblem(background_ode!, background_noise!, init_noise, tspan, p)
 noise_sol = solve(
     noise_prob, SRIW1(); callback = sde_cb, dt = param_dict[:dt],
     adaptive = false,
@@ -196,7 +196,7 @@ noise_arr = zeros(
     noise_prob = SDEProblem(
         background_ode!,
         background_noise!,
-        noise_u₀,
+        init_noise,
         tspan,
         p
     )

### 2023-08-16 scripts/ensemble-sim.jl

@@ -10,7 +10,7 @@ includet(funsdir("ensemble-functions.jl"))
 #%%
 N_vec = convert.(Int64, [5e5])
 nsims_vec = [1_000]
-u₀_prop_map = [
+init_states_prop_map = [
     Dict(:s => 0.1, :e => 0.01, :i => 0.01, :r => 0.88)
 ]
 dt_vec = [1.0]
@@ -37,7 +37,7 @@ ensemble_time_p = EnsembleTimeParameters(tmin, tmax, tstep)
 
 base_param_dict = @dict(
     N = N_vec,
-    u₀_prop = u₀_prop_map,
+    init_states_prop = init_states_prop_map,
     transmission_p = transmission_p,
     time_p = time_p,
     nsims = nsims_vec,

### 2023-08-16 scripts/single-sim.jl

@@ -19,7 +19,7 @@ s = 0.1
 e = 0.01
 i = 0.01
 r = 1.0 - (s + e + i)
-u₀ = convert.(Int64, [N * s, N * e, N * i, N * r, N])
+init_states = convert.(Int64, [N * s, N * e, N * i, N * r, N])
 singlesim_time_p = SimTimeParameters(0.0, 365.0 * 100, 1.0)
 
 latent_per = 8
@@ -39,7 +39,7 @@ Random.seed!(1234)
 
 #%%
 seir_array, change_array, jump_array, beta_arr = seir_mod(
-    u₀, p, trange; retbetaarr = true, type = "stoch"
+    init_states, p, trange; retbetaarr = true, type = "stoch"
 );
 
 seir_df = create_sir_df(

### 2023-08-16 scripts/single-sim_bifurcation.jl

@@ -17,8 +17,8 @@ mu_vec .= collect(mu_min:mu_step:mu_max) ./ (1_000 * 365)
 
 epsilon_vec = (1.06 .* mu_vec .* (R_0 - 1)) ./ sqrt(N)
 
-bifurc_mu_seir_arr = zeros(Float64, size(u₀, 1), tlength, n_mus);
-bifurc_mu_change_arr = zeros(Float64, size(u₀, 1), tlength, n_mus);
+bifurc_mu_seir_arr = zeros(Float64, size(init_states, 1), tlength, n_mus);
+bifurc_mu_change_arr = zeros(Float64, size(init_states, 1), tlength, n_mus);
 bifurc_mu_jump_arr = zeros(Float64, 9, tlength, n_mus);
 
 prog = Progress(n_mus)
@@ -31,7 +31,7 @@ prog = Progress(n_mus)
 
     params = (beta_mean, beta_force, sigma, gamma, mu_run, epsilon, R_0)
     seir_mod!(seir, change, jump,
-        u₀, params, trange; dt = τ, type = "det",
+        init_states, params, trange; dt = τ, type = "det",
     )
     next!(prog)
 end
@@ -76,8 +76,8 @@ n_beta_forces = length(beta_force_min:beta_force_step:beta_force_max)
 beta_force_vec = zeros(Float64, n_beta_forces)
 beta_force_vec .= collect(beta_force_min:beta_force_step:beta_force_max)
 
-bifurc_beta_force_seir_arr = zeros(Float64, size(u₀, 1), tlength, n_beta_forces);
-bifurc_beta_force_change_arr = zeros(Float64, size(u₀, 1), tlength, n_beta_forces);
+bifurc_beta_force_seir_arr = zeros(Float64, size(init_states, 1), tlength, n_beta_forces);
+bifurc_beta_force_change_arr = zeros(Float64, size(init_states, 1), tlength, n_beta_forces);
 bifurc_beta_force_jump_arr = zeros(Float64, 9, tlength, n_beta_forces);
 
 mu = 50 / (1_000 * 365)
@@ -90,7 +90,7 @@ epsilon = 1.06 * mu * (R_0 - 1) / sqrt(N)
 
     local p = (beta_mean, beta_force, sigma, gamma, mu, epsilon, R_0)
 
-    seir_mod!(seir, change, jump, u₀, p, trange; dt = τ, type = "det")
+    seir_mod!(seir, change, jump, init_states, p, trange; dt = τ, type = "det")
 end
 
[…]

### 2023-08-16 src/Julia/SEIR-model.jl

@@ -29,7 +29,7 @@ the model can return an array of transmission rates if `retbetaarr = true`, and 
 
 ```julia-repl
 seir_array, change_array, jump_array, beta_arr = seir_mod(
-    u₀, p, trange; retbetaarr = true, type = "stoch"
+    init_states, p, trange; retbetaarr = true, type = "stoch"
 );
 ```
 """

### 2023-08-16 src/Julia/ensemble-functions.jl

@@ -20,7 +20,7 @@ function run_ensemble_jump_prob(params_dict; prog = prog)
                 "seasonal-infectivity-import",
                 "tau-leaping",
                 "N_$(p[:N])",
-                "r_$(p[:u₀_prop][:r])",
+                "r_$(p[:init_states_prop][:r])",
                 "nsims_$(p[:nsims])",
                 "births_per_k_$(p[:births_per_k])",
                 "beta_force_$(p[:beta_force])",
@@ -32,9 +32,9 @@ function run_ensemble_jump_prob(params_dict; prog = prog)
                 p;
                 allowedtypes = (Symbol, Dict, String, Real),
                 accesses = [
-                    :N, :u₀_prop, :nsims, :tmax, :dt, :births_per_k, :beta_force
+                    :N, :init_states_prop, :nsims, :tmax, :dt, :births_per_k, :beta_force
                 ],
-                expand = ["u₀_prop"],
+                expand = ["init_states_prop"],
                 sort = false,
             ),
             loadfile = false
@@ -47,14 +47,14 @@ end
     run_jump_prob(param_dict)
 """
 function run_jump_prob(param_dict)
-    @unpack N, u₀_prop, transmission_p, time_p, nsims, dt, beta_force,
+    @unpack N, init_states_prop, transmission_p, time_p, nsims, dt, beta_force,
     births_per_k, seed = param_dict
-    @unpack s, e, i, r = u₀_prop
+    @unpack s, e, i, r = init_states_prop
     @unpack R_0, sigma, gamma = transmission_p
     @unpack tstep, tlength, trange = time_p
 
-    u₀ = convert.(Int64, [s * N, e * N, i * N, r * N, N])
-    u0_dict = Dict(zip([:S, :E, :I, :R, :N], u₀))
+    init_states = convert.(Int64, [s * N, e * N, i * N, r * N, N])
+    init_states_dict = Dict(zip([:S, :E, :I, :R, :N], init_states))
 
     mu = births_per_k / (1_000 * 365)
     beta_mean = calculate_beta(R_0, gamma, mu, 1, N)
@@ -62,8 +62,8 @@ function run_jump_prob(param_dict)
 
     p = (beta_mean, beta_force, sigma, gamma, mu, epsilon, R_0)
 
-    ensemble_seir_arr = zeros(Int64, size(u₀, 1), tlength, nsims)
-    ensemble_change_arr = zeros(Int64, size(u₀, 1), tlength, nsims)
[…]

### 2023-08-29 src/Julia/ensemble-functions.jl
Correct ensemble unpacking of time and initial state parameters

@@ -116,7 +116,7 @@ function summarize_ensemble_jump_prob(params_dict; prog = prog)
                 "seasonal-infectivity-import",
                 "tau-leaping",
                 "N_$(p[:N])",
-                "r_$(p[:init_states_prop][:r])",
+                "r_$(p[:init_states_prop][:r_prop])",
                 "nsims_$(p[:nsims])",
                 "births_per_k_$(p[:births_per_k])",
                 "beta_force_$(p[:beta_force])",
@@ -144,7 +144,7 @@ end
     jump_prob_summary(param_dict)
 """
 function jump_prob_summary(param_dict)
-    @unpack N, init_states_prop, nsims, beta_force, births_per_k, quantiles =
+    @unpack N, init_states_prop, nsims, beta_force, births_per_k, time_p, quantiles =
         param_dict
 
     sim_name = savename(
@@ -168,7 +168,7 @@ function jump_prob_summary(param_dict)
             "seasonal-infectivity-import",
             "tau-leaping",
             "N_$N",
-            "r_$(init_states_prop.r)",
+            "r_$(init_states_prop[:r_prop])",
             "nsims_$nsims",
             "births_per_k_$births_per_k",
             "beta_force_$beta_force",

### 2023-08-29 notebooks/Julia/simulation-detection.qmd

@@
 tmax = 10.0
-δt = 0.1/365
+tstep = 0.1/365
 tspan = (0.0, tmax)
-tlength = length(tspan[1]:δt:tspan[2])
+tlength = length(tspan[1]:tstep:tspan[2])
 Random.seed!(1234);
@@
         )
-    jump_sol = solve(jump_prob, SSAStepper(), saveat = δt)
+    jump_sol = solve(jump_prob, SSAStepper(), saveat = tstep)
     return jump_sol;

### 2024-09-26 manuscript/manuscript.qmd
Rewrite manuscript using quarto

---
title: "The Need to Develop a Holistic Infectious Disease Surveillance System"
header-title: true
authors:
  - name: "Callum R.K. Arnold"
    affiliations:
      - ref: "PSU-Bio"
      - ref: "CIDD"
    corresponding: true
    email: "cfa5228@psu.edu"
  - name: "Alex C. Kong"
    affiliations:
      - ref: "Hopkins-IH"
  - name: "Amy K. Winter"
    affiliations:
      - ref: "UGA"
  - name: "William J. Moss"
    affiliations:
      - ref: "Hopkins-IH"
      - ref: "Hopkins-Epi"
  - name: "Bryan N. Patenaude"
    affiliations:
      - ref: "Hopkins-IH"
  - name: "Matthew J. Ferrari"
    affiliations:
      - ref: "PSU-Bio"
      - ref: "CIDD"
affiliations:
  - id: "PSU-Bio"
    address: "Department of Biology, Pennsylvania State University, University Park, PA, USA 16802"
  - id: "CIDD"
    address: "Center for Infectious Disease Dynamics, Pennsylvania State University, University Park, PA, USA 16802"
  - id: "Hopkins-IH"
    address: "Department of International Health, Johns Hopkins Bloomberg School of Public Health, Baltimore, MD, USA 21205"
  - id: "UGA"
    address: "Department of Epidemiology, College of Public Health, University of Georgia, Athens, GA, USA 30602"
  - id: "Hopkins-Epi"
    address: "Department of Epidemiology, Johns Hopkins Bloomberg School of Public Health, Baltimore, MD, USA 21205"
execute:
  echo: false
  warning: false
  message: false
# bibliography: /Users/cfa5228/Documents/Repos/OutbreakDetection/manuscript/OD.bib
bibliography: OD.bib
bibliographystyle: springer-vancouver
keywords:
  - "Rapid-Diagnostic Tests"
  - "ELISA"
  - "Infectious Disease Surveillance"
  - "Outbreak Detection"
word-count: true
line-numbers: true
abstract: |
  ## Background

  Infectious disease surveillance and outbreak detection systems often utilize diagnostic testing to validate case identification.
  The metrics of sensitivity, specificity, and positive predictive value are commonly discussed when evaluating the performance of diagnostic tests, and to a lesser degree, the performance of outbreak detection systems.
  However, the interaction of the two levels’ (the test and the alert system) metrics, is typically overlooked.
  Here, we describe how equivalent regions of detection accuracy can exist over a range of diagnostic test characteristics, examining the sensitivity to background noise structure.

  ## Methods

  We generated a stochastic SEIR model with importation to simulate true measles and non-measles sources of febrile rash (noise) daily incidence.
  We simulated the noise to have in-phase, out-of-phase, non-seasonal dynamics, with respect to the measles simulation.
  We also examined noise where each day were independent Poisson draws.
  For rapid diagnostic test (RDT) and enzyme-linked immunosorbent assay (ELISA) like tests, we varied the reporting rate, and the observed daily incidence threshold required to trigger an alert, and computed the: percentages of alerts that were correct (analogous to the positive predictive value), the percentage of outbreaks detected (analogous to the sensitivity), the delay between outbreak initiation and alert, and the number of (un)avoidable outbreak cases.
  Summarizing the percentage of alerts correct and outbreaks detected into an accuracy metric, we selected the “optimal threshold” for each combination of diagnostic test and reporting rate, presenting the summary statistics above.

  ## Results

  The optimal threshold for each test increased monotonically as the percentage of clinic visited who were tested increased, though there are adjacent reporting rates that have identical thresholds due to the integer requirement for the daily threshold.
  With Poisson-only noise, similar outbreak detection accuracies could be achieved with imperfect RDT-like tests as with ELISA-like diagnostic tests (c.
  93%), given moderately high testing rates.
  With larger delays (14 days) between the ELISA test procurement and result date, RDTs could outperform the ELISA.
  Similar numbers of unavoidable cases and outbreak alert delays could be achieved between the test types.
  With dynamical, in-phase, noise, however, the accuracy of ELISA scenarios was far superior to those achieved with RDTs (c.
  93% vs. 73%). For dynamical noise, RDT-based scenarios typically favored more sensitive alert threshold than ELISA-based scenarios (at a given reporting rate), observed with lower numbers of unavoidable cases and detection delays.

  ## Conclusions

  The performance of an outbreak detection system is highly sensitive to the structure and the level of background noise.
  Under the assumption that the noise is relatively static over time (as influenza-like illness was early in the COVID-19 pandemic), RDTs can perform as well as ELISA in a surveillance system.
  However, when the noise is non-stationary, imperfect tests cannot overcome their accuracy limitations through higher testing rates. 
toc: false
tbl-cap-location: bottom
format:
  generic-manuscript-typst:
    keep-typ: true
---

# Background

At the heart of an outbreak detection system is a surveillance program built upon case detection, often utilizing individual diagnostic tests as necessary components of epidemiological investigations before the declaration of an outbreak [@murrayInfectiousDiseaseSurveillance2017; @zhouEliminationTropicalDisease2013; @pahoIntegratedApproachCommunicable2000; @worldhealthorganizationMeaslesOutbreakGuide2022; @craggOutbreakResponse2018].
When diseases have non-specific symptoms, without accurate measurement tools we cannot have confidence in our ability to correctly ascribe changes in symptoms within a population to a particular disease, and therefore detect outbreaks of specific pathogens.
As a result, it has been commonplace for surveillance systems to be developed around high-accuracy tests, such as Polymerase Chain Reaction (PCR) tests and immunoglobulin (Ig) tests, when financially and logistically feasible [@gastanaduyMeasles2019; @commissionerCoronavirusCOVID19Update2020; @grasslyComparisonMolecularTesting2020; @ezhilanSARSCoVMERSCoVSARSCoV22021; @worldhealthorganizationCholera2023; @essentialprogrammeonimmunizationepiimmunizationClinicalSpecimensLaboratory2018].
Depending on the disease in question, either sensitivity (the ability to correctly detect a true positive individual) or specificity (the ability to correctly discount a true negative) will be prioritized, as they often are at odds with each other [@westreichDiagnosticTestingScreening2019; @shrefflerDiagnosticTestingAccuracy2024; @parikhUnderstandingUsingSensitivity2008].
This balance is commonly defined within the Target Product Profile (TPP) of a test [@worldhealthorganizationTargetProductProfiles], which is a set of minimum characteristics that should be met for production and widespread use, helping to guide research and development.
For example, in the wake of the 2013 Ebola outbreak in Guinea a TPP was developed that listed the minimum acceptable sensitivity of 95% and specificity of 99% [@chuaCaseImprovedDiagnostic2015].
Recognizing that Ebola is not the major cause of fever and other non-specific symptoms in the region, it is arguably more important to prioritize the specificity of the disease, although the authors note that the severity of the infection requires a high level of sensitivity as the consequences of a missed case are dire at an individual and population level [@chuaCaseImprovedDiagnostic2015].

Much like the accuracy of an individual test, outbreak detection systems face the same issue regarding the prioritization of sensitive vs specific alerts [@germanSensitivityPredictiveValue2000; @worldhealthorganizationOperationalThresholds2014; @lewisTimelyDetectionMeningococcal2001].
For many disease systems, particularly in resource constrained environments where the burden of infectious diseases is highest [@gbd2019childandadolescentcommunicablediseasecollaboratorsUnfinishedAgendaCommunicable2023; @roserBurdenDisease2023], cases are counted and if they breach a threshold, be that weekly, monthly, or some combination of the two, an alert is triggered that may launch a further investigation and/or a response [@worldhealthorganizationMeaslesOutbreakGuide2022; @worldhealthorganizationOperationalThresholds2014].
In effect, this discretizes a distinctly continuous phenomenon into a binary measure, outbreak or no outbreak, for decision making purposes.
When considering reactive vaccination campaigns and non-pharmaceutical based interventions designed to reduce transmission and limit and suppress outbreaks through direct and indirect protection, early action has the potential to avert the most cases, and in some instances, can be the most cost effective as it provides the opportunity to incorporate later information within an “adaptive management” framework [@atkinsAnticipatingFutureLearning2020; @taoLogisticalConstraintsLead; @graisTimeEssenceExploring2008; @ferrariTimeStillEssence2014; @worldhealthorganizationConfirmingInvestigatingManaging2009; @minettiLessonsChallengesMeasles2013].
While this framing would point towards a sensitive surveillance system being optimal, with the benefits clearly defined in terms of morbidity and mortality avoided, each action comes with both direct and indirect financial and opportunity costs stemming from unnecessary activities that remove resources from future response activities.
Just as the balance of sensitivity and specificity of a test must be carefully evaluated, so must the balance at the outbreak level.

The concept of using incidence-based alert triggers to define the discrete event of an “outbreak” with characteristics analogous to individual tests has been well documented in the case of meningitis, measles, and malaria [@worldhealthorganizationMeaslesOutbreakGuide2022; @lewisTimelyDetectionMeningococcal2001; @worldhealthorganizationConfirmingInvestigatingManaging2009; @trotterResponseThresholdsEpidemic2015; @cooperReactiveVaccinationControl2019; @zalwangoEvaluationMalariaOutbreak2024; @kanindaEffectivenessIncidenceThresholds2000].
However, an overlooked, yet critical, aspect of an outbreak detection system is the interplay between the individual test and outbreak alert characteristics.
With their success within malaria surveillance systems, and particularly since the COVID-19 pandemic, rapid diagnostic tests (RDTs) have garnered wider acceptance, and their potential for use in other disease systems has been gaining interest [@warrenerEvaluationRapidDiagnostic2023].
Despite concerns of their lower diagnostic accuracy slowing their adoption until recently [@millerAddressingBarriersDevelopment2015], the reduced cold-chain requirements [@brownRapidDiagnosticTests2020], and gains in speed of result and the impact on the timeliness of an alert system may outweigh the costs in some settings [@warrenerEvaluationRapidDiagnostic2023; @mcmorrowMalariaRapidDiagnostic2011; @larremoreTestSensitivitySecondary2021; @middletonModelingTransmissionMitigation2023].

In this paper we examine how the interaction between imperfect diagnostic tests affects the optimal outbreak alert method for the detection of measles outbreaks, in the context of a febrile rash surveillance system that detects and tests suspected measles cases i.e., an individual with fever and maculopapular (non-vesicular) rash (or who the healthcare worker suspects has measles) [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018].
Because measles symptoms are non-specific, it is important to account for non-measles sources of febrile rash e.g., rubella, parvovirus, varicella, etc., producing the potential for false positive results in the context of imperfect tests.
Currently, measles outbreaks are declared on the basis of either suspected measles cases alone, the use of an enzyme-linked immunosorbent assay (ELISA) to detect the presence of measles-specific IgM antibodies, or a combination of the two, depending on the elimination status of the region, with countries nearer elimination increasing the use of ELISA diagnostic confirmation for suspected cases [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018].
Each of these detection systems have their flaws.
Although clinical case definition is very fast and requires limited resources, providing a highly sensitive alert system, in the face of high “background noise” from non-measles sources of febrile rash relative to true measles prevalence, it can lead to great uncertainty in the alert system and low positive predictive value (PPV), i.e., the probability that an alert accurately captures an outbreak [@hutchinsEvaluationMeaslesClinical2004].
And while ELISA confirmation is the standard diagnostic test for measles surveillance [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018], it is expensive due to its cold-chain and transport costs and delays resulting from the requirement to transfer samples from point of care to a central laboratory, leading to a slower and less sensitive (but more specific) alert system.
A rapid diagnostic test offers the opportunity to provide a compromise between diagnostic accuracy and timeliness, and recent developments show encouraging signs in the field as well as in theory [@warrenerEvaluationRapidDiagnostic2023; @brownRapidDiagnosticTests2020].

By examining the combination of alert threshold and individual test characteristic in a modeling study that explicitly incorporates dynamical background noise, we aim to illustrate the need to develop a TPP for the whole detection system, not just one component.
To evaluate the alert system performance, we develop a set of outbreak definition criteria and surveillance metrics, drawing inspiration from acceptance sampling, ecological surveillance systems, and epidemiological surveillance systems guidelines and reviews [@yemshanovAcceptanceSamplingCosteffective2020; @christensenHerdlevelInterpretationTest2000; @muratoEvaluationSamplingMethods2020; @sternAutomatedOutbreakDetection1999; @calbaSurveillanceSystemsEvaluation2015; @guidelinesworkinggroupUpdatedGuidelinesEvaluating2001].
Using these metrics we overcome issues encountered by early warning systems that rely on dynamical values such as $R_{\text{effective}}$ in defining outbreaks [@jombartRealtimeMonitoringCOVID192021; @proverbioPerformanceEarlyWarning2022; @stolermanUsingDigitalTraces2023; @brettAnticipatingEpidemicTransitions2018; @salmonMonitoringCountTime2016], for example characterizing the end of an epidemic period is important in a timeseries where multiple outbreaks will occur.

# Methods

## Model Structure

We constructed a stochastic compartmental non-age structured SEIR model of measles and simulated using a modified Tau-leaping algorithm with a time step of 1 day, utilizing binomial draws to ensure compartment sizes remained positive valued [@chatterjeeBinomialDistributionBased2005; @gillespieApproximateAcceleratedStochastic2001].
To stochastically reintroduce infections and cause recurring epidemics, commuter-style imports were added that are proportional to the size of the population and $R_{0}$ [@keelingModelingInfectiousDiseases2008], and the transmission parameter ($\beta_{t}$) is sinusoidal with a period of one year.
$R_{0}$ was set to 16, with a latent period of 10 days and infectious period of 8 days. The population was initialized with 500_000 individuals with Ghana-like birth and vaccination rates, and the final results were scaled up to the 2022 population size of Ghana (33 million). The full table of parameters can be found in @tbl-model-parameters. All simulations and analysis was completed in Julia version 1.10.3 [@bezansonJuliaFreshApproach2017], with all code stored at [https://github.com/arnold-c/OutbreakDetection](https://github.com/arnold-c/OutbreakDetection).

::: {#tbl-model-parameters}

```{=typst}
#let parameters = csv("parameters.csv")
#let import_rate = $(1.06*μ*R_0)/(√(N))$
#let parameter_labels = ( "Parameters", $R_0$, $"Latent period ("#sym.sigma")"$, $"Infectious period ("#sym.gamma")"$, "Seasonal amplitude", $"Birth rate ("#sym.mu")"$, $"Vaccination rate at birth ("#sym.rho")"$)

#figure(
	table(
		columns: 3,
		..for ((label), (.., measles, nonmeasles)) in parameter_labels.zip(parameters) {
			(label, measles, nonmeasles)
		},
		[Importation rate], table.cell(colspan: 2, align: center, $(1.06*μ*R_0)/(√(N))$),
		[Population size (N)], table.cell(colspan: 2, align: center, "500_000, scaled to 33M"),
		[Initial proportion susceptible], table.cell(colspan: 2, align: center, "0.05"),
		[Initial proportion exposed], table.cell(colspan: 2, align: center, "0.0"),
		[Initial proportion infected], table.cell(colspan: 2, align: center, "0.0"),
		[Initial proportion recovered], table.cell(colspan: 2, align: center, "0.95"),
	)
)
```

Compartmental model parameters

:::

To examine the sensitivity of the detection system to background noise, we layered the measles incidence timeseries with one of four noise time series: Poisson-only noise, or dynamical noise with rubella-like parameters that could be in- or out-of-phase or with independent seasonality to the measles dynamics.
The total annual incidence of noise in each system was approximately 8 times the average annual measles incidence, with the dynamical noise systems also containing low levels of Poisson noise, reflecting Ghana-like dynamics.
Despite the average incidence of noise in each scenario being the same, the relative proportion of measles to noise varies differently throughout the timeseries, exacerbating the effects of imperfect diagnostic tests that will produce higher rates of false positives and negatives than ELISA-like diagnostics.
Imperfect tests are also more susceptible to incorrect results because of the varying prevalence throughout the timeseries, so the PPV of the system will not be static throughout the timeseries.
To account for the stochasticity of the model, we simulated 100 time series for each scenario, each lasting 100 years, before summarizing the distributions of outbreak detection methods.

## Defining Outbreaks

It is common to use expert review to define outbreaks when examining empirical data, but this is not feasible in a modeling study where tens of thousands of years are being simulated.
To account for this, many studies only simulate a single outbreak within a timeseries (repeating this short stochastic simulation multiple times to ensemble results), define an outbreak as a period where $R_{\text{effective}}$ > 1, or use a threshold of > 2 standard deviations (s.d.) over the mean seasonal incidence observed in empirical data [@sternAutomatedOutbreakDetection1999;; @jombartRealtimeMonitoringCOVID192021; @stolermanUsingDigitalTraces2023; @salmonMonitoringCountTime2016; @teklehaimanotAlertThresholdAlgorithms2004; @leclereAutomatedDetectionHospital2017].
Each method has its uses, but to evaluate the performance of an outbreak detection system in an endemic region where multiple sequential epidemics are expected it is important to clearly define the bounds of the outbreak, which can only be achieved by one of these methods (2 s.d. > mean).
This, however, assumes strong seasonal forcing and regular periodicity of incidence to produce a smooth enough baseline, which is not present as countries near measles elimination status [@grahamMeaslesCanonicalPath2019].
Here we define a true measles outbreak as a region of the timeseries that meets the following three criteria:

- The daily measles incidence must be greater than, or equal to, 5 cases
- The daily measles incidence must remain above 5 cases for greater than, or equal to, 30 consecutive days
- The total measles incidence must be great than, or equal to, 500 cases within the bounds of the outbreak

Each of these is a necessary, but not sufficient, condition to the definition of an outbreak; all must be met.
For example, a region may produce 1000 total cases within the lower and upper bounds defined by the daily incidence being $\ge$ 5 cases, but if this period only lasts for 20 days it would not be considered an outbreak as it would not be possible to mount a response to.
The incidence of non-measles febrile rash (i.e., noise) does not affect the outbreak status of a region but does affect the alert status triggered by the testing protocol.

## Triggering Alerts

The use of alert thresholds is common in locations burdened by measles, where access to high performance computing is limited making nowcasting-style approaches to outbreak detection impractical.
For this reason, we define an “alert” as any single day where the 7-day moving average of the daily incidence is greater than, or equal to, a threshold of X.
To examine the interaction of test and surveillance systems we vary the alert threshold, X, between 1 and 15 cases per day, before computing the evaluation metrics and identifying the optimal threshold for each individual test.
We also examined an alert method that required either the daily incidence or the 7-day moving average of daily incidence to exceed the alert threshold X (see supplemental results).

Each day, 60% of the measles and non-measles febrile rash cases visit the clinic for treatment, and Y% percentage of these clinic visits are tested, as all clinic visits are deemed to be suspected measles cases because they meet the clinical case definition.
This percentage of clinic visits, Y, that are tested is varied between 10% and 60%, in 10% increments, for all combinations of diagnostic test (apart from the clinical case definition) and alert threshold, defining the “testing scenario”.
While a reporting rate of 6-48% is high, we are only simulating a population of 500_000 individuals i.e., approximately 1.5% of Ghana’s population, which produces absolute incidence values that are consistent with observed reporting on the national scale. Each testing scenario uses one of the following tests:

- No test, where every suspected measles case is counted as positive for measles and used to trigger an outbreak alert, which is equivalent to the use of the clinical case definition. This is implemented as a diagnostic test with 100% sensitivity, 0% specificity, a 0-day lag in returning the result, and all individuals who visit the clinic are tested (not varied between 10-60%, like the other diagnostic tests described).
- An RDT equivalent with 85% sensitivity and specificity, and 0-day lag in result return. That is, 85% of true measles cases will be correctly labelled as positive, and 15% of non-measles febrile rash individuals that are tested will be incorrectly labelled as positive for measles. This acts as a lower bound of acceptability for a new measles RDT.
- An ELISA equivalent test with 100% sensitivity and specificity with a 0-day test result delay.
- An ELISA equivalent test with 100% sensitivity and specificity with a 7-day test result delay.
- An ELISA equivalent test with 100% sensitivity and specificity with a 14-day test result delay.

## Optimal Thresholds

For each of the previously defined tests, we calculate the number of test positive cases (that will include false positive and negative cases resulting from imperfect diagnostic tests), which is used to categorize the time series by alert status in conjunction with the alert threshold X.
The overlap with true outbreak status, or lack thereof, is used to classify the following metrics, and has been illustrated in @fig-outbreak-schematic:

- The percentage of alerts that are correct. This is analogous to the PPV of the alert system. Note that it is possible to trigger multiple alerts within a single outbreak, and each would be considered correct.
- The percentage of outbreaks that are detected. This is analogous to the sensitivity of the alert system.
- The detection delay. Note that if an alert precedes the start of the outbreak, as long as it successfully “captures” the outbreak within its bounds, it is considered to be correct, resulting in a negative detection delay i.e., an early warning triggered by false positives.
- The number of unavoidable and avoidable outbreak cases and deaths. Avoidable cases are defined as those that occur within an outbreak after a correct alert is first triggered i.e., cases that could theoretically be prevented with a perfectly effective and timely response. Unavoidable cases are the inverse: those that occur before a correct alert, or those that occur in an undetected outbreak. Using a fitted age distribution of measles cases in Ghana [@mintaProgressMeaslesElimination2023] and the age-specific case fatality ratio for Ghana in 2022 [@sbarraEstimatingNationallevelMeasles2023], we scaled the values for each one-year age cohort, per annum. In practice, not all cases deemed avoidable are (due to imperfect and delays in responses), but to minimize the sensitivity of the results to the response implementation and operational constraints we are counting them as such.

![A schematic of the outbreak definition and alert detection system. A) Noise time series. B) Measles incidence time series. C) Observed time series resulting from testing noise & measles cases that visit the healthcare facility. The orange bands/vertical lines represent regions of the measles time series that meet the outbreak definition criteria. The green bands/vertical lines represent regions of the observed (measles - noise) time series that breach the alert threshold (the horizontal dashed line), and constitute an alert.](./schematic-simulation_with-shade.png){#fig-outbreak-schematic}

For each combination of diagnostic test and testing rate, the optimal alert threshold is calculated by selecting the threshold that produces the highest accuracy.
Each of the 100 simulations produces an accuracy (the mean of the percentages of alerts that are correct and the percentage of the outbreaks that are detected), and the median accuracy of this distribution is used to determine the optimal alert threshold.
Each testing scenario (the combination of the diagnostic test, testing rate, and alert threshold), is then compared (using the metrics above) at the optimal threshold that is specific for the scenario.
This allows for conclusions to be made about the surveillance system as a whole, instead of just single components.

# Results

## Optimal Thresholds, Accuracy, and Unavoidable Cases

The threshold that maximized outbreak detection accuracy given a testing scenario depends heavily on the noise structure.
For example, the optimal threshold for an RDT with 90% sensitivity/specificity and 40% testing of clinic visits was 4 cases per day with the dynamical in-phase noise, whereas it is 5 cases per day when Poisson-only noise was simulated (@tbl-optimal-thresholds).
This corresponds to an accuracy of 72% and 93%, respectively (@tbl-optimal-thresholds-accuracy).
Much higher accuracy can be achieved with imperfect tests under Poisson noise because large spikes of non-measles febrile rash do not occur that would lead to many false positives, triggering an erroneous alert.
With dynamical noise, this possibility can and does occur, and when the ratio of measles to non-measles cases is sufficiently low, drastically impacts the accuracy of an alert system.
This is why the values observed among ELISA tests are identical between the different noise structures: with a perfect test there are no false positive results.
For dynamical noise, RDTs never perform as well as an ELISA, even scenarios that include a 14-day lag between test procurement and result receipt.
However, with Poisson noise, once testing reaches 20% of clinic visits, equivalence can be observed at approximately 91% accuracy (@tbl-optimal-thresholds-accuracy).

::: {#tbl-optimal-thresholds}

```{=typst}
#let optimal_thresholds = csv("optimal-thresholds.csv")

#figure(
	table(
		columns: 10,
		  fill: (x, y) => {
		    if y == 0 {gray}
		    if y == 1 {gray}
		  },
		[], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
		..optimal_thresholds.flatten()
	)
)
```

Optimal threshold for each testing scenario. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only

:::

Moving across a single row of @tbl-optimal-thresholds, as the testing rate increases, so does the optimal threshold, because more individuals are testing positive.
However, the changes in accuracy are not monotonically increasing, as one might anticipate.
Even with a perfect test, there are sequential cells that increase the testing rate, yet accuracy decreases. For example, the ELISA equivalent with a 0-day test result lag goes from an outbreak detection accuracy of 91% down to 87% when testing increases from 30% to 40% of clinic visits.
The reason behind this unintuitive result stems from the use of integer valued thresholds.
While computing the moving average incidence helps to smooth out the alert system, requiring the daily alert to be an integer value results in the optimal threshold for alert accuracy being stuck at a lower level than a truly optimal intermediate floating point number, because the increase in the alert system’s PPV from a higher threshold is outweighed by the larger decrease in alert sensitivity, or the threshold is forced to increase.
Examining the number of unavoidable cases (@tbl-optimal-thresholds-unavoidable) and the detection delays (@tbl-optimal-thresholds-delays), we can see that the decrease in accuracy results from a decrease in the sensitivity of the alert system: the detection delay increases by 6 days, and the unavoidable cases by c. 2300.
Although the testing rate increases, so does the optimal threshold, from 3 test positives to 4 test positives per day.

::: {#tbl-optimal-thresholds-accuracy}

```{=typst}
#let optimal_thresholds = csv("optimal-thresholds-accuracy.csv")

#figure(
	table(
		columns: 10,
		  fill: (x, y) => {
		    if y == 0 {gray}
		    if y == 1 {gray}
		  },
		[], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
		..optimal_thresholds.flatten()
	)
)
```

Optimal threshold for each testing scenario. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only

:::

::: {#tbl-optimal-thresholds-unavoidable}

```{=typst}
#let optimal_thresholds = csv("optimal-thresholds-unavoidable.csv")

#figure(
	table(
		columns: 10,
		  fill: (x, y) => {
		    if y == 0 {gray}
		    if y == 1 {gray}
		  },
		[], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
		..optimal_thresholds.flatten()
	)
)
```

Unavoidable cases per annum of each testing scenario at their specific optimal thresholds, scaled up to Ghana’s 2022 population. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only.

:::

::: {#tbl-optimal-thresholds-delays}

```{=typst}
#let optimal_thresholds = csv("optimal-thresholds-delays.csv")

#figure(
	table(
		columns: 10,
		  fill: (x, y) => {
		    if y == 0 {gray}
		    if y == 1 {gray}
		  },
		[], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
		..optimal_thresholds.flatten()
	)
)
```

Outbreak alert delay (days) of each testing scenario at their specific optimal thresholds. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only.

:::

# Discussion

The performance of an outbreak detection system is highly sensitive to the structure and level of background noise in the simulation.
Despite the mean daily noise incidence set to equivalent values between the dynamical and Poisson-only simulations, drastically different results are observed.
Under the assumption that non-measles febrile rash is relatively static in time (Poisson), RDTs can perform as well, if not better than ELISA tests at moderate to high testing rates, and at a fraction of the cost.
However, if it is expected that the noise is non-stationary, imperfect tests cannot overcome their accuracy limitations through higher testing rates, saturating at c. 74% accuracy, relative to ELISA’s 93%.

The purpose of routine surveillance is to characterize the infection landscape.
With strong public health infrastructure and infectious disease surveillance programs, it is possible to develop a strong understanding of the shape of febrile rash cases, regardless of source.
With this information, countries can tailor their future activities to rely more or less heavily upon RDTs, depending on the dynamics of the target disease and its relationship to background noise, favoring RDTs when there are low levels of noise and ELISAs during large rubella outbreaks, for example.

## Limitations and Strengths

To our knowledge, this is one of the first simulation studies to examine the relationship between individual test characteristics and the wider surveillance program.
By explicitly modeling the interaction between the two, we make a case that surveillance systems should take a holistic approach; ignoring one component can lead to drastically different, and suboptimal, results.
Additionally, by defining outbreak bounds concretely we have been able to calculate metrics of outbreak detection performance that draw parallels to those used when evaluating individual diagnostic tests, allowing for intuitive and simple implementation of this method in resource-constrained environments, something that is not possible with most outbreak detection and early warning system simulations in the literature.
An evaluation of all outbreak detection algorithms is beyond the scope of this work, but a more computationally expensive approach based on nowcasting incidence may help overcome the shortcomings of RDTs in high-noise scenarios.

For computational simplicity, this paper did not include demography in the model structure.
And while a simulation-based approach allows for complete determination of true infection status i.e., measle vs non-measles febrile rash cases, and therefore an accurate accounting of the outbreak and alert bounds, these simulations do not specifically represent any real-world setting.
The evaluation of empirical data does provide this opportunity, but at the cost of knowing the true infection status of individuals and confounding of multiple variables, limiting analysis to only those who are observed (i.e., not those in the community who do not visit a healthcare center), and removing the possibility to explore the sensitivity of the results to parameters of interest to a surveillance program e.g., testing rate, and the test itself.

Additionally, is has been well documented that the performance of an individual test is highly sensitive to its timing within a person’s infection cycle [@gastanaduyMeasles2019; @larremoreTestSensitivitySecondary2021; @middletonModelingTransmissionMitigation2023; @kisslerViralDynamicsAcute2021; @ratnamPerformanceIndirectImmunoglobulin2000], so it is possible that different conclusions would be drawn if temporal information about the test acquisition was included in the simulation.

Finally, the optimal threshold for a testing scenario depends heavily on the costs ascribed to incorrect actions, be that failing to detect an outbreak or incorrectly mounting a response for an outbreak that doesn’t exist.
In the simulations we have weighted them equally, but it is likely that they should not be deemed equivalent: missing an outbreak leads to many thousands of cases and associated DALYs, whereas an unnecessary alert would generally launch an initial low-cost investigation for full determination of the outbreak status.
This is particularly important in countries with vast heterogeneity in transmission: different weightings should be applied to higher vs. lower priority/risk regions to account for discrepancies in consequences of incorrect decisions.
For example, a high priority zone could still benefit from a false alert as many high-risk healthcare regions within a country are targeted for Supplemental Immunization Activities, so a false alert would just hasten the (proactive) vaccination response.

Given these limitations, the explicit values (i.e., optimal thresholds, accuracies etc.) should be interpreted with caution, and the exact results observed in the real-world will likely be highly dependent on unseen factors, such as the proportion of measles and non-measles sources of febrile rash that seek healthcare.
However, the general patterns should hold, and more importantly, the analysis framework provides a consistent and holistic approach to evaluating the tradeoff between individual level tests and the alert system enacted to detect outbreaks.

{{< pagebreak >}}

# Funding

This work was supported by funding from the Office of the Provost and the Clinical and Translational Science Institute, Huck Life Sciences Institute, and Social Science Research Institutes at the Pennsylvania State University.
The project described was supported by the National Center for Advancing Translational Sciences, National Institutes of Health, through Grant UL1 TR002014.
The content is solely the responsibility of the authors and does not necessarily represent the official views of the NIH. The funding sources had no role in the collection, analysis, interpretation, or writing of the report.

# Acknowledgements

## Author Contributions

*Conceptualization:* CA, MJF

*Data curation:* MJF, CA

*Formal analysis:* CA, MJF

*Funding acquisition:* MJF, WM, AW

*Investigation:* CA, MJF

*Methodology:* CA, MJF

*Project administration:* MJF

*Software:* CA

*Supervision:* MJF, WM, AW, BP

*Validation:* CA, MJF

*Visualization:* CA

*Writing - original draft:* CA, MJF

*Writing - review and editing:* all authors.

## Conflicts of Interest and Financial Disclosures

The authors declare no conflicts of interest.

## Data Access, Responsibility, and Analysis

Callum Arnold and Dr. Matthew J. Ferrari had full access to all the data in the study and take responsibility for the integrity of the data and the accuracy of the data analysis.
Callum Arnold and Dr. Matthew J. Ferrari (Department of Biology, Pennsylvania State University) conducted the data analysis.

## Data Availability

All code and data for the simulations can be found at [https://github.com/arnold-c/OutbreakDetection](https://github.com/arnold-c/OutbreakDetection).

{{< pagebreak >}}

::: {#refs}
:::

### 2024-09-26 manuscript/supplemental-appendix.qmd
Rewrite manuscript using quarto

---
title: "The Need to Develop a Holistic Infectious Disease Surveillance System"
header-title: true
authors:
  - name: "Callum R.K. Arnold"
    affiliations:
      - ref: "PSU-Bio"
      - ref: "CIDD"
    corresponding: true
    email: "cfa5228@psu.edu"
  - name: "Alex C. Kong"
    affiliations:
      - ref: "Hopkins-IH"
  - name: "Amy K. Winter"
    affiliations:
      - ref: "UGA"
  - name: "William J. Moss"
    affiliations:
      - ref: "Hopkins-IH"
      - ref: "Hopkins-Epi"
  - name: "Bryan N. Patenaude"
    affiliations:
      - ref: "Hopkins-IH"
  - name: "Matthew J. Ferrari"
    affiliations:
      - ref: "PSU-Bio"
      - ref: "CIDD"
affiliations:
  - id: "PSU-Bio"
    address: "Department of Biology, Pennsylvania State University, University Park, PA, USA 16802"
  - id: "CIDD"
    address: "Center for Infectious Disease Dynamics, Pennsylvania State University, University Park, PA, USA 16802"
  - id: "Hopkins-IH"
    address: "Department of International Health, Johns Hopkins Bloomberg School of Public Health, Baltimore, MD, USA 21205"
  - id: "UGA"
    address: "Department of Epidemiology, College of Public Health, University of Georgia, Athens, GA, USA 30602"
  - id: "Hopkins-Epi"
    address: "Department of Epidemiology, Johns Hopkins Bloomberg School of Public Health, Baltimore, MD, USA 21205"
execute:
  echo: false
  warning: false
  message: false
# bibliography: /Users/cfa5228/Documents/Repos/OutbreakDetection/manuscript/OD.bib
bibliography: OD.bib
bibliographystyle: springer-vancouver
toc: false
format:
  generic-manuscript-typst:
    keep-typ: true
---

# Results

### 2024-09-27 manuscript/manuscript.qmd
Include Julia scripts in manuscript file for plots

@@
     keep-typ: true
+jupyter: julia-1.10
 ---
@@
 
-![A schematic of the outbreak definition and alert detection system.
+```{julia}
+#| fig-cap: "A schematic of the outbreak definition and alert detection system.
 A) Noise time series.
@@
 The orange bands/vertical lines represent regions of the measles time series that meet the outbreak definition criteria.
-The green bands/vertical lines represent regions of the observed (measles - noise) time series that breach the alert threshold (the horizontal dashed line), and constitute an alert.](./schematic-simulation_with-shade.png){#fig-outbreak-schematic}
+The green bands/vertical lines represent regions of the observed (measles - noise) time series that breach the alert threshold (the horizontal dashed line), and constitute an alert."
+#| label: fig-outbreak-schematic
+include("schematic-plot.jl")
+```
 
@@
 
+```{julia}
+#| fig-cap: "The accuracy of outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| label: fig-accuracy
+include("line-plot.jl")
+```
+
 # Discussion

### 2024-09-27 OutbreakDetectionUtils/src/OutbreakDetectionUtils.jl

@@ -21,7 +21,8 @@ export POPULATION_N, LATENT_PER_DAYS, DUR_INF_DAYS, R0, SIGMA, GAMMA,
     MU, BETA_MEAN, BETA_FORCE, EPSILON
 
 include("test-constants.jl")
-export CLINICAL_CASE_TEST_SPEC, EPI_LINKED_CASE_TEST_SPEC, CLINICAL_TEST_SPECS, PROTOTYPE_RDT_TEST_SPECS
+export CLINICAL_CASE_TEST_SPEC,
+    EPI_LINKED_CASE_TEST_SPEC, CLINICAL_TEST_SPECS, PROTOTYPE_RDT_TEST_SPECS
 
 include("SEIR-model.jl")
 export seir_mod, seir_mod!, seir_mod_loop!,
@@ -33,7 +34,9 @@ export create_sir_df, create_sir_beta_dfs,
 
 include("detection-thresholds.jl")
 export create_inc_infec_arr,
-    create_inc_infec_arr!, calculate_outbreak_thresholds
+    create_inc_infec_arr!, calculate_outbreak_thresholds,
+    classify_all_outbreaks!, filter_only_outbreaks, calculate_positives,
+    calculate_true_positives!
 
 include("diag-testing-functions.jl")
 export create_testing_arrs, create_testing_arrs!, calculate_tested!,

### 2024-09-27 manuscript/line-plot.jl

@@ -1,0 +1,97 @@
+#%%
+using DrWatson
+@quickactivate "OutbreakDetection"
+
+using OutbreakDetectionUtils
+using OutbreakDetection
+using Revise
+
+include(srcdir("makie-plotting-setup.jl"))
+include(srcdir("ensemble-parameters.jl"))
+
+if false
+    include("../src/ensemble-parameters.jl")
+end
+
+#%%
+optimal_threshold_test_spec_vec = [
+    IndividualTestSpecification(0.85, 0.85, 0),
+    IndividualTestSpecification(0.9, 0.9, 0),
+    IndividualTestSpecification(1.0, 1.0, 0),
+    IndividualTestSpecification(1.0, 1.0, 14),
+]
+
+optimal_threshold_alertthreshold_vec = collect(1:1:15)
+
+R_0_vec = [16.0]
+
+ensemble_dynamics_spec_vec = create_combinations_vec(
+    DynamicsParameters,
+    (
+        [ensemble_state_specification.init_states.N],
+        [27],
+        [0.2],
+        [SIGMA],
+        [GAMMA],
+        R_0_vec,
+        [0.8],
+    ),
+)
+
+ensemble_spec_vec = create_combinations_vec(
+    EnsembleSpecification,
+    (
+        [ensemble_model_type],
+        [ensemble_state_specification],
+        ensemble_dynamics_spec_vec,
+        [ensemble_time_specification],
+        [ensemble_nsims],
+    ),
+)
+
+alert_method_vec = ["movingavg"]
+
+#%%
+ensemble_noise_specification =
+    filter(ensemble_noise_specification_vec) do noise_spec
+        noise_spec.noise_type == "poisson" ||
+            noise_spec.correlation == "in-phase"
+    end
+
+ensemble_specification = ensemble_spec_vec[1]
+alertmethod = alert_method_vec[1]
+
+optimal_threshold_core_params = (
+    alertthreshold_vec = optimal_threshold_alertthreshold_vec,
+    ensemble_specification = ensemble_specification,
+    outbreak_specification = ensemble_outbreak_specification,
+    moving_avg_detection_lag = ensemble_moving_avg_detection_lag,
+    percent_visit_clinic = ensemble_percent_visit_clinic,
+    alertmethod = alertmethod,
+)
+
+basedirpath = joinpath(
+    "R0_$(ensemble_specification.dynamics_parameters.R_0)"
[…]

### 2024-09-27 manuscript/schematic-plot.jl

@@ -1,7 +1,5 @@
-#%%
 using DrWatson
 @quickactivate "OutbreakDetection"
-using Revise: includet
 
 using GLMakie
 using StatsBase
@@ -9,11 +7,10 @@ using StatsBase
 using OutbreakDetection
 using OutbreakDetectionUtils
 
-includet(srcdir("makie-plotting-setup.jl"))
+include(srcdir("makie-plotting-setup.jl"))
 
-#%%
 states_p = StateParameters(;
-    N=500_000, s_prop=0.05, e_prop=0.00, i_prop=0.00
+    N = 500_000, s_prop = 0.05, e_prop = 0.00, i_prop = 0.00
 )
 
 dynamics_p = DynamicsParameters(
@@ -30,7 +27,7 @@ dynamics_p = DynamicsParameters(
 )
 
 noise_states_p = StateParameters(;
-    N=500_000, s_prop=0.15, e_prop=0.00, i_prop=0.00
+    N = 500_000, s_prop = 0.15, e_prop = 0.00, i_prop = 0.00
 )
 
 noise_dynamics_p = DynamicsParameters(
@@ -41,12 +38,16 @@ noise_dynamics_p = DynamicsParameters(
     1 / 14,
     5.0,
     0.65;
-    seasonality=sin
+    seasonality = sin,
 )
 
-test_specification = IndividualTestSpecification(0.85, 0.85, 0)
+test_specification = IndividualTestSpecification(
+    0.85, 0.85, 0
+)
 
-time_p = SimTimeParameters(; tmin=0.0, tmax=365.0 * 20, tstep=1.0)
+time_p = SimTimeParameters(;
+    tmin = 0.0, tmax = 365.0 * 20, tstep = 1.0
+)
 
 outbreak_specification = OutbreakSpecification(5, 30, 500)
 
@@ -57,7 +58,7 @@ outbreak_detection_specification = OutbreakDetectionSpecification(
     movingavg_window,
     1.0,
     0.75,
-    "movingavg"
+    "movingavg",
 )
 
 function get_outbreak_status(
@@ -69,7 +70,9 @@ function get_outbreak_status(
 
     abovethresholdrle = rle(abovethreshold_vec)
 
-    all_outbreak_bounds = calculate_outbreak_thresholds(abovethresholdrle)
+    all_outbreak_bounds = calculate_outbreak_thresholds(
+        abovethresholdrle
+    )
 
     classify_all_outbreaks!(
         inc_vec,
[…]

### 2024-09-27 src/line_plots.jl

@@ -20,11 +20,13 @@ function line_accuracy_plot(
     colors = lineplot_colors,
     xlabel = "Proportion Tested",
     ylabel = "Outbreak Detection\nAccuracy",
+    facet_fontsize = 24,
     labelsize = 30,
     show_x_facet_label = true,
     show_y_facet_label = true,
     clinical_hline = true,
     force = false,
+    save_plot = true,
     kwargs...,
 )
     mkpath(plotdirpath)
@@ -74,6 +76,7 @@ function line_accuracy_plot(
                     colors = colors,
                     xlabel = xlabel,
                     ylabel = ylabel,
+                    facet_fontsize = facet_fontsize,
                     show_x_facet_label = show_x_facet_label,
                     kwargs...,
                 )
@@ -111,7 +114,9 @@ function line_accuracy_plot(
         )
         rowsize!(fig.layout, 0, Relative(0.03))
 
-        Makie.save(plotpath, fig; size = size)
+        if save_plot
+            Makie.save(plotpath, fig; size = size)
+        end
         return fig
     end
 
@@ -129,6 +134,7 @@ function _line_accuracy_plot!(
     xlabel = "Proportion Tested",
     ylabel = "Accuracy",
     show_x_facet_label = true,
+    facet_fontsize = 24,
     num_noise_descriptions = 1,
     kwargs...,
 )
@@ -181,6 +187,7 @@ function _line_accuracy_plot!(
         colors = colors,
         xlabel = xlabel,
         ylabel = ylabel,
+        facet_fontsize = facet_fontsize,
         kwargs_dict...,
     )
 
@@ -195,6 +202,7 @@ function _line_accuracy_facet!(
     colors = lineplot_colors,
     xlabel = "Proportion Tested",
     ylabel = "Accuracy",
+    facet_fontsize = 24,
     kwargs...,
 )
     kwargs_dict = Dict(kwargs)
@@ -213,7 +221,7 @@ function _line_accuracy_facet!(
         Label(
             gl[1, xpos],
             kwargs_dict[:x_facet_label];
-            fontsize = 24,
+            fontsize = facet_fontsize,
             padding = (0, 0, 0, 0),
             valign = :bottom,
             tellwidth = false,

### 2024-09-27 manuscript/manuscript.qmd
Update background and methods sections

@@
 At the heart of an outbreak detection system is a surveillance program built upon case detection, often utilizing individual diagnostic tests as necessary components of epidemiological investigations before the declaration of an outbreak [@murrayInfectiousDiseaseSurveillance2017; @zhouEliminationTropicalDisease2013; @pahoIntegratedApproachCommunicable2000; @worldhealthorganizationMeaslesOutbreakGuide2022; @craggOutbreakResponse2018].
-When diseases have non-specific symptoms, without accurate measurement tools we cannot have confidence in our ability to correctly ascribe changes in symptoms within a population to a particular disease, and therefore detect outbreaks of specific pathogens.
+For diseases with non-specific symptoms, accurate measurement tools are often necessary to confidently and correctly ascribe changes in symptom prevalence within a population to a particular disease, and therefore detect outbreaks of specific pathogens.
 As a result, it has been commonplace for surveillance systems to be developed around high-accuracy tests, such as Polymerase Chain Reaction (PCR) tests and immunoglobulin (Ig) tests, when financially and logistically feasible [@gastanaduyMeasles2019; @commissionerCoronavirusCOVID19Update2020; @grasslyComparisonMolecularTesting2020; @ezhilanSARSCoVMERSCoVSARSCoV22021; @worldhealthorganizationCholera2023; @essentialprogrammeonimmunizationepiimmunizationClinicalSpecimensLaboratory2018].
@@
 Much like the accuracy of an individual test, outbreak detection systems face the same issue regarding the prioritization of sensitive vs specific alerts [@germanSensitivityPredictiveValue2000; @worldhealthorganizationOperationalThresholds2014; @lewisTimelyDetectionMeningococcal2001].
-For many disease systems, particularly in resource constrained environments where the burden of infectious diseases is highest [@gbd2019childandadolescentcommunicablediseasecollaboratorsUnfinishedAgendaCommunicable2023; @roserBurdenDisease2023], cases are counted and if they breach a threshold, be that weekly, monthly, or some combination of the two, an alert is triggered that may launch a further investigation and/or a response [@worldhealthorganizationMeaslesOutbreakGuide2022; @worldhealthorganizationOperationalThresholds2014].
+For many disease systems, particularly in resource constrained environments where the burden of infectious diseases is highest [@gbd2019childandadolescentcommunicablediseasecollaboratorsUnfinishedAgendaCommunicable2023; @roserBurdenDisease2023], cases are counted and if a pre-determined threshold is breached, be that weekly, monthly, or some combination of the two, an alert is triggered that may launch a further investigation and/or a response [@worldhealthorganizationMeaslesOutbreakGuide2022; @worldhealthorganizationOperationalThresholds2014].
 In effect, this discretizes a distinctly continuous phenomenon into a binary measure, outbreak or no outbreak, for decision making purposes.
 When considering reactive vaccination campaigns and non-pharmaceutical based interventions designed to reduce transmission and limit and suppress outbreaks through direct and indirect protection, early action has the potential to avert the most cases, and in some instances, can be the most cost effective as it provides the opportunity to incorporate later information within an “adaptive management” framework [@atkinsAnticipatingFutureLearning2020; @taoLogisticalConstraintsLead; @graisTimeEssenceExploring2008; @ferrariTimeStillEssence2014; @worldhealthorganizationConfirmingInvestigatingManaging2009; @minettiLessonsChallengesMeasles2013].
-While this framing would point towards a sensitive surveillance system being optimal, with the benefits clearly defined in terms of morbidity and mortality avoided, each action comes with both direct and indirect financial and opportunity costs stemming from unnecessary activities that remove resources from future response activities.
-Just as the balance of sensitivity and specificity of a test must be carefully evaluated, so must the balance at the outbreak level.
+While this framing would point towards a sensitive surveillance system being optimal, with the benefits clearly defined in terms of morbidity and mortality avoided, each action comes with both direct and indirect financial and opportunity costs stemming from unnecessary activities that remove resources from future response capabilities.
+Just as the balance of sensitivity and specificity of a test for an individual must be carefully evaluated, so must the balance at the outbreak level.
 
@@
 To evaluate the alert system performance, we develop a set of outbreak definition criteria and surveillance metrics, drawing inspiration from acceptance sampling, ecological surveillance systems, and epidemiological surveillance systems guidelines and reviews [@yemshanovAcceptanceSamplingCosteffective2020; @christensenHerdlevelInterpretationTest2000; @muratoEvaluationSamplingMethods2020; @sternAutomatedOutbreakDetection1999; @calbaSurveillanceSystemsEvaluation2015; @guidelinesworkinggroupUpdatedGuidelinesEvaluating2001].
-Using these metrics we overcome issues encountered by early warning systems that rely on dynamical values such as $R_{\text{effective}}$ in defining outbreaks [@jombartRealtimeMonitoringCOVID192021; @proverbioPerformanceEarlyWarning2022; @stolermanUsingDigitalTraces2023; @brettAnticipatingEpidemicTransitions2018; @salmonMonitoringCountTime2016], for example characterizing the end of an epidemic period is important in a timeseries where multiple outbreaks will occur.
+Using these metrics we overcome issues encountered by early warning systems that rely on dynamical values such as $R_{\text{effective}}$ in defining outbreaks [@jombartRealtimeMonitoringCOVID192021; @proverbioPerformanceEarlyWarning2022; @stolermanUsingDigitalTraces2023; @brettAnticipatingEpidemicTransitions2018; @salmonMonitoringCountTime2016], for example, characterizing the end of an epidemic period is important in a time series where multiple outbreaks will occur.
 
@@
 We constructed a stochastic compartmental non-age structured SEIR model of measles and simulated using a modified Tau-leaping algorithm with a time step of 1 day, utilizing binomial draws to ensure compartment sizes remained positive valued [@chatterjeeBinomialDistributionBased2005; @gillespieApproximateAcceleratedStochastic2001].
-To stochastically reintroduce infections and cause recurring epidemics, commuter-style imports were added that are proportional to the size of the population and $R_{0}$ [@keelingModelingInfectiousDiseases2008], and the transmission parameter ($\beta_{t}$) is sinusoidal with a period of one year.
+To stochastically reintroduce infections and cause recurring epidemics, commuter-style imports were added that are proportional to the size of the population and $R_{0}$ [@keelingModelingInfectiousDiseases2008], and the transmission parameter ($\beta_{t}$) is sinusoidal with a period of one year and 20% seasonal amplitude.
 $R_{0}$ was set to 16, with a latent period of 10 days and infectious period of 8 days.
-The population was initialized with 500_000 individuals with Ghana-like birth and vaccination rates, and the final results were scaled up to the 2022 population size of Ghana (33 million).
+The population was initialized with 500,000 individuals with Ghana-like birth and vaccination rates, and the final results were scaled up to the approximate 2022 population size of Ghana (33 million) [@worldbankGhana].
 The full table of parameters can be found in @tbl-model-parameters.
@@
 
-#figure(
-	table(
+#table(
   columns: 3,
@@
   [Importation rate], table.cell(colspan: 2, align: center, $(1.06*μ*R_0)/(√(N))$),
-		[Population size (N)], table.cell(colspan: 2, align: center, "500_000, scaled to 33M"),
+  [Population size (N)], table.cell(colspan: 2, align: center, "500,000, scaled to 33M"),
   [Initial proportion susceptible], table.cell(colspan: 2, align: center, "0.05"),
@@
   [Initial proportion recovered], table.cell(colspan: 2, align: center, "0.95"),
-	)
 )
@@
 
-To examine the sensitivity of the detection system to background noise, we layered the measles incidence timeseries with one of four noise time series: Poisson-only noise, or dynamical noise with rubella-like parameters that could be in- or out-of-phase or with independent seasonality to the measles dynamics.
-The total annual incidence of noise in each system was approximately 8 times the average annual measles incidence, with the dynamical noise systems also containing low levels of Poisson noise, reflecting Ghana-like dynamics.
-Despite the average incidence of noise in each scenario being the same, the relative proportion of measles to noise varies differently throughout the timeseries, exacerbating the effects of imperfect diagnostic tests that will produce higher rates of false positives and negatives than ELISA-like diagnostics.
-Imperfect tests are also more susceptible to incorrect results because of the varying prevalence throughout the timeseries, so the PPV of the system will not be static throughout the timeseries.
-To account for the stochasticity of the model, we simulated 100 time series for each scenario, each lasting 100 years, before summarizing the distributions of outbreak detection methods.
+To examine the sensitivity of the detection system to background noise, we layered the measles incidence time series with one of four noise time series structures: Poisson-only noise; or dynamical noise with rubella-like parameters that could be in- or out-of-phase, or with independent seasonality to the measles dynamics.
+Only  dynamical in-phase noise and Poisson-only noise are presented in this manuscript.
+For each noise structure, five different magnitudes of noise were simulated and evaluated: the total annual incidence of noise in each system was either the same magnitude as the average annual measles incidence, two times, four times, six times, or eight times the average annual measles incidence.
+100 time series of 100 years were simulated for each scenario, before summarizing the distributions of outbreak detection methods.
 
@@
 It is common to use expert review to define outbreaks when examining empirical data, but this is not feasible in a modeling study where tens of thousands of years are being simulated.
-To account for this, many studies only simulate a single outbreak within a timeseries (repeating this short stochastic simulation multiple times to ensemble results), define an outbreak as a period where $R_{\text{effective}}$ > 1, or use a threshold of > 2 standard deviations (s.d.) over the mean seasonal incidence observed in empirical data [@sternAutomatedOutbreakDetection1999;; @jombartRealtimeMonitoringCOVID192021; @stolermanUsingDigitalTraces2023; @salmonMonitoringCountTime2016; @teklehaimanotAlertThresholdAlgorithms2004; @leclereAutomatedDetectionHospital2017].
-Each method has its uses, but to evaluate the performance of an outbreak detection system in an endemic region where multiple sequential epidemics are expected it is important to clearly define the bounds of the outbreak, which can only be achieved by one of these methods (2 s.d.
-> mean).
+To account for this, many studies only simulate a single outbreak within a time series (repeating this short stochastic simulation multiple times to ensemble results), define an outbreak as a period where $R_{\text{effective}}$ > 1, or use a threshold of > 2 standard deviations (s.d.) over the mean seasonal incidence observed in empirical data (or from a 'burn-in' period of the simulation) [@sternAutomatedOutbreakDetection1999; @jombartRealtimeMonitoringCOVID192021; @stolermanUsingDigitalTraces2023; @salmonMonitoringCountTime2016; @teklehaimanotAlertThresholdAlgorithms2004; @leclereAutomatedDetectionHospital2017].
+Each method has its uses, but to evaluate the performance of an outbreak detection system in an endemic region where multiple sequential epidemics are expected it is important to clearly define the bounds of the outbreak, which can only be achieved by 2 s.d.
+> mean ($R_\text{effective}$ will be less than 1 after an outbreak's peak, but still within what can be reasonably defined as the outbreak's bounds).
 This, however, assumes strong seasonal forcing and regular periodicity of incidence to produce a smooth enough baseline, which is not present as countries near measles elimination status [@grahamMeaslesCanonicalPath2019].
-Here we define a true measles outbreak as a region of the timeseries that meets the following three criteria:
+Here we define a true measles outbreak as a region of the time series that meets the following three criteria:
 
@@
 For example, a region may produce 1000 total cases within the lower and upper bounds defined by the daily incidence being $\ge$ 5 cases, but if this period only lasts for 20 days it would not be considered an outbreak as it would not be possible to mount a response to.
-The incidence of non-measles febrile rash (i.e., noise) does not affect the outbreak status of a region but does affect the alert status triggered by the testing protocol.
+The incidence of non-measles febrile rash (i.e., noise) does not affect the outbreak status of a region but may affect the alert status triggered by the testing protocol.
 
@@
 The use of alert thresholds is common in locations burdened by measles, where access to high performance computing is limited making nowcasting-style approaches to outbreak detection impractical.
-For this reason, we define an “alert” as any single day where the 7-day moving average of the daily incidence is greater than, or equal to, a threshold of X.
-To examine the interaction of test and surveillance systems we vary the alert threshold, X, between 1 and 15 cases per day, before computing the evaluation metrics and identifying the optimal threshold for each individual test.
-We also examined an alert method that required either the daily incidence or the 7-day moving average of daily incidence to exceed the alert threshold X (see supplemental results).
+For this reason, we define an “alert” as any single day where the 7-day moving average of the daily incidence is greater than, or equal to, a pre-specified alert threshold.
+To examine the interaction of test and surveillance systems we vary the alert threshold, between 1 and 15 cases per day, before computing the evaluation metrics and identifying the optimal threshold for each combination of individual test, testing rate, noise structure, and noise magnitude.
+We also examined an alert method that required either the daily incidence or the 7-day moving average of daily incidence to exceed the alert threshold (see supplemental results).
 
-Each day, 60% of the measles and non-measles febrile rash cases visit the clinic for treatment, and Y% percentage of these clinic visits are tested, as all clinic visits are deemed to be suspected measles cases because they meet the clinical case definition.
-This percentage of clinic visits, Y, that are tested is varied between 10% and 60%, in 10% increments, for all combinations of diagnostic test (apart from the clinical case definition) and alert threshold, defining the “testing scenario”.
-While a reporting rate of 6-48% is high, we are only simulating a population of 500_000 individuals i.e., approximately 1.5% of Ghana’s population, which produces absolute incidence values that are consistent with observed reporting on the national scale.
+Each day, 60% of the measles and non-measles febrile rash cases visit the clinic for treatment, and a percentage of these clinic visits are tested, as all clinic visits are deemed to be suspected measles cases because they meet the clinical case definition.
+This percentage of clinic visits that are tested is varied between 10% and 60%, in 10% increments, for all combinations of diagnostic test (except clinical case definition) and alert threshold, defining the “testing scenario”.
+<!--While a reporting rate of 6-48% is high, we are only simulating a population of 500,000 individuals i.e., approximately 1.5% of Ghana’s population, which produces absolute incidence values that are consistent with observed reporting on the national scale.-->
 Each testing scenario uses one of the following tests:
@@
 This acts as a lower bound of acceptability for a new measles RDT.
+- An RDT equivalent with 90% sensitivity and specificity, and 0-day lag in result return.
 - An ELISA equivalent test with 100% sensitivity and specificity with a 0-day test result delay.
@@
 
-For each of the previously defined tests, we calculate the number of test positive cases (that will include false positive and negative cases resulting from imperfect diagnostic tests), which is used to categorize the time series by alert status in conjunction with the alert threshold X.
+For each of the previously defined tests, we calculate the number of test positive cases (that will include false positive and negative cases resulting from imperfect diagnostic tests), which is used to categorize the time series by alert status in conjunction with the alert threshold.
 The overlap with true outbreak status, or lack thereof, is used to classify the following metrics, and has been illustrated in @fig-outbreak-schematic:
 
-- The percentage of alerts that are correct.
+- The percentage of alerts that are 'correct'.
 This is analogous to the PPV of the alert system.
@@
 - The detection delay.
-Note that if an alert precedes the start of the outbreak, as long as it successfully “captures” the outbreak within its bounds, it is considered to be correct, resulting in a negative detection delay i.e., an early warning triggered by false positives.
+If an alert precedes the start of the outbreak, as long as it successfully “captures” the outbreak within its bounds, it is considered to be a 'correct' alert, resulting in a negative detection delay i.e., an early warning triggered by false positives.
 - The number of unavoidable and avoidable outbreak cases and deaths.
@@
 The threshold that maximized outbreak detection accuracy given a testing scenario depends heavily on the noise structure.
-For example, the optimal threshold for an RDT with 90% sensitivity/specificity and 40% testing of clinic visits was 4 cases per day with the dynamical in-phase noise, whereas it is 5 cases per day when Poisson-only noise was simulated (@tbl-optimal-thresholds).
+For example, when the ratio of average noise:measles incidence is 8, the optimal threshold for an RDT with 90% sensitivity/specificity and 40% testing of clinic visits was 4 cases per day with the dynamical in-phase noise, whereas it is 5 cases per day when Poisson-only noise was simulated (@tbl-optimal-thresholds).
 This corresponds to an accuracy of 72% and 93%, respectively (@tbl-optimal-thresholds-accuracy).
@@
 
-#figure(
-	table(
+#table(
   columns: 10,
@@
   ..optimal_thresholds.flatten()
-	)
 )
@@
 Despite the mean daily noise incidence set to equivalent values between the dynamical and Poisson-only simulations, drastically different results are observed.
+
+> Despite the same average incidence of noise in each (comparable) scenario, the relative proportion of measles to noise varies differently throughout the time series, exacerbating the effects of imperfect diagnostic tests that will produce higher rates of false positives and negatives than ELISA-like diagnostics.
+Imperfect tests are also more susceptible to incorrect results because of the varying prevalence throughout the time series, so the PPV of the system will not be static throughout the time series.
+
 Under the assumption that non-measles febrile rash is relatively static in time (Poisson), RDTs can perform as well, if not better than ELISA tests at moderate to high testing rates, and at a fraction of the cost.
@@
 For computational simplicity, this paper did not include demography in the model structure.
-And while a simulation-based approach allows for complete determination of true infection status i.e., measle vs non-measles febrile rash cases, and therefore an accurate accounting of the outbreak and alert bounds, these simulations do not specifically represent any real-world setting.
+And while a simulation-based approach allows for complete determination of true infection status i.e., measles vs non-measles febrile rash cases, and therefore an accurate accounting of the outbreak and alert bounds, these simulations do not specifically represent any real-world setting.
 The evaluation of empirical data does provide this opportunity, but at the cost of knowing the true infection status of individuals and confounding of multiple variables, limiting analysis to only those who are observed (i.e., not those in the community who do not visit a healthcare center), and removing the possibility to explore the sensitivity of the results to parameters of interest to a surveillance program e.g., testing rate, and the test itself.
@@
 Given these limitations, the explicit values (i.e., optimal thresholds, accuracies etc.) should be interpreted with caution, and the exact results observed in the real-world will likely be highly dependent on unseen factors, such as the proportion of measles and non-measles sources of febrile rash that seek healthcare.
-However, the general patterns should hold, and more importantly, the analysis framework provides a consistent and holistic approach to evaluating the tradeoff between individual level tests and the alert system enacted to detect outbreaks.
+However, the general patterns should hold, and more importantly, the analysis framework provides a consistent and holistic approach to evaluating the trade-off between individual level tests and the alert system enacted to detect outbreaks.
 

### 2024-09-28 manuscript/manuscript.qmd
Update lineplot code to work with other outcomes and create tables and plots in the manuscript
Use separate files that are run with each manuscript rendering to ensure
correct versions are included

@@
 
+```{julia}
+#| echo: false
+include("optimal-thresholds.jl");
+```
+
 ::: {#tbl-optimal-thresholds}
@@
 ```{=typst}
-#let optimal_thresholds = csv("optimal-thresholds.csv")
+#let optimal_thresholds = csv("thresholds.csv")
 
@@
   },
+  align: center,
   [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
@@
 ```{=typst}
-#let optimal_thresholds = csv("optimal-thresholds-accuracy.csv")
+#let accuracy = csv("optimal-thresholds_accuracy.csv")
 
@@
   [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
-  ..optimal_thresholds.flatten()
+  ..accuracy.flatten()
 )
@@
 
-Optimal threshold for each testing scenario. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only
+The outbreak detection accuracy for each testing scenario, at the optimal thresholds.
 
@@
 ```{julia}
+#| echo: false
+include("line-plot.jl");
+```
+
+```{julia}
 #| fig-cap: "The accuracy of outbreak detection systems under different testing rates and noise structures.
@@
 #| label: fig-accuracy
-include("line-plot.jl")
+accuracy_line_plot
+```
+
+```{julia}
+#| fig-cap: "The number of unavoidable cases of outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| label: fig-unavoidable
+unavoidable_line_plot
+```
+
+```{julia}
+#| fig-cap: "The detection delay of outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| label: fig-delay
+delay_line_plot
 ```

### 2024-09-28 OutbreakDetectionUtils/src/structs.jl

@@ -354,6 +354,10 @@ struct PoissonNoiseSpecification{
     noise_mean_scaling::T2
 end
 
+function PoissonNoiseSpecification(noise_mean_scaling)
+    return PoissonNoiseSpecification("Poisson", noise_mean_scaling)
+end
+
 struct DynamicalNoiseSpecification{
     T1<:AbstractString,T2<:AbstractFloat,T3<:Integer
 } <: NoiseSpecification

### 2024-09-28 manuscript/line-plot.jl

@@ -80,18 +80,57 @@ baseplotdirpath = joinpath(
     basedirpath,
 )
 
+
 #%%
-line_accuracy_plot(
+accuracy_line_plot = line_accuracy_plot(
     ensemble_noise_specification,
     ensemble_percent_clinic_tested_vec,
     optimal_threshold_test_spec_vec,
     optimal_threshold_core_params;
+    outcome = :accuracy,
+    ylabel = "Outbreak Detection\nAccuracy",
     plotdirpath = baseplotdirpath,
     facet_fontsize = 20,
     labelsize = 20,
     show_x_facet_label = true,
     show_y_facet_label = false,
-    clinical_hline = true,
+    ylims = (0.6, 1.0),
     force = true,
     save_plot = false,
 )
+
+#%%
+unavoidable_line_plot = line_accuracy_plot(
+    ensemble_noise_specification,
+    ensemble_percent_clinic_tested_vec,
+    optimal_threshold_test_spec_vec,
+    optimal_threshold_core_params;
+    outcome = :unavoidable_cases,
+    ylabel = "Unavoidable Cases",
+    plotdirpath = baseplotdirpath,
+    facet_fontsize = 20,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    force = true,
+    save_plot = false,
+)
+
+#%%
+delay_line_plot = line_accuracy_plot(
+    ensemble_noise_specification,
+    ensemble_percent_clinic_tested_vec,
+    optimal_threshold_test_spec_vec,
+    optimal_threshold_core_params;
+    outcome = :detectiondelays,
+    ylabel = "Detection Delays",
+    plotdirpath = baseplotdirpath,
+    clinical_hline = false,
+    facet_fontsize = 20,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    force = true,
+    save_plot = false,
+)
+

### 2024-09-28 manuscript/optimal-thresholds.jl

@@ -1,0 +1,317 @@
+#%%
+using DrWatson
+@quickactivate "OutbreakDetection"
+
+using ProgressMeter
+using FLoops
+using NaNMath: NaNMath
+using CSV: CSV
+using DataFrames
+using CategoricalArrays
+using DataFramesMeta
+using Statistics
+using Match
+
+using OutbreakDetectionUtils
+
+includet(srcdir("makie-plotting-setup.jl"))
+include(srcdir("ensemble-parameters.jl"))
+
+#%%
+rdt_spec_vec = [
+    IndividualTestSpecification(0.85, 0.85, 0),
+    IndividualTestSpecification(0.9, 0.9, 0),
+    IndividualTestSpecification(1.0, 0.0, 0),
+]
+
+elisa_test_spec_vec = [
+    IndividualTestSpecification(1.0, 1.0, 0),
+    IndividualTestSpecification(1.0, 1.0, 3),
+    IndividualTestSpecification(1.0, 1.0, 7),
+    IndividualTestSpecification(1.0, 1.0, 14),
+]
+
+optimal_threshold_alertthreshold_vec = collect(1:1:15)
+
+R_0_vec = collect(16.0)
+
+ensemble_dynamics_spec_vec = create_combinations_vec(
+    DynamicsParameters,
+    (
+        [ensemble_state_specification.init_states.N],
+        [27],
+        [0.2],
+        [SIGMA],
+        [GAMMA],
+        R_0_vec,
+        [0.8],
+    ),
+)
+
+ensemble_spec_vec = create_combinations_vec(
+    EnsembleSpecification,
+    (
+        [ensemble_model_type],
+        [ensemble_state_specification],
+        ensemble_dynamics_spec_vec,
+        [ensemble_time_specification],
+        [ensemble_nsims],
+    ),
+)
+
+# alert_method_vec = ["movingavg", "dailythreshold_movingavg"]
+alert_method_vec = ["movingavg"]
+
+#%%
+cfr_df = CSV.read(
+    datadir("CFR_2022.csv"),
+    DataFrame; delim = ',',
+    header = true,
+    types = [String, Int64, Float64],
+    silencewarnings = true,
+)
+dropmissing!(cfr_df)
+
+gha_cfr = only(cfr_df[cfr_df.country .== "GHA", :CFR])
+
+population_df = CSV.read(
+    datadir("input-populations.csv"),
+    DataFrame; delim = ',',
+    header = true,
+)
+
[…]

### 2024-09-28 src/line_plots.jl

@@ -13,6 +13,7 @@ function line_accuracy_plot(
     ensemble_percent_clinic_tested_vec,
     optimal_threshold_test_spec_vec,
     optimal_threshold_core_params;
+    outcome = :accuracy,
     plotdirpath = DrWatson.plotsdir(),
     plotname = "line_accuracy_plot",
     plotformat = "png",
@@ -24,6 +25,7 @@ function line_accuracy_plot(
     labelsize = 30,
     show_x_facet_label = true,
     show_y_facet_label = true,
+    ylims = (nothing, nothing),
     clinical_hline = true,
     force = false,
     save_plot = true,
@@ -72,10 +74,12 @@ function line_accuracy_plot(
                     optimal_thresholds_vec,
                     i,
                     j;
+                    outcome = outcome,
                     num_noise_descriptions = num_noise_descriptions,
                     colors = colors,
                     xlabel = xlabel,
                     ylabel = ylabel,
+                    ylims = ylims,
                     facet_fontsize = facet_fontsize,
                     show_x_facet_label = show_x_facet_label,
                     kwargs...,
@@ -98,7 +102,9 @@ function line_accuracy_plot(
             end
             colsize!(fig.layout, 0, Relative(0.03))
         end
-        unique_test_specifications = unique_test_specifications[1:(end - 1)]
+        if clinical_hline
+            push!(colors, "green")
+        end
         Legend(
             fig[0, :],
             map(enumerate(unique_test_specifications)) do (i, test_spec)
@@ -130,19 +136,21 @@ function _line_accuracy_plot!(
     optimal_thresholds_vec,
     i,
     j;
+    outcome = :accuracy,
     colors = lineplot_colors,
     xlabel = "Proportion Tested",
     ylabel = "Accuracy",
     show_x_facet_label = true,
     facet_fontsize = 24,
     num_noise_descriptions = 1,
+    ylims = (nothing, nothing),
     kwargs...,
 )
     kwargs_dict = Dict{Symbol,Any}(kwargs)
 
     long_df = create_optimal_threshold_summary_df(
         optimal_thresholds_vec,
-        :accuracy;
+        outcome;
         percentiles = [0.1, 0.9],
     )
 
@@ -153,7 +161,7 @@ function _line_accuracy_plot!(
             :sensitivity,
             :specificity,
             :test_lag,
-            :accuracy_mean,
[…]

### 2024-10-20 OutbreakDetectionUtils/src/diag-testing-functions.jl
Update OutbreakThresholdChars struct to include proportion of time series in alert/outbreak

@@ -310,11 +310,23 @@ function calculate_OutbreakThresholdChars(
         )
 
         n_outbreak_tests = calculate_n_outbreak_tests(
-            @view(testarr[:, 1, sim]), @view(testarr[:, 2, sim]), outbreakbounds
+            @view(testarr[:, 1, sim]), @view(testarr[:, 2, sim]),
+            outbreakbounds,
         )
 
         mean_noise_incidence_ratio =
             noise_means.mean_noise / StatsBase.mean(@view(infecarr[:, 1, sim]))
+
+        proportion_timeseries_in_outbreak = calculate_proportion_timeseries_in_outbreak(
+            @view(infecarr[:, 3, sim])
+        )
+
+        proportion_timeseries_in_alert = calculate_proportion_timeseries_in_outbreak(
+            @view(testarr[:, 6, sim])
+        )
+
+        alert_outbreak_timeseries_prop_diff =
+            proportion_timeseries_in_alert - proportion_timeseries_in_outbreak
 
         OutbreakThresholdChars(
             dailychars...,
@@ -332,17 +344,32 @@ function calculate_OutbreakThresholdChars(
             mean_noise_incidence_ratio,
             noise_means.mean_poisson_noise,
             noise_means.poisson_noise_prop,
+            proportion_timeseries_in_outbreak,
+            proportion_timeseries_in_alert,
+            alert_outbreak_timeseries_prop_diff,
         )
     end
 
     return StructArrays.StructArray(OT_chars)
 end
 
+function calculate_proportion_timeseries_in_outbreak(outbreak_status_vec)
+    return sum(outbreak_status_vec) / length(outbreak_status_vec)
+end
+
+function calculate_proportion_timeseries_in_outbreak(
+    outbreak_bounds_periodsum_vec, time_parameters
+)
+    return sum(outbreak_bounds_periodsum_vec) / time_parameters.tlength
+end
+
 function calculate_n_tests(infectious_tested_vec, noise_tested_vec)
     return sum(infectious_tested_vec) + sum(noise_tested_vec)
 end
 
-function calculate_n_outbreak_tests(infectious_tested_vec, noise_tested_vec, outbreakbounds)
+function calculate_n_outbreak_tests(
+    infectious_tested_vec, noise_tested_vec, outbreakbounds
+)
     tested = 0
     for (lower, upper) in eachrow(@view(outbreakbounds[:, 1:2]))
         @views tested += sum(infectious_tested_vec[lower:upper])

### 2024-10-20 OutbreakDetectionUtils/src/structs.jl

@@ -239,6 +239,9 @@ struct OutbreakThresholdChars{
     mean_noise_incidence_ratio::T1
     mean_poisson_noise::T1
     poisson_noise_prop::T1
+    proportion_timeseries_in_outbreak::T1
+    proportion_timeseries_in_alert::T1
+    alert_outbreak_timeseries_prop_diff::T1
 end
 
 struct OutbreakSpecification{T1<:Integer,T2<:AbstractString}

### 2024-10-20 manuscript/manuscript.qmd
Update manuscript plots

@@
 
+```{julia}
+#| fig-cap: "The difference in the proportion of the time series in alert - proportion in an outbreak,of outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| label: fig-delay
+alert_outbreak_proportion_line_plot
+```
+
 # Discussion

### 2024-10-20 manuscript/line-plot.jl

@@ -80,7 +80,6 @@ baseplotdirpath = joinpath(
     basedirpath,
 )
 
-
 #%%
 accuracy_line_plot = line_accuracy_plot(
     ensemble_noise_specification,
@@ -90,7 +89,7 @@ accuracy_line_plot = line_accuracy_plot(
     outcome = :accuracy,
     ylabel = "Outbreak Detection\nAccuracy",
     plotdirpath = baseplotdirpath,
-    facet_fontsize = 20,
+    facet_fontsize = 18,
     labelsize = 20,
     show_x_facet_label = true,
     show_y_facet_label = false,
@@ -108,10 +107,12 @@ unavoidable_line_plot = line_accuracy_plot(
     outcome = :unavoidable_cases,
     ylabel = "Unavoidable Cases",
     plotdirpath = baseplotdirpath,
-    facet_fontsize = 20,
+    clinical_hline = false,
+    facet_fontsize = 18,
     labelsize = 20,
     show_x_facet_label = true,
     show_y_facet_label = false,
+    ylims = (0, 3.5e4),
     force = true,
     save_plot = false,
 )
@@ -126,11 +127,30 @@ delay_line_plot = line_accuracy_plot(
     ylabel = "Detection Delays\n(Days)",
     plotdirpath = baseplotdirpath,
     clinical_hline = false,
-    facet_fontsize = 20,
+    facet_fontsize = 18,
     labelsize = 20,
     show_x_facet_label = true,
     show_y_facet_label = false,
+    ylims = (-100, 100),
     force = true,
     save_plot = false,
 )
 
+#%%
+alert_outbreak_proportion_line_plot = line_accuracy_plot(
+    ensemble_noise_specification,
+    ensemble_percent_clinic_tested_vec,
+    optimal_threshold_test_spec_vec,
+    optimal_threshold_core_params;
+    outcome = :alert_outbreak_timeseries_prop_diff,
+    ylabel = "Proportion of Time Series\nIn Alert - Outbreak",
+    plotdirpath = baseplotdirpath,
+    clinical_hline = false,
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (-100, 100),
+    force = true,
+    save_plot = false,
+)

### 2024-10-20 src/line_plots.jl

@@ -27,6 +27,7 @@ function line_accuracy_plot(
     show_x_facet_label = true,
     show_y_facet_label = true,
     ylims = (nothing, nothing),
+    hidedecorations = (true, true),
     clinical_hline = true,
     force = false,
     save_plot = true,
@@ -51,11 +52,10 @@ function line_accuracy_plot(
 
         for (i, noise_description) in pairs(unique_noise_descriptions)
 
-
-
             label_noise_description = Match.@match noise_description begin
                 "poisson" => "Poisson Noise"
                 "dynamical, in-phase" => "Dynamical Noise: In-Phase"
+                _ => "Other Noise"
             end
 
             shape_noise_specification = filter(
@@ -89,6 +89,7 @@ function line_accuracy_plot(
                     xlabel = xlabel,
                     ylabel = "$label_noise_description\n" * ylabel,
                     ylims = ylims,
+                    hidedecorations = hidedecorations,
                     facet_fontsize = facet_fontsize,
                     show_x_facet_label = show_x_facet_label,
                     kwargs...,
@@ -160,6 +161,7 @@ function _line_accuracy_plot!(
     facet_fontsize = 24,
     num_noise_descriptions = 1,
     ylims = (nothing, nothing),
+    hidedecorations = (true, true),
     kwargs...,
 )
     kwargs_dict = Dict{Symbol,Any}(kwargs)
@@ -203,6 +205,7 @@ function _line_accuracy_plot!(
         ylabel = ""
     end
 
+
     _line_accuracy_facet!(
         gl,
         noise_spec,
@@ -216,6 +219,14 @@ function _line_accuracy_plot!(
         ylims = ylims,
         kwargs_dict...,
     )
+
+    if ylims != (nothing, nothing) && hidedecorations[2] && j != 1
+        hideydecorations!(contents(gl)[1])
+    end
+
+    if hidedecorations[1] && i < num_noise_descriptions
+        hidexdecorations!(contents(gl)[1])
+    end
 
     return nothing
 end

### 2024-10-20 manuscript/manuscript.qmd
Update manuscript to show outbreak proportion figure

@@
 The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
-#| label: fig-delay
+#| label: fig-outbreak-proportion
+outbreak_proportion_line_plot
+```
+
+```{julia}
+#| fig-cap: "The difference in the proportion of the time series in alert - proportion in an outbreak,of outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| label: fig-alert-proportion
+alert_proportion_line_plot
+```
+
+```{julia}
+#| fig-cap: "The difference in the proportion of the time series in alert - proportion in an outbreak,of outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| label: fig-alert-outbreak-proportion
 alert_outbreak_proportion_line_plot

### 2024-10-21 manuscript/manuscript.qmd
Move tables out of manuscript into supplemental appendix

@@
 
-::: {#tbl-optimal-thresholds-accuracy}
-
-```{=typst}
-#let accuracy = csv("optimal-thresholds_accuracy.csv")
-
-#table(
-  columns: 10,
-    fill: (x, y) => {
-      if y == 0 {gray}
-      if y == 1 {gray}
-    },
-  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
-  ..accuracy.flatten()
-)
-```
-
-The outbreak detection accuracy for each testing scenario, at the optimal thresholds.
-
-:::
-
-::: {#tbl-optimal-thresholds-unavoidable}
-
-```{=typst}
-#let optimal_thresholds = csv("optimal-thresholds-unavoidable.csv")
-
-#table(
-  columns: 10,
-    fill: (x, y) => {
-      if y == 0 {gray}
-      if y == 1 {gray}
-    },
-  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
-  ..optimal_thresholds.flatten()
-)
-```
-
-Unavoidable cases per annum of each testing scenario at their specific optimal thresholds, scaled up to Ghana’s 2022 population.
-A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence.
-B) the noise structure is Poisson only.
-
-:::
-
-::: {#tbl-optimal-thresholds-delays}
-
-```{=typst}
-#let optimal_thresholds = csv("optimal-thresholds-delays.csv")
-
-#table(
-  columns: 10,
-    fill: (x, y) => {
-      if y == 0 {gray}
-      if y == 1 {gray}
-    },
-  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
-  ..optimal_thresholds.flatten()
-)
-```
-
-Outbreak alert delay (days) of each testing scenario at their specific optimal thresholds.
-A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence.
-B) the noise structure is Poisson only.
-
-:::
-
 ```{julia}

### 2024-10-21 manuscript/supplemental-appendix.qmd
Move tables out of manuscript into supplemental appendix

@@
 # Results
+
+## Tables
+
+::: {#tbl-optimal-thresholds-accuracy}
+
+```{=typst}
+#let accuracy = csv("optimal-thresholds_accuracy.csv")
+
+#table(
+  columns: 10,
+    fill: (x, y) => {
+      if y == 0 {gray}
+      if y == 1 {gray}
+    },
+  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
+  ..accuracy.flatten()
+)
+```
+
+The outbreak detection accuracy for each testing scenario, at the optimal thresholds.
+
+:::
+
+::: {#tbl-optimal-thresholds-unavoidable}
+
+```{=typst}
+#let optimal_thresholds = csv("optimal-thresholds-unavoidable.csv")
+
+#table(
+  columns: 10,
+    fill: (x, y) => {
+      if y == 0 {gray}
+      if y == 1 {gray}
+    },
+  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
+  ..optimal_thresholds.flatten()
+)
+```
+
+Unavoidable cases per annum of each testing scenario at their specific optimal thresholds, scaled up to Ghana’s 2022 population.
+A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence.
+B) the noise structure is Poisson only.
+
+:::
+
+::: {#tbl-optimal-thresholds-delays}
+
+```{=typst}
+#let optimal_thresholds = csv("optimal-thresholds-delays.csv")
+
+#table(
+  columns: 10,
+    fill: (x, y) => {
+      if y == 0 {gray}
+      if y == 1 {gray}
+    },
+  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
+  ..optimal_thresholds.flatten()
+)
+```
+
+Outbreak alert delay (days) of each testing scenario at their specific optimal thresholds.
+A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence.
+B) the noise structure is Poisson only.
+
+:::
+
+## Figures
+
+```{julia}
+#| echo: false
+include("line-plots-script.jl");
+```
+
+```{julia}
+#| fig-cap: "The accuracy of outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| label: fig-accuracy
+accuracy_line_plot
+```
+
+```{julia}
+#| fig-cap: "The number of unavoidable cases of outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| label: fig-unavoidable
+unavoidable_line_plot
+```
+
+```{julia}
+#| fig-cap: "The detection delay of outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| label: fig-delay
+delay_line_plot
+```
+
+```{julia}
+#| fig-cap: "The difference between proportion of the time series in alert for outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| label: fig-alert-proportion
+alert_proportion_line_plot
+```

### 2024-10-21 manuscript/manuscript.qmd
Update manuscript text and use jupyter cache

@@
 execute:
+  cache: true
   echo: false
@@
 The full table of parameters can be found in @tbl-model-parameters.
-All simulations and analysis was completed in Julia version 1.10.3 [@bezansonJuliaFreshApproach2017], with all code stored at [https://github.com/arnold-c/OutbreakDetection](https://github.com/arnold-c/OutbreakDetection).
+All simulations and analysis was completed in Julia version 1.10.5 [@bezansonJuliaFreshApproach2017], with all code stored at [https://github.com/arnold-c/OutbreakDetection](https://github.com/arnold-c/OutbreakDetection).
 
@@
 
-## Optimal Thresholds, Accuracy, and Unavoidable Cases
-
-The threshold that maximized outbreak detection accuracy given a testing scenario depends heavily on the noise structure.
-For example, when the ratio of average noise:measles incidence is 8, the optimal threshold for an RDT with 90% sensitivity/specificity and 40% testing of clinic visits was 4 cases per day with the dynamical in-phase noise, whereas it is 5 cases per day when Poisson-only noise was simulated (@tbl-optimal-thresholds).
-This corresponds to an accuracy of 72% and 93%, respectively (@tbl-optimal-thresholds-accuracy).
-Much higher accuracy can be achieved with imperfect tests under Poisson noise because large spikes of non-measles febrile rash do not occur that would lead to many false positives, triggering an erroneous alert.
-With dynamical noise, this possibility can and does occur, and when the ratio of measles to non-measles cases is sufficiently low, drastically impacts the accuracy of an alert system.
-This is why the values observed among ELISA tests are identical between the different noise structures: with a perfect test there are no false positive results.
-For dynamical noise, RDTs never perform as well as an ELISA, even scenarios that include a 14-day lag between test procurement and result receipt.
-However, with Poisson noise, once testing reaches 20% of clinic visits, equivalence can be observed at approximately 91% accuracy (@tbl-optimal-thresholds-accuracy).
+The threshold that maximized outbreak detection accuracy given a testing scenario depends heavily on the noise structure and magnitude.
+For example, when the ratio of average noise incidence was 8 times higher than the average measles incidence, the optimal threshold for an RDT with 90% sensitivity/specificity and 40% testing of clinic visits was 4 cases per day with the dynamical in-phase noise, whereas it is 5 cases per day when Poisson-only noise was simulated (@tbl-optimal-thresholds).
+This corresponds to an accuracy of 72% and 93%, respectively (@fig-accuracy).
+Much higher accuracy can be achieved with imperfect tests under Poisson noise because large spikes of non-measles febrile rash that would lead to many false positives and trigger erroneous alerts do not occur.
+With dynamical noise, this possibility can and does occur and, when the ratio of non-measles to measles cases is sufficiently high, drastically impacts the accuracy of an alert system.
+The results from ELISA tests are identical between the different noise structures: with a perfect test there are no false positive results.
+For dynamical noise, RDTs never perform as well as an ELISA, even under scenarios that include a 14-day lag between test procurement and result receipt.
+However, with Poisson noise, once testing reaches 20% of clinic visits, equivalence can be observed at approximately 91% accuracy (@fig-accuracy).
 
@@
 
-Optimal threshold for each testing scenario. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only
+Optimal threshold for each testing scenario, when the average noise incidence is 8 times higher than the average measles incidence.
+A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence.
+B) the noise structure is Poisson only
 
@@
 However, the changes in accuracy are not monotonically increasing, as one might anticipate.
-Even with a perfect test, there are sequential cells that increase the testing rate, yet accuracy decreases.
-For example, the ELISA equivalent with a 0-day test result lag goes from an outbreak detection accuracy of 91% down to 87% when testing increases from 30% to 40% of clinic visits.
+Even with a perfect test, there are sequential cells that increase the testing rate, yet accuracy decreases (@fig-accuracy).
+For example, the ELISA equivalent with a 0-day test result lag outbreak detection accuracy decreases from 91% to 87% when clinic testing rates increase from 30% to 40% of clinic visits.
 The reason behind this unintuitive result stems from the use of integer valued thresholds.
-While computing the moving average incidence helps to smooth out the alert system, requiring the daily alert to be an integer value results in the optimal threshold for alert accuracy being stuck at a lower level than a truly optimal intermediate floating point number, because the increase in the alert system’s PPV from a higher threshold is outweighed by the larger decrease in alert sensitivity, or the threshold is forced to increase.
-Examining the number of unavoidable cases (@tbl-optimal-thresholds-unavoidable) and the detection delays (@tbl-optimal-thresholds-delays), we can see that the decrease in accuracy results from a decrease in the sensitivity of the alert system: the detection delay increases by 6 days, and the unavoidable cases by c.
+While computing the moving average test positive incidence helps to smooth out the alert system, an integer valued threshold can result in step-changes of accuracy between two threshold values, and the expected increase in the alert system's PPV from a higher threshold value is outweighed by the loss in the alert system's sensitivity to detecting outbreaks.
+Even with a perfect test, the alert system needs to discriminate between endemic/imported cases and epidemic cases; increasing the testing rate will result in higher numbers of test positive individuals, and a lower threshold can result in an overly sensitive alert system, triggering for measles infections, but not those within an outbreak.
+A higher threshold will face the opposite issue; not triggering for smaller outbreaks, which is more likely to be an issue at lower testing rates.
+Examining the number of unavoidable cases (@fig-unavoidable) and the detection delays (@fig-delay), we can see that the decrease in accuracy results from a decrease in the sensitivity of the alert system: the detection delay increases by 6 days, and the unavoidable cases by c.
 2300.
 Although the testing rate increases, so does the optimal threshold, from 3 test positives to 4 test positives per day.
+The subsequent change in testing rate (40 - 50%) is associated with no change in the optimal threshold for the ELISA, and as expected, the number of unavoidable cases and detection delay decrease, resulting in a more sensitive alert system, with an higher system accuracy indicating that the increase more than offset the decrease in the system's PPV.
 
@@
 #| echo: false
-include("line-plot.jl");
+include("line-plots-script.jl");
 ```
@@
 
-```{julia}
-#| fig-cap: "The difference in the proportion of the time series in alert - proportion in an outbreak,of outbreak detection systems under different testing rates and noise structures.
-The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
-#| label: fig-outbreak-proportion
-outbreak_proportion_line_plot
-```
+For Poisson noise, similar detection delays are observed for all test and noise magnitudes, with variation by testing rate (mean of -3.7 to 36.1 days).
+Under dynamical noise, there are clearer differences in the performance of ELISA and RDTs, with the separation of outcomes occurring later than observed for detection accuracy (8 time noise magnitude vs.
+2 times, respectively) (@fig-accuracy).
+With large amounts of dynamical noise (8 times the measles incidence), the mean detection delay of the 90% and 85% RDTs range from -17.5 days to 3.2 days, and from -25.2 days to -3.4 days, respectively.
+Negative delays indicate that alerts are being triggered before the start of the outbreak and is correlated with the proportion of the time series that is under alert, with larger negative delays associated with more and/or longer alert periods (@fig-alert-proportion).
 
 ```{julia}
-#| fig-cap: "The difference in the proportion of the time series in alert - proportion in an outbreak,of outbreak detection systems under different testing rates and noise structures.
+#| fig-cap: "The difference between proportion of the time series in alert for outbreak detection systems under different testing rates and noise structures.
 The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
@@
 alert_proportion_line_plot
-```
-
-```{julia}
-#| fig-cap: "The difference in the proportion of the time series in alert - proportion in an outbreak,of outbreak detection systems under different testing rates and noise structures.
-The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
-#| label: fig-alert-outbreak-proportion
-alert_outbreak_proportion_line_plot
 ```

### 2024-10-22 manuscript/line-plots-script.jl
Rename optimal thresholds scripts for manuscript and remove duplicated optimal threshold calculation
Just load a filter a subset of all optimal thresholds for the creation
of the tables in the supplement

@@ -2,155 +2,19 @@
 using DrWatson
 @quickactivate "OutbreakDetection"
 
-using OutbreakDetectionUtils
-using OutbreakDetection
-using Revise
+using OutbreakDetectionUtils:
+    IndividualTestSpecification, DynamicalNoiseSpecification
+using StatsBase: mean
 
 include(srcdir("makie-plotting-setup.jl"))
-include(srcdir("ensemble-parameters.jl"))
+
+include(projectdir("manuscript", "optimal-thresholds-loading.jl"));
 
 if false
-    include("../src/ensemble-parameters.jl")
+    include("optimal-thresholds-loading.jl")
 end
 
 #%%
-optimal_threshold_test_spec_vec = [
-    IndividualTestSpecification(0.85, 0.85, 0),
-    IndividualTestSpecification(0.9, 0.9, 0),
-    IndividualTestSpecification(1.0, 1.0, 0),
-    IndividualTestSpecification(1.0, 1.0, 14),
-]
-
-optimal_threshold_alertthreshold_vec = collect(1:1:15)
-
-R_0_vec = [16.0]
-
-ensemble_dynamics_spec_vec = create_combinations_vec(
-    DynamicsParameters,
-    (
-        [ensemble_state_specification.init_states.N],
-        [27],
-        [0.2],
-        [SIGMA],
-        [GAMMA],
-        R_0_vec,
-        [0.8],
-    ),
-)
-
-ensemble_spec_vec = create_combinations_vec(
-    EnsembleSpecification,
-    (
-        [ensemble_model_type],
-        [ensemble_state_specification],
-        ensemble_dynamics_spec_vec,
-        [ensemble_time_specification],
-        [ensemble_nsims],
-    ),
-)
-
-alert_method_vec = ["movingavg"]
-
-#%%
-ensemble_noise_specification =
-    filter(ensemble_noise_specification_vec) do noise_spec
-        noise_spec.noise_type == "poisson" ||
-            noise_spec.correlation == "in-phase"
-    end
-
-ensemble_specification = ensemble_spec_vec[1]
-alertmethod = alert_method_vec[1]
-
-optimal_threshold_core_params = (
-    alertthreshold_vec = optimal_threshold_alertthreshold_vec,
-    ensemble_specification = ensemble_specification,
-    outbreak_specification = ensemble_outbreak_specification,
-    moving_avg_detection_lag = ensemble_moving_avg_detection_lag,
[…]

### 2024-10-22 manuscript/optimal-thresholds-loading.jl

@@ -1,0 +1,111 @@
+#%%
+using DrWatson
+@quickactivate "OutbreakDetection"
+
+using OutbreakDetectionUtils:
+    IndividualTestSpecification, create_combinations_vec,
+    DynamicalNoiseSpecification, PoissonNoiseSpecification
+using OutbreakDetection: collect_OptimalThresholdCharacteristics
+
+include(srcdir("ensemble-parameters.jl"))
+
+if false
+    include("../src/ensemble-parameters.jl")
+end
+
+#%%
+rdt_test_spec_vec = [IndividualTestSpecification(0.85, 0.85, 0),
+    IndividualTestSpecification(0.9, 0.9, 0)]
+
+elisa_test_spec_vec = [IndividualTestSpecification(1.0, 1.0, 0),
+    IndividualTestSpecification(1.0, 1.0, 14)]
+
+optimal_threshold_test_spec_vec = vcat(rdt_test_spec_vec, elisa_test_spec_vec)
+
+optimal_threshold_alertthreshold_vec = collect(1:1:15)
+
+R_0_vec = [16.0]
+
+ensemble_dynamics_spec_vec = create_combinations_vec(
+    DynamicsParameters,
+    (
+        [ensemble_state_specification.init_states.N],
+        [27],
+        [0.2],
+        [SIGMA],
+        [GAMMA],
+        R_0_vec,
+        [0.8],
+    ),
+)
+
+ensemble_spec_vec = create_combinations_vec(
+    EnsembleSpecification,
+    (
+        [ensemble_model_type],
+        [ensemble_state_specification],
+        ensemble_dynamics_spec_vec,
+        [ensemble_time_specification],
+        [ensemble_nsims],
+    ),
+)
+
+alert_method_vec = ["movingavg"]
+
+#%%
+ensemble_noise_specification =
+    filter(ensemble_noise_specification_vec) do noise_spec
+        noise_spec.noise_type == "poisson" ||
+            noise_spec.correlation == "in-phase"
+    end
+
+ensemble_specification = ensemble_spec_vec[1]
+alertmethod = alert_method_vec[1]
+
+optimal_threshold_core_params = (
+    alertthreshold_vec = optimal_threshold_alertthreshold_vec,
+    ensemble_specification = ensemble_specification,
+    outbreak_specification = ensemble_outbreak_specification,
+    moving_avg_detection_lag = ensemble_moving_avg_detection_lag,
+    percent_visit_clinic = ensemble_percent_visit_clinic,
[…]

### 2024-10-22 manuscript/optimal-thresholds-tables.jl

@@ -1,0 +1,173 @@
+#%%
+using DrWatson
+@quickactivate "OutbreakDetection"
+
+using DataFrames
+using CategoricalArrays
+using Match: Match
+using CSV: CSV
+
+using OutbreakDetectionUtils: create_optimal_thresholds_df
+
+include(projectdir("manuscript", "optimal-thresholds-loading.jl"));
+
+if false
+    include("optimal-thresholds-loading.jl")
+end
+
+#%%
+cfr_df = CSV.read(
+    datadir("CFR_2022.csv"),
+    DataFrame; delim = ',',
+    header = true,
+    types = [String, Int64, Float64],
+    silencewarnings = true,
+)
+dropmissing!(cfr_df)
+
+gha_cfr = only(cfr_df[cfr_df.country .== "GHA", :CFR])
+
+population_df = CSV.read(
+    datadir("input-populations.csv"),
+    DataFrame; delim = ',',
+    header = true,
+)
+
+gha_2022_pop = only(
+    population_df[population_df.ISO3_code .== "GHA", "2022"]
+)
+gha_2022_scale_population =
+    gha_2022_pop / ensemble_state_specification.init_states.N
+
+countries = [
+    (;
+        name = "Ghana",
+        code = "GHA",
+        cfr = gha_cfr,
+        year = "2022",
+        population_size = gha_2022_pop,
+        scale_population = gha_2022_scale_population,
+    ),
+]
+
+#%%
+dynamical_noise_rdt_optimal_solutions = filter(
+    chars -> in(chars.individual_test_specification, rdt_test_spec_vec),
+    dynamical_noise_optimal_solutions,
+);
+
+poisson_noise_rdt_optimal_solutions = filter(
+    chars -> in(chars.individual_test_specification, rdt_test_spec_vec),
+    poisson_noise_optimal_solutions,
+);
+
+elisa_optimal_solutions = filter(
+    chars -> in(chars.individual_test_specification, elisa_test_spec_vec),
+    dynamical_noise_optimal_solutions,
+);
+
+#%%
+elisa_df = create_optimal_thresholds_df(elisa_optimal_solutions)
+elisa_df[!, :noise_spec] .= "Common"
+
+rdt_dynamical_df = create_optimal_thresholds_df(
+    dynamical_noise_rdt_optimal_solutions
+)
+rdt_dynamical_df[!, :noise_spec] .= "Dynamical noise: in-phase"
+
+rdt_poisson_df = create_optimal_thresholds_df(
[…]

### 2024-10-22 manuscript/optimal-thresholds.jl

@@ -1,317 +1,0 @@
-#%%
-using DrWatson
-@quickactivate "OutbreakDetection"
-
-using ProgressMeter
-using FLoops
-using NaNMath: NaNMath
-using CSV: CSV
-using DataFrames
-using CategoricalArrays
-using DataFramesMeta
-using Statistics
-using Match
-
-using OutbreakDetectionUtils
-
-includet(srcdir("makie-plotting-setup.jl"))
-include(srcdir("ensemble-parameters.jl"))
-
-#%%
-rdt_spec_vec = [
-    IndividualTestSpecification(0.85, 0.85, 0),
-    IndividualTestSpecification(0.9, 0.9, 0),
-    IndividualTestSpecification(1.0, 0.0, 0),
-]
-
-elisa_test_spec_vec = [
-    IndividualTestSpecification(1.0, 1.0, 0),
-    IndividualTestSpecification(1.0, 1.0, 3),
-    IndividualTestSpecification(1.0, 1.0, 7),
-    IndividualTestSpecification(1.0, 1.0, 14),
-]
-
-optimal_threshold_alertthreshold_vec = collect(1:1:15)
-
-R_0_vec = collect(16.0)
-
-ensemble_dynamics_spec_vec = create_combinations_vec(
-    DynamicsParameters,
-    (
-        [ensemble_state_specification.init_states.N],
-        [27],
-        [0.2],
-        [SIGMA],
-        [GAMMA],
-        R_0_vec,
-        [0.8],
-    ),
-)
-
-ensemble_spec_vec = create_combinations_vec(
-    EnsembleSpecification,
-    (
-        [ensemble_model_type],
-        [ensemble_state_specification],
-        ensemble_dynamics_spec_vec,
-        [ensemble_time_specification],
-        [ensemble_nsims],
-    ),
-)
-
-# alert_method_vec = ["movingavg", "dailythreshold_movingavg"]
-alert_method_vec = ["movingavg"]
-
-#%%
-cfr_df = CSV.read(
-    datadir("CFR_2022.csv"),
-    DataFrame; delim = ',',
-    header = true,
-    types = [String, Int64, Float64],
-    silencewarnings = true,
-)
-dropmissing!(cfr_df)
-
-gha_cfr = only(cfr_df[cfr_df.country .== "GHA", :CFR])
-
-population_df = CSV.read(
-    datadir("input-populations.csv"),
-    DataFrame; delim = ',',
-    header = true,
-)
-
[…]

### 2024-10-22 manuscript/manuscript.qmd
Create manuscript figures directly within the manuscript file
Helps split the work into smaller sections and reduce recomputations
when sharing objects between plot and table creation scripts

@@
 #| label: fig-accuracy
-accuracy_line_plot
+line_plot(
+    optimal_threshold_characteristics;
+    outcome = :accuracy,
+    ylabel = "Outbreak Detection\nAccuracy",
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (0.5, 1.0),
+    force = true,
+    save_plot = false,
+    clinical_hline = clinical_hline,
+)
 ```
@@
 #| label: fig-unavoidable
-unavoidable_line_plot
+line_plot(
+    optimal_threshold_characteristics;
+    outcome = :unavoidable_cases,
+    ylabel = "Unavoidable Cases",
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (0, 3.5e4),
+    force = true,
+    save_plot = false,
+    clinical_hline = clinical_hline,
+)
 ```
@@
 #| label: fig-delay
-delay_line_plot
+line_plot(
+    optimal_threshold_characteristics;
+    outcome = :detectiondelays,
+    ylabel = "Detection Delays\n(Days)",
+    hlines = (0.0),
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (-100, 100),
+    force = true,
+    save_plot = false,
+    clinical_hline = clinical_hline,
+)
 ```
@@
 With large amounts of dynamical noise (8 times the measles incidence), the mean detection delay of the 90% and 85% RDTs range from -17.5 days to 3.2 days, and from -25.2 days to -3.4 days, respectively.
-Negative delays indicate that alerts are being triggered before the start of the outbreak and is correlated with the proportion of the time series that is under alert, with larger negative delays associated with more and/or longer alert periods (@fig-alert-proportion).
+Negative delays indicate that alerts are being triggered before the start of the outbreak and is correlated with the proportion of the time series that is under alert, with larger negative delays associated with more and/or longer alert periods (@fig-alert-proportion, Supplemental Figure 2).
 

### 2024-10-22 manuscript/supplemental-appendix.qmd
Update supplement figures to create directly in file

@@
 ```{julia}
-#| fig-cap: "The accuracy of outbreak detection systems under different testing rates and noise structures.
+#| fig-cap: "The difference between proportion of the time series in outbreak for outbreak detection systems under different testing rates and noise structures.
 The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
-#| label: fig-accuracy
-accuracy_line_plot
+#| label: fig-outbreak-proportion
+line_plot(
+    optimal_threshold_characteristics;
+    outcome = :proportion_timeseries_in_outbreak,
+    ylabel = "Proportion of Time\nSeries In Outbreak",
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (0.0, 0.35),
+    force = true,
+    save_plot = false,
+    clinical_hline = clinical_hline,
+)
 ```
@@
 ```{julia}
-#| fig-cap: "The number of unavoidable cases of outbreak detection systems under different testing rates and noise structures.
+#| fig-cap: "The alert durations of outbreak detection systems under different testing rates and noise structures.
 The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
-#| label: fig-unavoidable
-unavoidable_line_plot
+#| label: fig-alert-duration
+line_plot(
+    optimal_threshold_characteristics;
+    outcome = :alert_duration_vec,
+    ylabel = "Alert Duration\n(Days)",
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (0, 170),
+    force = true,
+    save_plot = false,
+    clinical_hline = clinical_hline,
+)
 ```
-
-```{julia}
-#| fig-cap: "The detection delay of outbreak detection systems under different testing rates and noise structures.
-The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
-#| label: fig-delay
-delay_line_plot
-```
-
-```{julia}
-#| fig-cap: "The difference between proportion of the time series in alert for outbreak detection systems under different testing rates and noise structures.
-The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
-#| label: fig-alert-proportion
-alert_proportion_line_plot
-```

### 2024-10-22 manuscript/schematic-plot.jl
Update schematic plot to use new outbreak and alert bounds array format and clean up xlims handling

@@ -1,3 +1,4 @@
+#%%
 using DrWatson
 @quickactivate "OutbreakDetection"
 
@@ -9,6 +10,7 @@ using OutbreakDetectionUtils
 
 include(srcdir("makie-plotting-setup.jl"))
 
+#%%
 states_p = StateParameters(;
     N = 500_000, s_prop = 0.05, e_prop = 0.00, i_prop = 0.00
 )
@@ -61,6 +63,7 @@ outbreak_detection_specification = OutbreakDetectionSpecification(
     "movingavg",
 )
 
+#%%
 function get_outbreak_status(
     inc_vec, outbreak_specification
 )
@@ -95,6 +98,7 @@ function get_outbreak_status(
     return outbreak_status, outbreak_bounds
 end
 
+#%%
 function shift_vec(invec, shift::T) where {T<:Integer}
     if shift == 0
         return invec
@@ -113,6 +117,7 @@ function shift_vec(invec, shift::T) where {T<:Integer}
     return outvec
 end
 
+#%%
 function create_schematic_simulation(
     states_p, dynamics_p, noise_states_p, noise_dynamics_p, test_specification,
     time_p; seed = 1234,
@@ -189,8 +194,9 @@ function create_schematic_simulation(
         outbreak_detection_specification.alert_threshold,
     )
     alert_bounds = calculate_outbreak_thresholds(
-        rle(alertstatus_vec .> 0); ncols = 2
+        rle(alertstatus_vec .> 0); ncols = 3
     )
+    OutbreakDetectionUtils.calculate_outbreak_duration!(alert_bounds)
 
     return inc_vec,
     outbreak_status,
@@ -201,6 +207,7 @@ function create_schematic_simulation(
     alert_bounds
 end
 
+#%%
 inc_vec, outbreak_status, outbreak_bounds, noise_vec, movingavg_testpositives, alertstatus_vec, alert_bounds = create_schematic_simulation(
     states_p,
     dynamics_p,
@@ -215,6 +222,7 @@ inc_vec, outbreak_status, outbreak_bounds, noise_vec, movingavg_testpositives, a
     shift_noise = -100,
 );
 
+#%%
 function plot_schematic(
     inc_vec,
     outbreakstatus_vec,
@@ -240,8 +248,8 @@ function plot_schematic(
     times = collect(time_p.trange)
 
     if haskey(kwargs_dict, :xlims)
[…]

### 2024-10-22 src/line_plots.jl
Update lineplot to allow scaling of avoidable and unavoidable cases
May want to scale to a population size/per annum

@@ -44,7 +44,7 @@ function line_plot(
             ensemble_percent_clinic_tested_vec,
             optimal_threshold_test_spec_vec,
             optimal_threshold_core_params;
-            clinical_hline = clinical_hline
+            clinical_hline = clinical_hline,
         )
 
         return line_plot(
@@ -67,7 +67,7 @@ function line_plot(
             hlines = hlines,
             force = force,
             save_plot = save_plot,
-            kwargs...
+            kwargs...,
         )
     end
 
@@ -100,6 +100,13 @@ function line_plot(
     plotpath = joinpath(plotdirpath, "$plotname.$plotformat")
 
     if !isfile(plotpath) || force
+        if in(outcome, [:avoidable_cases, :unavoidable_cases])
+            kwargs_dict = Dict{Symbol,Any}(kwargs)
+            if !haskey(kwargs_dict, :cases_scaling)
+                @error "Cases scaling not provided. Please provide kwarg `cases_scaling`"
+            end
+        end
+
         fig = Figure()
 
         unique_noise_descriptions = unique(
@@ -284,6 +291,7 @@ function _line_plot(
         optimal_thresholds_vec,
         outcome;
         percentiles = [0.1, 0.9],
+        nboots = nothing,
     )
 
     select!(
@@ -297,6 +305,23 @@ function _line_plot(
             x -> endswith(x, "th"),
         ),
     )
+
+    if in(outcome, [:avoidable_cases, :unavoidable_cases])
+        DataFrames.transform!(
+            long_df,
+            DataFrames.Cols(r".*_mean", r".*_[0-9]+th") .=>
+                (
+                    x ->
+                        Int64.(
+                            round.(
+                                x * kwargs_dict[:cases_scaling];
+                                digits = 0,
+                            ),
+                        )
+                );
+            renamecols = false,
+        )
+    end
 
     if show_x_facet_label && i == 1
         x_facet_label = "$(Int64(round(

### 2024-10-22 manuscript/manuscript.qmd
Implement lineplot scaling for unavoidable cases in manuscript

@@
     show_y_facet_label = false,
-    ylims = (0, 3.5e4),
+    ylims = (0, 2.4e4),
     force = true,
@@
     clinical_hline = clinical_hline,
+    cases_scaling = gha_2022_scale_population_per_annum
 )

### 2024-10-22 manuscript/manuscript.qmd
Clean up manuscript and supplement tables and figures

@@
 
-```{julia}
+```{julia optimal-threshold-setup}
 #| echo: false
-include("optimal-thresholds.jl");
+include("optimal-thresholds-tables.jl");
 ```
@@
 ```{=typst}
-#let optimal_thresholds = csv("thresholds.csv")
+#let optimal_thresholds = csv("optimal-thresholds.csv")
 
 #table(
-  columns: 10,
+  columns: 9,
   fill: (x, y) => {
@@
   align: center,
-  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
+  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 6, align: center, "Testing Rate"),
   ..optimal_thresholds.flatten()
@@
 
-```{julia}
+```{julia plotting-setup}
 #| echo: false
-include("line-plots-script.jl");
+include(srcdir("makie-plotting-setup.jl"));
+using OutbreakDetection: line_plot
 ```
@@
 #| label: fig-alert-proportion
-alert_proportion_line_plot
+line_plot(
+    optimal_threshold_characteristics;
+    outcome = :proportion_timeseries_in_alert,
+    ylabel = "Proportion of Time Series\nIn Alert",
+    hlines = (0.0),
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (0, 0.35),
+    force = true,
+    save_plot = false,
+    clinical_hline = clinical_hline,
+)
 ```

### 2024-10-22 manuscript/supplemental-appendix.qmd
Clean up manuscript and supplement tables and figures

@@
 #table(
-  columns: 10,
+  columns: 9,
     fill: (x, y) => {
@@
     },
-  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
+  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 6, align: center, "Testing Rate"),
   ..accuracy.flatten()
@@
 ```{=typst}
-#let optimal_thresholds = csv("optimal-thresholds-unavoidable.csv")
+#let optimal_thresholds = csv("optimal-thresholds_unavoidable-cases.csv")
 
 #table(
-  columns: 10,
+  columns: 9,
     fill: (x, y) => {
@@
     },
-  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 7, align: center, "Testing Rate"),
+  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 6, align: center, "Testing Rate"),
   ..optimal_thresholds.flatten()
@@
 ```{=typst}
-#let optimal_thresholds = csv("optimal-thresholds-delays.csv")
+#let optimal_thresholds = csv("optimal-thresholds_detection-delays.csv")
 
 #table(
-  columns: 10,
+  columns: 9,
     fill: (x, y) => {

### 2024-10-22 manuscript/line-plots-values-check.jl

@@ -6,7 +6,6 @@ using OutbreakDetectionUtils:
     IndividualTestSpecification, DynamicalNoiseSpecification
 using StatsBase: mean
 
-include(srcdir("makie-plotting-setup.jl"))
 
 include(projectdir("manuscript", "optimal-thresholds-loading.jl"));
 
@@ -97,70 +96,3 @@ mapreduce(
 ) do (label, mean_alert_duration_vec)
     Dict(label => round.(extrema(mean_alert_duration_vec); digits = 1))
 end
-
-#%%
-# outbreak_proportion_line_plot = line_plot(
-#     optimal_threshold_characteristics;
-#     outcome = :proportion_timeseries_in_outbreak,
-#     ylabel = "Proportion of Time Series\nIn Outbreak",
-#     plotdirpath = baseplotdirpath,
-#     facet_fontsize = 18,
-#     labelsize = 20,
-#     show_x_facet_label = true,
-#     show_y_facet_label = false,
-#     ylims = (0.0, 0.25),
-#     force = true,
-#     save_plot = false,
-# clinical_hline = clinical_hline,
-# )
-#
-# #%%
-
-# #%%
-# nalerts_line_plot = line_plot(
-#     optimal_threshold_characteristics;
-#     outcome = :nalerts,
-#     ylabel = "Number of Alerts",
-#     plotdirpath = baseplotdirpath,
-#     facet_fontsize = 18,
-#     labelsize = 20,
-#     show_x_facet_label = true,
-#     show_y_facet_label = false,
-#     ylims = (0, 350),
-#     force = true,
-#     save_plot = false,
-# clinical_hline = clinical_hline,
-# )
-#
-# #%%
-# nalerts_per_outbreak_line_plot = line_plot(
-#     optimal_threshold_characteristics;
-#     outcome = :n_alerts_per_outbreak,
-#     ylabel = "Number of Alerts per Outbreak",
-#     plotdirpath = baseplotdirpath,
-#     facet_fontsize = 18,
-#     labelsize = 20,
-#     show_x_facet_label = true,
-#     show_y_facet_label = false,
-#     ylims = (0, 9),
-#     force = true,
-#     save_plot = false,
-# clinical_hline = clinical_hline,
-# )
-#
-# #%%
-# alert_outbreak_proportion_line_plot = line_plot(
-#     optimal_threshold_characteristics;
-#     outcome = :alert_outbreak_timeseries_prop_diff,
[…]

### 2024-10-22 manuscript/optimal-thresholds-tables.jl

@@ -62,24 +62,6 @@ elisa_optimal_solutions = filter(
 );
 
 #%%
-elisa_df = create_optimal_thresholds_df(elisa_optimal_solutions)
-elisa_df[!, :noise_spec] .= "Common"
-
-rdt_dynamical_df = create_optimal_thresholds_df(
-    dynamical_noise_rdt_optimal_solutions
-)
-rdt_dynamical_df[!, :noise_spec] .= "Dynamical noise: in-phase"
-
-rdt_poisson_df = create_optimal_thresholds_df(
-    poisson_noise_rdt_optimal_solutions
-)
-rdt_poisson_df[!, :noise_spec] .= "Poisson noise"
-
-thresholds_df = vcat(
-    rdt_dynamical_df, rdt_poisson_df, elisa_df
-)
-
-#%%
 elisa_df = create_optimal_threshold_summary_df(
     elisa_optimal_solutions,
     [:unavoidable_cases, :detectiondelays, :alert_duration_vec];
@@ -91,7 +73,7 @@ elisa_df
 
 rdt_dynamical_df = create_optimal_threshold_summary_df(
     dynamical_noise_rdt_optimal_solutions,
-    [:unavoidable_cases, :detectiondelays, :alert_duration_vec],
+    [:unavoidable_cases, :detectiondelays, :alert_duration_vec];
     percentiles = nothing,
     nboots = nothing,
 )
@@ -99,7 +81,7 @@ rdt_dynamical_df[!, :noise_spec] .= "Dynamical noise: in-phase"
 
 rdt_poisson_df = create_optimal_threshold_summary_df(
     poisson_noise_rdt_optimal_solutions,
-    [:unavoidable_cases, :detectiondelays, :alert_duration_vec],
+    [:unavoidable_cases, :detectiondelays, :alert_duration_vec];
     percentiles = nothing,
     nboots = nothing,
 )
@@ -109,7 +91,14 @@ thresholds_df = vcat(
     rdt_dynamical_df, rdt_poisson_df, elisa_df
 )
 
-thresholds_df[!, :unavoidable_cases] = Int64.(round.(thresholds_df[!, :unavoidable_cases] * gha_2022_scale_population_per_annum; digits = 0))
+thresholds_df[!, :unavoidable_cases] =
+    Int64.(
+        round.(
+            thresholds_df[!, :unavoidable_cases] *
+            gha_2022_scale_population_per_annum;
+            digits = 0,
+        )
+    )
 
[…]

### 2024-10-24 manuscript/manuscript.qmd
Add comments of things to update in manuscript

@@
 
+_Mention that the same model was used to define the noise as the outbreaks._
+
 ## Defining Outbreaks
@@
 
+_Need to mention that an alert is any time the threshold is exceeded, regardless of the number of consecutive days of exceedance._
+_Clearly state that 1 day of exceedance is and 30 days of consecutive exceedance are both treated as a single alert i.e., the same._
+
 The use of alert thresholds is common in locations burdened by measles, where access to high performance computing is limited making nowcasting-style approaches to outbreak detection impractical.
@@
 
-- No test, where every suspected measles case is counted as positive for measles and used to trigger an outbreak alert, which is equivalent to the use of the clinical case definition.
+<!-- - No test, where every suspected measles case is counted as positive for measles and used to trigger an outbreak alert, which is equivalent to the use of the clinical case definition.
 This is implemented as a diagnostic test with 100% sensitivity, 0% specificity, a 0-day lag in returning the result, and all individuals who visit the clinic are tested (not varied between 10-60%, like the other diagnostic tests described).
+-->
 - An RDT equivalent with 85% sensitivity and specificity, and 0-day lag in result return.
@@
 - An ELISA equivalent test with 100% sensitivity and specificity with a 0-day test result delay.
-- An ELISA equivalent test with 100% sensitivity and specificity with a 7-day test result delay.
+<!--- An ELISA equivalent test with 100% sensitivity and specificity with a 7-day test result delay.-->
 - An ELISA equivalent test with 100% sensitivity and specificity with a 14-day test result delay.
@@
 
-> Despite the same average incidence of noise in each (comparable) scenario, the relative proportion of measles to noise varies differently throughout the time series, exacerbating the effects of imperfect diagnostic tests that will produce higher rates of false positives and negatives than ELISA-like diagnostics.
-Imperfect tests are also more susceptible to incorrect results because of the varying prevalence throughout the time series, so the PPV of the system will not be static throughout the time series.
-
 Under the assumption that non-measles febrile rash is relatively static in time (Poisson), RDTs can perform as well, if not better than ELISA tests at moderate to high testing rates, and at a fraction of the cost.
@@
 74% accuracy, relative to ELISA’s 93%.
+This occurs because, despite the same average incidence of noise in each (comparable) scenario, the relative proportion of measles to noise varies differently throughout the time series, exacerbating the effects of imperfect diagnostic tests that will produce higher rates of false positives and negatives than ELISA-like diagnostics.
+Imperfect tests are also more susceptible to incorrect results due to the varying prevalence throughout the time series, so the PPV of the system will not be static throughout the time series.
+
+- _Make point about balance of false positive/negative outbreak detection_
+- _Analysis isn't true optimization:_
+  - _Requires explicit decisions about preference for speed / false alerts vs higher PPV_
+  - _Surveillance is counting for action (WHO quote)_
 
@@
 # Funding
+
+- _Something about GAVI/Gates_
 

### 2024-10-24 manuscript/supplemental-appendix.qmd
Add comments of things to update in manuscript

@@
 
-The outbreak detection accuracy for each testing scenario, at the optimal thresholds.
+Mean outbreak detection accuracy of each testing scenario at their specific optimal thresholds, when the average noise incidence is 8 times higher than the average measles incidence.
+A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence.
+B) the noise structure is Poisson only.
 
@@
 
-Unavoidable cases per annum of each testing scenario at their specific optimal thresholds, scaled up to Ghana’s 2022 population.
+Mean unavoidable cases per annum of each testing scenario at their specific optimal thresholds, scaled up to Ghana’s 2022 population, when the average noise incidence is 8 times higher than the average measles incidence.
 A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence.
@@
 
-Outbreak alert delay (days) of each testing scenario at their specific optimal thresholds.
+Mean outbreak alert delay (days) of each testing scenario at their specific optimal thresholds, when the average noise incidence is 8 times higher than the average measles incidence.
 A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence.

### 2024-10-28 manuscript/manuscript.qmd
Add Matt's abstract suggestions

@@
   However, the interaction of the two levels’ (the test and the alert system) metrics, is typically overlooked.
-  Here, we describe how equivalent regions of detection accuracy can exist over a range of diagnostic test characteristics, examining the sensitivity to background noise structure.
+  Here, we describe how equivalent regions of detection accuracy can exist over a range of diagnostic test characteristics, examining the sensitivity to background noise structure and magnitude.
 
@@
   We generated a stochastic SEIR model with importation to simulate true measles and non-measles sources of febrile rash (noise) daily incidence.
-  We simulated the noise to have in-phase, out-of-phase, non-seasonal dynamics, with respect to the measles simulation.
-  We also examined noise where each day were independent Poisson draws.
-  For rapid diagnostic test (RDT) and enzyme-linked immunosorbent assay (ELISA) like tests, we varied the reporting rate, and the observed daily incidence threshold required to trigger an alert, and computed the: percentages of alerts that were correct (analogous to the positive predictive value), the percentage of outbreaks detected (analogous to the sensitivity), the delay between outbreak initiation and alert, and the number of (un)avoidable outbreak cases.
-  Summarizing the percentage of alerts correct and outbreaks detected into an accuracy metric, we selected the “optimal threshold” for each combination of diagnostic test and reporting rate, presenting the summary statistics above.
+  We generated time series of febrile rash (i.e., measles clinical case definition) by summing the daily incidence of measles and either independent Poisson noise or non-measles dynamical noise (consistent with rubella virus).
+  For each time series we assumed a fraction of all cases were seen at a healthcare clinic, and a subset of those were diagnostically confirmed using a test with sensitivity and specificity consistent with either a rapid diagnostic test (RDT) or an enzyme-linked immunosorbent assay (ELISA).
+  From the resulting time series of test-positive cases, we define an outbreak alert as the exceedance of a threshold by the 7-day rolling average of observed (test positive) cases.
+  For each threshold level, we calculated percentages of alerts that were aligned with an outbreak (analogous to the positive predictive value), the percentage of outbreaks detected (analogous to the sensitivity), and combined these two measures into an accuracy metric for outbreak detection.
+  We selected the optimal threshold as the value that maximizes accuracy.
+  We show how the optimal threshold and resulting accuracy depend on the diagnostic test, reporting rate, and the type and magnitude of the non-measles noise.
 
@@
 
-  The optimal threshold for each test increased monotonically as the percentage of clinic visited who were tested increased, though there are adjacent reporting rates that have identical thresholds due to the integer requirement for the daily threshold.
+  The optimal threshold for each test increased monotonically as the percentage of clinic visits who were tested increased.
   With Poisson-only noise, similar outbreak detection accuracies could be achieved with imperfect RDT-like tests as with ELISA-like diagnostic tests (c.
   93%), given moderately high testing rates.
-  With larger delays (14 days) between the ELISA test procurement and result date, RDTs could outperform the ELISA.
+  With larger delays (14 days) between the ELISA test administration and result date, RDTs could outperform the ELISA.
   Similar numbers of unavoidable cases and outbreak alert delays could be achieved between the test types.
-  With dynamical, in-phase, noise, however, the accuracy of ELISA scenarios was far superior to those achieved with RDTs (c.
+  With dynamical noise, however, the accuracy of ELISA scenarios was far superior to those achieved with RDTs (c.
 93% vs.
@@
 
-  The performance of an outbreak detection system is highly sensitive to the structure and the level of background noise.
-  Under the assumption that the noise is relatively static over time (as influenza-like illness was early in the COVID-19 pandemic), RDTs can perform as well as ELISA in a surveillance system.
-  However, when the noise is non-stationary, imperfect tests cannot overcome their accuracy limitations through higher testing rates. 
+  The performance of an outbreak detection system is highly sensitive to the structure and the magnitude of background noise.
+  Under the assumption that the noise is relatively static over time, RDTs can perform as well as ELISA in a surveillance system.
+  However, when the noise is temporally correlated, as from a separate SEIR process, imperfect tests cannot overcome their accuracy limitations through higher testing rates. 
 toc: false

### 2024-10-28 manuscript/manuscript.qmd
Add Matt's suggestions for background

@@
 
-Much like the accuracy of an individual test, outbreak detection systems face the same issue regarding the prioritization of sensitive vs specific alerts [@germanSensitivityPredictiveValue2000; @worldhealthorganizationOperationalThresholds2014; @lewisTimelyDetectionMeningococcal2001].
+Much like the accuracy of an individual test, outbreak detection systems face the same issue regarding the prioritization of sensitive or specific alerts [@germanSensitivityPredictiveValue2000; @worldhealthorganizationOperationalThresholds2014; @lewisTimelyDetectionMeningococcal2001].
 For many disease systems, particularly in resource constrained environments where the burden of infectious diseases is highest [@gbd2019childandadolescentcommunicablediseasecollaboratorsUnfinishedAgendaCommunicable2023; @roserBurdenDisease2023], cases are counted and if a pre-determined threshold is breached, be that weekly, monthly, or some combination of the two, an alert is triggered that may launch a further investigation and/or a response [@worldhealthorganizationMeaslesOutbreakGuide2022; @worldhealthorganizationOperationalThresholds2014].
-In effect, this discretizes a distinctly continuous phenomenon into a binary measure, outbreak or no outbreak, for decision making purposes.
-When considering reactive vaccination campaigns and non-pharmaceutical based interventions designed to reduce transmission and limit and suppress outbreaks through direct and indirect protection, early action has the potential to avert the most cases, and in some instances, can be the most cost effective as it provides the opportunity to incorporate later information within an “adaptive management” framework [@atkinsAnticipatingFutureLearning2020; @taoLogisticalConstraintsLead; @graisTimeEssenceExploring2008; @ferrariTimeStillEssence2014; @worldhealthorganizationConfirmingInvestigatingManaging2009; @minettiLessonsChallengesMeasles2013].
-While this framing would point towards a sensitive surveillance system being optimal, with the benefits clearly defined in terms of morbidity and mortality avoided, each action comes with both direct and indirect financial and opportunity costs stemming from unnecessary activities that remove resources from future response capabilities.
+In effect, this discretizes a distinctly continuous phenomenon (observed cases) into a binary measure, outbreak or no outbreak, for decision making purposes.
+Reactive interventions, such as vaccination campaigns and non-pharmaceutical based interventions, designed to reduce transmission or limit and suppress outbreaks, early action has the potential to avert the most cases [@atkinsAnticipatingFutureLearning2020; @taoLogisticalConstraintsLead; @graisTimeEssenceExploring2008; @ferrariTimeStillEssence2014; @worldhealthorganizationConfirmingInvestigatingManaging2009; @minettiLessonsChallengesMeasles2013].
+While this framing would point towards a sensitive (i.e., early alert) surveillance system being optimal, each action comes with both direct and indirect financial and opportunity costs stemming from unnecessary activities that limit resources for future response capabilities.
 Just as the balance of sensitivity and specificity of a test for an individual must be carefully evaluated, so must the balance at the outbreak level.
@@
 With their success within malaria surveillance systems, and particularly since the COVID-19 pandemic, rapid diagnostic tests (RDTs) have garnered wider acceptance, and their potential for use in other disease systems has been gaining interest [@warrenerEvaluationRapidDiagnostic2023].
-Despite concerns of their lower diagnostic accuracy slowing their adoption until recently [@millerAddressingBarriersDevelopment2015], the reduced cold-chain requirements [@brownRapidDiagnosticTests2020], and gains in speed of result and the impact on the timeliness of an alert system may outweigh the costs in some settings [@warrenerEvaluationRapidDiagnostic2023; @mcmorrowMalariaRapidDiagnostic2011; @larremoreTestSensitivitySecondary2021; @middletonModelingTransmissionMitigation2023].
+Despite concerns of their lower diagnostic accuracy slowing their adoption until recently [@millerAddressingBarriersDevelopment2015], RDTs generally have reduced cold-chain requirements [@brownRapidDiagnosticTests2020] and faster speed of result that has been show to outweigh the cost of false positive/negative results in some settings [@warrenerEvaluationRapidDiagnostic2023; @mcmorrowMalariaRapidDiagnostic2011; @larremoreTestSensitivitySecondary2021; @middletonModelingTransmissionMitigation2023].
 
-In this paper we examine how the interaction between imperfect diagnostic tests affects the optimal outbreak alert method for the detection of measles outbreaks, in the context of a febrile rash surveillance system that detects and tests suspected measles cases i.e., an individual with fever and maculopapular (non-vesicular) rash (or who the healthcare worker suspects has measles) [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018].
+In this paper we examine how the use of imperfect diagnostic tests affects the performance of outbreak detection for measles outbreaks in the context of a febrile rash surveillance system that includes both measles and non-measles cases.
 Because measles symptoms are non-specific, it is important to account for non-measles sources of febrile rash e.g., rubella, parvovirus, varicella, etc., producing the potential for false positive results in the context of imperfect tests.
-Currently, measles outbreaks are declared on the basis of either suspected measles cases alone, the use of an enzyme-linked immunosorbent assay (ELISA) to detect the presence of measles-specific IgM antibodies, or a combination of the two, depending on the elimination status of the region, with countries nearer elimination increasing the use of ELISA diagnostic confirmation for suspected cases [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018].
+Currently, measles outbreaks are declared on the basis of either suspected measles cases (i.e., an individual with fever and maculopapular rash [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018]) alone, cases confirmed by enzyme-linked immunosorbent assay (ELISA) to detect the presence of measles-specific IgM antibodies, or a combination of the two, depending on the elimination status of the region, with countries nearer elimination increasing the use of ELISA diagnostic confirmation for suspected cases [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018].
 Each of these detection systems have their flaws.
-Although clinical case definition is very fast and requires limited resources, providing a highly sensitive alert system, in the face of high “background noise” from non-measles sources of febrile rash relative to true measles prevalence, it can lead to great uncertainty in the alert system and low positive predictive value (PPV), i.e., the probability that an alert accurately captures an outbreak [@hutchinsEvaluationMeaslesClinical2004].
-And while ELISA confirmation is the standard diagnostic test for measles surveillance [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018], it is expensive due to its cold-chain and transport costs and delays resulting from the requirement to transfer samples from point of care to a central laboratory, leading to a slower and less sensitive (but more specific) alert system.
+Although clinical case definition is very fast and requires limited resources, it is highly sensitive, and in the face of high “background noise” from non-measles sources of febrile rash, can lead to low positive predictive value (PPV), i.e., the probability that an alert accurately captures an outbreak [@hutchinsEvaluationMeaslesClinical2004].
+And while ELISA confirmation is the standard diagnostic test for measles surveillance and has higher specificity [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018], the training and facility requirements generally mean that samples must be transported from the point of care to a separate laboratory, which incurs both costs and delays _**[ADD REFERENCE HERE]**_.
+In resource-poor settings these delays may be days to weeks.
 A rapid diagnostic test offers the opportunity to provide a compromise between diagnostic accuracy and timeliness, and recent developments show encouraging signs in the field as well as in theory [@warrenerEvaluationRapidDiagnostic2023; @brownRapidDiagnosticTests2020].

### 2024-10-29 scripts/ensemble-sim.jl
Update ensemble sim setup to use newly calculated values for dynamic vaccination

@@ -103,7 +103,7 @@ dynamical_noise_correlation = [
 ]
 dynamical_noise_mean_scaling_vec = [0.15]
 dynamical_noise_vaccination_coverage_vec = [
-    0.854, 0.738, 0.508, 0.279, 0.050
+    0.8538, 0.7383, 0.5088, 0.2789, 0.0492
 ]
 dynamical_noise_spec_vec = create_combinations_vec(
     DynamicalNoiseSpecification,

### 2024-10-29 manuscript/manuscript.qmd
Start to update methods of manuscript with clearer description of noise scaling

@@
 
-We constructed a stochastic compartmental non-age structured SEIR model of measles and simulated using a modified Tau-leaping algorithm with a time step of 1 day, utilizing binomial draws to ensure compartment sizes remained positive valued [@chatterjeeBinomialDistributionBased2005; @gillespieApproximateAcceleratedStochastic2001].
-To stochastically reintroduce infections and cause recurring epidemics, commuter-style imports were added that are proportional to the size of the population and $R_{0}$ [@keelingModelingInfectiousDiseases2008], and the transmission parameter ($\beta_{t}$) is sinusoidal with a period of one year and 20% seasonal amplitude.
+We constructed a stochastic compartmental non-age structured Susceptible-Exposed-Infected-Recovered (SEIR) model of measles and simulated using a modified Tau-leaping algorithm with a time step of 1 day, utilizing binomial draws to ensure compartment sizes remained positive valued [@chatterjeeBinomialDistributionBased2005; @gillespieApproximateAcceleratedStochastic2001].
+We assumed that transmission rate ($\beta_{t}$) is sinusoidal with a period of one year and 20% seasonal amplitude.
 $R_{0}$ was set to 16, with a latent period of 10 days and infectious period of 8 days.
 The population was initialized with 500,000 individuals with Ghana-like birth and vaccination rates, and the final results were scaled up to the approximate 2022 population size of Ghana (33 million) [@worldbankGhana].
+ We assumed commuter-style imports at each time step to avoid extinction; the number of imports each day were drawn from a Poisson distribution with mean proportional to the size of the population and $R_{0}$ [@keelingModelingInfectiousDiseases2008].
 The full table of parameters can be found in @tbl-model-parameters.
@@
 #let import_rate = $(1.06*μ*R_0)/(√(N))$
-#let parameter_labels = ( "Parameters", $R_0$, $"Latent period ("#sym.sigma")"$, $"Infectious period ("#sym.gamma")"$, "Seasonal amplitude", $"Birth rate ("#sym.mu")"$, $"Vaccination rate at birth ("#sym.rho")"$)
+#let parameter_labels = ( "Parameters", $R_0$, $"Latent period ("#sym.sigma")"$, $"Infectious period ("#sym.gamma")"$, "Seasonal amplitude", $"Birth/death rate ("#sym.mu")"$, $"Vaccination rate at birth ("#sym.rho")"$)
 
@@
 To examine the sensitivity of the detection system to background noise, we layered the measles incidence time series with one of four noise time series structures: Poisson-only noise; or dynamical noise with rubella-like parameters that could be in- or out-of-phase, or with independent seasonality to the measles dynamics.
-Only  dynamical in-phase noise and Poisson-only noise are presented in this manuscript.
-For each noise structure, five different magnitudes of noise were simulated and evaluated: the total annual incidence of noise in each system was either the same magnitude as the average annual measles incidence, two times, four times, six times, or eight times the average annual measles incidence.
+For Poisson-only noise, the number of non-measles febrile rash cases each day were independent draws from a Poisson distribution with mean $\lambda$.
+For dynamical noise, we generated time series of cases from an SEIR model that matched the measles model in structure, but had $R_0 = 5$, mean latent period of 7 days, and mean infectious period of 14 days (@tbl-model-parameters).
+The seasonality for the dynamical noise was assumed to be in-phase with measles, anti-phase with measles (peak timing 6 months later), or non-seasonal.
+Only dynamical in-phase noise and Poisson-only noise are presented in the main text; the anti-phase and non-seasonal dynamical noise scenarios are presented in the supplement.
+
+For each noise structure, we simulated five different magnitudes of noise ($\Lambda$).
+$\Lambda$ was calculated as a multiple ($c$) of the average daily measles incidence ($\langle \Delta I_{M} \rangle$): $\Lambda = c \cdot \langle \Delta I_{M} \rangle \text{where} c \in	\{1, 2, 4, 6, 8 \}$.
+For the Poisson-noise scenarios, independent draws from a Poisson distribution with mean $\lambda = \Lambda$ were simulated to produce the noise time series.
+For the dynamical noise scenarios, the vaccination rate at birth was set to 85.4%, 73.8%, 50.8%, 27.9%, and 5.0% to produce equivalent mean daily noise values (to within 2 decimal places): $\Lambda = \langle \Delta I_{R} \rangle = c \cdot \langle \Delta I_{M} \rangle$.
+
+the total annual incidence of noise in each system was either the same magnitude as the average annual measles incidence, two times, four times, six times, or eight times the average annual measles incidence.
 100 time series of 100 years were simulated for each scenario, before summarizing the distributions of outbreak detection methods.
 
-_Mention that the same model was used to define the noise as the outbreaks._
+_**Mention that the same model was used to define the noise as the outbreaks.**_
 

### 2024-10-29 manuscript/manuscript.qmd
Update dynamic noise vaccination values for manuscript

@@
 For the Poisson-noise scenarios, independent draws from a Poisson distribution with mean $\lambda = \Lambda$ were simulated to produce the noise time series.
-For the dynamical noise scenarios, the vaccination rate at birth was set to 85.4%, 73.8%, 50.8%, 27.9%, and 5.0% to produce equivalent mean daily noise values (to within 2 decimal places): $\Lambda = \langle \Delta I_{R} \rangle = c \cdot \langle \Delta I_{M} \rangle$.
+For the dynamical noise scenarios, the vaccination rate at birth was set to 85.38%, 73.83%, 50.88%, 27.89%, and 4.92% to produce equivalent mean daily noise values (to within 2 decimal places): $\Lambda = \langle \Delta I_{R} \rangle = c \cdot \langle \Delta I_{M} \rangle$.
 

### 2024-10-29 src/ensemble-parameters.jl

@@ -48,7 +48,7 @@ dynamical_noise_correlation = ["in-phase", "out-of-phase", "none"]
 dynamical_noise_mean_scaling_vec = [0.15]
 # dynamical_noise_vaccination_coverage_vec = collect(0.0:0.2:0.8)
 dynamical_noise_vaccination_coverage_vec = [
-    0.854, 0.738, 0.508, 0.279, 0.050
+    0.8538, 0.7383, 0.5088, 0.2789, 0.0492
 ]
 dynamical_noise_spec_vec = create_combinations_vec(
     DynamicalNoiseSpecification,

### 2024-10-30 manuscript/manuscript.qmd
Add Matt's suggestions to methods section

@@
 For Poisson-only noise, the number of non-measles febrile rash cases each day were independent draws from a Poisson distribution with mean $\lambda$.
-For dynamical noise, we generated time series of cases from an SEIR model that matched the measles model in structure, but had $R_0 = 5$, mean latent period of 7 days, and mean infectious period of 14 days (@tbl-model-parameters).
+For dynamical noise, we generated time series of cases from an SEIR model that matched the measles model in structure, but had $R_0 = 5$, mean latent period of 7 days, and mean infectious period of 14 days, and added some additional noise independently drawn from a Poisson distribution with mean equal to 15% of the average daily rubella incidence from the SEIR time series to account for non-rubella sources of febrile rash (@tbl-model-parameters) [_**REFERENCES**_].
 The seasonality for the dynamical noise was assumed to be in-phase with measles, anti-phase with measles (peak timing 6 months later), or non-seasonal.
@@
 For the dynamical noise scenarios, the vaccination rate at birth was set to 85.38%, 73.83%, 50.88%, 27.89%, and 4.92% to produce equivalent mean daily noise values (to within 2 decimal places): $\Lambda = \langle \Delta I_{R} \rangle = c \cdot \langle \Delta I_{M} \rangle$.
-
-the total annual incidence of noise in each system was either the same magnitude as the average annual measles incidence, two times, four times, six times, or eight times the average annual measles incidence.
 100 time series of 100 years were simulated for each scenario, before summarizing the distributions of outbreak detection methods.
-
-_**Mention that the same model was used to define the noise as the outbreaks.**_
 
@@
 
-Each of these is a necessary, but not sufficient, condition to the definition of an outbreak; all must be met.
-For example, a region may produce 1000 total cases within the lower and upper bounds defined by the daily incidence being $\ge$ 5 cases, but if this period only lasts for 20 days it would not be considered an outbreak as it would not be possible to mount a response to.
+Only events meeting all 3 criteria are classified as outbreaks
 The incidence of non-measles febrile rash (i.e., noise) does not affect the outbreak status of a region but may affect the alert status triggered by the testing protocol.
 
-## Triggering Alerts
-
-_Need to mention that an alert is any time the threshold is exceeded, regardless of the number of consecutive days of exceedance._
-_Clearly state that 1 day of exceedance is and 30 days of consecutive exceedance are both treated as a single alert i.e., the same._
-
-The use of alert thresholds is common in locations burdened by measles, where access to high performance computing is limited making nowcasting-style approaches to outbreak detection impractical.
-For this reason, we define an “alert” as any single day where the 7-day moving average of the daily incidence is greater than, or equal to, a pre-specified alert threshold.
-To examine the interaction of test and surveillance systems we vary the alert threshold, between 1 and 15 cases per day, before computing the evaluation metrics and identifying the optimal threshold for each combination of individual test, testing rate, noise structure, and noise magnitude.
-We also examined an alert method that required either the daily incidence or the 7-day moving average of daily incidence to exceed the alert threshold (see supplemental results).
-
-Each day, 60% of the measles and non-measles febrile rash cases visit the clinic for treatment, and a percentage of these clinic visits are tested, as all clinic visits are deemed to be suspected measles cases because they meet the clinical case definition.
+Each day, 60% of the measles and non-measles febrile rash cases visit the clinic for treatment, and a percentage (P) of these clinic visits are tested, as all clinic visits are deemed to be suspected measles cases because they meet the clinical case definition.
 This percentage of clinic visits that are tested is varied between 10% and 60%, in 10% increments, for all combinations of diagnostic test (except clinical case definition) and alert threshold, defining the “testing scenario”.
-<!--While a reporting rate of 6-48% is high, we are only simulating a population of 500,000 individuals i.e., approximately 1.5% of Ghana’s population, which produces absolute incidence values that are consistent with observed reporting on the national scale.-->
 Each testing scenario uses one of the following tests:
@@
 
-## Optimal Thresholds
+The time series of "test-positive cases" is the daily count of those tested cases that return a positive test result.
+Thus, for each measles (@fig-outbreak-schematic b) and non-measles noise (@fig-outbreak-schematic a) time series, we have 1 time series of outbreaks (@fig-outbreak-schematic a) and 5 possible time series of test-positive cases (@fig-outbreak-schematic c) (all clinically compatible cases, plus the 4 testing scenarios), which will include false positive and negative cases resulting from imperfect diagnostic tests, that can be used to trigger outbreak alerts.
 
-For each of the previously defined tests, we calculate the number of test positive cases (that will include false positive and negative cases resulting from imperfect diagnostic tests), which is used to categorize the time series by alert status in conjunction with the alert threshold.
-The overlap with true outbreak status, or lack thereof, is used to classify the following metrics, and has been illustrated in @fig-outbreak-schematic:
+## Triggering Alerts
 
-- The percentage of alerts that are 'correct'.
-This is analogous to the PPV of the alert system.
-Note that it is possible to trigger multiple alerts within a single outbreak, and each would be considered correct.
-- The percentage of outbreaks that are detected. This is analogous to the sensitivity of the alert system.
-- The detection delay.
-If an alert precedes the start of the outbreak, as long as it successfully “captures” the outbreak within its bounds, it is considered to be a 'correct' alert, resulting in a negative detection delay i.e., an early warning triggered by false positives.
-- The number of unavoidable and avoidable outbreak cases and deaths.
-Avoidable cases are defined as those that occur within an outbreak after a correct alert is first triggered i.e., cases that could theoretically be prevented with a perfectly effective and timely response.
-Unavoidable cases are the inverse: those that occur before a correct alert, or those that occur in an undetected outbreak.
-Using a fitted age distribution of measles cases in Ghana [@mintaProgressMeaslesElimination2023] and the age-specific case fatality ratio for Ghana in 2022 [@sbarraEstimatingNationallevelMeasles2023], we scaled the values for each one-year age cohort, per annum.
-In practice, not all cases deemed avoidable are (due to imperfect and delays in responses), but to minimize the sensitivity of the results to the response implementation and operational constraints we are counting them as such.
+_**Need to mention that an alert is any time the threshold is exceeded, regardless of the number of consecutive days of exceedance.**_
+_**Clearly state that 1 day of exceedance is and 30 days of consecutive exceedance are both treated as a single alert i.e., the same.**_
+
+We define an "alert" as any contiguous string of 1 or more days where the 7-day moving average of the test-positive cases is greater than, or equal to, a pre-specified alert threshold, T.
+For each time series of test-positive cases, we calculate the percentage of alerts that are "correct", defined as any overlap of 1 or more days between the alert and outbreak period (@fig-outbreak-schematic c); this is analogous to the PPV of the alert system.
+Note that it is possible to have multiple alerts within a single outbreak, if the 7-day moving average of test positive cases drops below the threshold, and each would be considered correct.
+For all outbreaks that occur in the measles time series, we calculate the percentage for which at least 1 alert occurred between its start and end date (@fig-outbreak-schematic b); this is analogous to the sensitivity of the alert system.
+We also calculate the detection delay as the time from the start of an outbreak to the start of its first alert.
+If the alert starts before the outbreak and continues past the start date of the outbreak, this could be considered a correct alert with a negative delay i.e., an early warning triggered by a false positive.
 
@@
 
+Finally, for each time series we calculate the number of unavoidable and avoidable outbreak cases and deaths.
+Avoidable cases are defined as those that occur within an outbreak after a correct alert is first triggered i.e., cases that could theoretically be prevented with a perfectly effective and timely response.
+Unavoidable cases are the inverse: those that occur before a correct alert, or those that occur in an undetected outbreak.
+Using a fitted age distribution of measles cases in Ghana [@mintaProgressMeaslesElimination2023] and the age-specific case fatality ratio for Ghana in 2022 [@sbarraEstimatingNationallevelMeasles2023], we scaled the values for each one-year age cohort, per annum.
+In practice, not all cases deemed avoidable are (due to imperfect and delays in responses), but to minimize the sensitivity of the results to the response implementation and operational constraints we are counting them as such.
+
+We define the accuracy of the surveillance system for a given time series as the mean of the percentage of alerts that are correct and the proportion of the outbreaks with an alert.
+To examine the interaction of test and surveillance systems we varied the alert threshold, between 1 and 15 cases per day, and identified the threshold T for each combination of individual test, testing rate, noise structure and magnitude that maximizes accuracy.
+We also examined an alert method that required either the daily incidence or the 7-day moving average of daily incidence to exceed the alert threshold (see supplemental results).
+
 For each combination of diagnostic test and testing rate, the optimal alert threshold is calculated by selecting the threshold that produces the highest accuracy.
-Each of the 100 simulations produces an accuracy (the mean of the percentages of alerts that are correct and the percentage of the outbreaks that are detected), and the median accuracy of this distribution is used to determine the optimal alert threshold.
-Each testing scenario (the combination of the diagnostic test, testing rate, and alert threshold), is then compared (using the metrics above) at the optimal threshold that is specific for the scenario.
+Each of the 100 simulations per scenario produces an accuracy and the median accuracy of this distribution is used to determine the optimal alert threshold.
+We then compare testing scenarios (the combination of the diagnostic test, testing rate, and alert threshold) at the optimal threshold that is specific for the scenario.
 This allows for conclusions to be made about the surveillance system as a whole, instead of just single components.
@@
     force = true,
-    save_plot = false,
+    plotdirpath = DrWatson.projectdir("manuscript"),
+    plotname = "line_accuracy_plot",
+    save_plot = true,
     clinical_hline = clinical_hline,
@@
     force = true,
-    save_plot = false,
+    plotdirpath = DrWatson.projectdir("manuscript"),
+    plotname = "line_unavoidable_plot",
+    save_plot = true,
     clinical_hline = clinical_hline,
@@
     force = true,
-    save_plot = false,
+    plotdirpath = DrWatson.projectdir("manuscript"),
+    plotname = "line_delays_plot",
+    save_plot = true,
     clinical_hline = clinical_hline,
@@
     force = true,
-    save_plot = false,
+    plotdirpath = DrWatson.projectdir("manuscript"),
+    plotname = "line_prop_alert_plot",
+    save_plot = true,
     clinical_hline = clinical_hline,

### 2024-10-30 manuscript/optimal-thresholds-loading.jl

@@ -106,6 +106,6 @@ poisson_noise_optimal_solutions = filter_optimal_threshold_characteristics_by_no
 dynamical_noise_optimal_solutions = filter_optimal_threshold_characteristics_by_noise(
     optimal_threshold_characteristics,
     DynamicalNoiseSpecification(
-        "dynamical", 5.0, 7, 14, "in-phase", 0.15, 0.05
+        "dynamical", 5.0, 7, 14, "in-phase", 0.15, 0.0492
     ),
 );

### 2024-10-30 manuscript/manuscript.qmd
Update results with suggestions from Matt

@@
 
-The threshold that maximized outbreak detection accuracy given a testing scenario depends heavily on the noise structure and magnitude.
-For example, when the ratio of average noise incidence was 8 times higher than the average measles incidence, the optimal threshold for an RDT with 90% sensitivity/specificity and 40% testing of clinic visits was 4 cases per day with the dynamical in-phase noise, whereas it is 5 cases per day when Poisson-only noise was simulated (@tbl-optimal-thresholds).
-This corresponds to an accuracy of 72% and 93%, respectively (@fig-accuracy).
-Much higher accuracy can be achieved with imperfect tests under Poisson noise because large spikes of non-measles febrile rash that would lead to many false positives and trigger erroneous alerts do not occur.
-With dynamical noise, this possibility can and does occur and, when the ratio of non-measles to measles cases is sufficiently high, drastically impacts the accuracy of an alert system.
-The results from ELISA tests are identical between the different noise structures: with a perfect test there are no false positive results.
-For dynamical noise, RDTs never perform as well as an ELISA, even under scenarios that include a 14-day lag between test procurement and result receipt.
-However, with Poisson noise, once testing reaches 20% of clinic visits, equivalence can be observed at approximately 91% accuracy (@fig-accuracy).
+The threshold that maximized outbreak detection accuracy depends on diagnostic test characteristics, the testing rate, and the structure of the non-measles noise (@tbl-optimal-thresholds).
+When the average noise incidence was 8 times higher than the average measles incidence, the optimal threshold ranged between 1 and 7 test-positive cases per day.
+Not surprisingly, the biggest driver of this difference was the testing rate; as a large fraction of suspected cases are tested the optimal threshold increases monotonically for all test and noise types (@tbl-optimal-thresholds).
+The maximal attainable outbreak detection accuracy at the optimal threshold depends strongly on the structure and magnitude of the background noise.
+For Poisson noise, at all magnitudes, the maximum outbreak detection accuracy increases rapidly from 65% at 10% of suspected cases tested to ~90% accuracy at ≥20% testing for all test types (@fig-accuracy).
+For dynamical SEIR noise, the ELISA-like tests perform similarly to the Poisson noise case at all magnitudes (@fig-accuracy).
+For RDT-like tests, which have lower individual sensitivity and specificity, the maximal attainable accuracy is lower than the ELISA-like test for all testing rates when the magnitude of background noise is 2-times measles incidence or higher (@fig-accuracy).
+Notably, the maximal attainable accuracy declines with increasing noise and, at all noise levels, decreases with higher testing rates as the signal becomes dominated by false positive tests (@fig-accuracy).
 
@@
 
-Optimal threshold for each testing scenario, when the average noise incidence is 8 times higher than the average measles incidence.
-A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence.
-B) the noise structure is Poisson only
-
 :::
-
-Moving across a single row of @tbl-optimal-thresholds, as the testing rate increases, so does the optimal threshold, because more individuals are testing positive.
-However, the changes in accuracy are not monotonically increasing, as one might anticipate.
-Even with a perfect test, there are sequential cells that increase the testing rate, yet accuracy decreases (@fig-accuracy).
-For example, the ELISA equivalent with a 0-day test result lag outbreak detection accuracy decreases from 91% to 87% when clinic testing rates increase from 30% to 40% of clinic visits.
-The reason behind this unintuitive result stems from the use of integer valued thresholds.
-While computing the moving average test positive incidence helps to smooth out the alert system, an integer valued threshold can result in step-changes of accuracy between two threshold values, and the expected increase in the alert system's PPV from a higher threshold value is outweighed by the loss in the alert system's sensitivity to detecting outbreaks.
-Even with a perfect test, the alert system needs to discriminate between endemic/imported cases and epidemic cases; increasing the testing rate will result in higher numbers of test positive individuals, and a lower threshold can result in an overly sensitive alert system, triggering for measles infections, but not those within an outbreak.
-A higher threshold will face the opposite issue; not triggering for smaller outbreaks, which is more likely to be an issue at lower testing rates.
-Examining the number of unavoidable cases (@fig-unavoidable) and the detection delays (@fig-delay), we can see that the decrease in accuracy results from a decrease in the sensitivity of the alert system: the detection delay increases by 6 days, and the unavoidable cases by c.
-2300.
-Although the testing rate increases, so does the optimal threshold, from 3 test positives to 4 test positives per day.
-The subsequent change in testing rate (40 - 50%) is associated with no change in the optimal threshold for the ELISA, and as expected, the number of unavoidable cases and detection delay decrease, resulting in a more sensitive alert system, with an higher system accuracy indicating that the increase more than offset the decrease in the system's PPV.
 
@@
 
-```{julia}
-#| fig-cap: "The number of unavoidable cases of outbreak detection systems under different testing rates and noise structures.
-The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
-#| label: fig-unavoidable
-line_plot(
-    optimal_threshold_characteristics;
-    outcome = :unavoidable_cases,
-    ylabel = "Unavoidable Cases",
-    facet_fontsize = 18,
-    labelsize = 20,
-    show_x_facet_label = true,
-    show_y_facet_label = false,
-    ylims = (0, 2.4e4),
-    force = true,
-    plotdirpath = DrWatson.projectdir("manuscript"),
-    plotname = "line_unavoidable_plot",
-    save_plot = true,
-    clinical_hline = clinical_hline,
-    cases_scaling = gha_2022_scale_population_per_annum
-)
-```
+Introducing a lag in test result reporting necessarily decreases outbreak detection accuracy because an alert can only begin once the test results are in-hand, which increases the chance that an outbreak will end before the result.
+For the conditions simulated here, introducing a 14-day lag in test reporting for an ELISA-like test reduces the outbreak detection accuracy by ###%.
+For all simulated scenarios, this is consistent with, or higher than the accuracy achievable with an RDT-like test.
+This always leads to an increase in the median delay from outbreak start to alert relative to an ELISA-like test with no detection and frequently leads to a detection delay relative to an RDT-like test (@fig-delay).
 
@@
 
+Optimal threshold for each testing scenario, when the average noise incidence is 8 times higher than the average measles incidence.
+A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence.
+B) the noise structure is Poisson only
+
+It is notable that outbreak detection accuracy and detection delays do not increase monotonically with an increase in testing rate, and this holds regardless of the type of test.
+The reason behind this unintuitive results stems from the use of integer-valued thresholds.
+An integer valued threshold can result in step-changes of accuracy between two threshold values, and the expected increase in the alert system's PPV from a higher threshold value is outweighed by the loss in the alert system's sensitivity to detecting outbreaks.
+Even with a perfect test, the alert system must discriminate between endemic/imported cases and epidemic cases; increasing the testing rate will result in higher numbers of test positive individuals, and a lower threshold can result in an overly sensitive alert system, triggering for measles infections, but not those within an outbreak.
+A higher threshold will face the opposite issue; not triggering for smaller outbreaks, which is more likely to be an issue at lower testing rates.
+
+In general, the increase in accuracy with higher testing rates, is accompanied with longer testing delays.
+This reflects the change from highly sensitive systems with low thresholds to more specific systems with higher thresholds at higher testing rates.
 For Poisson noise, similar detection delays are observed for all test and noise magnitudes, with variation by testing rate (mean of -3.7 to 36.1 days).
@@
 With large amounts of dynamical noise (8 times the measles incidence), the mean detection delay of the 90% and 85% RDTs range from -17.5 days to 3.2 days, and from -25.2 days to -3.4 days, respectively.
+ Long detection delays manifest as large numbers of unavoidable cases (i.e., cases that occur between the outbreak start and its detection.
+Given the initial exponential rate of increase of outbreaks, the pattern of unavoidable cases follows the same shape as for detection delays, but more exaggerated.
 Negative delays indicate that alerts are being triggered before the start of the outbreak and is correlated with the proportion of the time series that is under alert, with larger negative delays associated with more and/or longer alert periods (@fig-alert-proportion, Supplemental Figure 2).
@@
     clinical_hline = clinical_hline,
+)
+```
+
+The non-monotic relationship between testing rate and both detection delay and unavoidable cases can be explained by the step changes in sensitivity due to integer thresholds and the overall sensitivity of the system.
+Notably, for all noise and testing combinations, the detection delay and number of unavoidable cases is lowest at low testing rates.
+This occurs because the system performs best with a highly sensitive threshold (1 test positive case) that means that the system in “alert” 15-30% of the time (@fig-alert-proportion, Supplemental Figure 2).
+Higher testing rates give rise to more a more specific outbreak detection system and a decrease in the proportion of the time series that is designated as alert.
+Note that for high testing rates and high ratios of dynamical noise to signal RDTs dominate the system with false positives leading to a correspondingly high proportion of the time series in alert (@fig-alert-proportion).
+
+Examining the number of unavoidable cases (@fig-unavoidable) and the detection delays (@fig-delay), we can see that the decrease in accuracy results from a decrease in the sensitivity of the alert system: the detection delay increases by 6 days, and the unavoidable cases by c.
+2300.
+Although the testing rate increases, so does the optimal threshold, from 3 test positives to 4 test positives per day.
+The subsequent change in testing rate (40 - 50%) is associated with no change in the optimal threshold for the ELISA, and as expected, the number of unavoidable cases and detection delay decrease, resulting in a more sensitive alert system, with a higher system accuracy indicating that the increase more than offset the decrease in the system's PPV.
+
+```{julia}
+#| fig-cap: "The number of unavoidable cases of outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| label: fig-unavoidable
+line_plot(
+    optimal_threshold_characteristics;
+    outcome = :unavoidable_cases,
+    ylabel = "Unavoidable Cases",
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (0, 2.4e4),
+    force = true,
+    plotdirpath = DrWatson.projectdir("manuscript"),
+    plotname = "line_unavoidable_plot",
+    save_plot = true,
+    clinical_hline = clinical_hline,
+    cases_scaling = gha_2022_scale_population_per_annum
 )

### 2024-10-30 manuscript/manuscript.qmd
Make methods noise description more explicit

@@
 
-For each noise structure, we simulated five different magnitudes of noise ($\Lambda$).
+For each noise structure, we simulated five different magnitudes of noise ($\Lambda$), representing the average daily noise incidence.
 $\Lambda$ was calculated as a multiple ($c$) of the average daily measles incidence ($\langle \Delta I_{M} \rangle$): $\Lambda =   c \cdot \langle \Delta I_{M} \rangle \text{where} c \in \{1, 2, 4, 6, 8 \}$.
-For the Poisson-noise scenarios, independent draws from a Poisson distribution with mean $\lambda = \Lambda$ were simulated to produce the noise time series.
-For the dynamical noise scenarios, the vaccination rate at birth was set to 85.38%, 73.83%, 50.88%, 27.89%, and 4.92% to produce equivalent mean daily noise values (to within 2 decimal places): $\Lambda = \langle \Delta I_{R} \rangle = c \cdot \langle \Delta I_{M} \rangle$.
+For the Poisson-noise scenarios, independent draws from a Poisson distribution with mean $c \cdot \langle \Delta I_{M} \rangle$ were simulated to produce the noise time series i.e., $\Lambda = \text{Pois}\left( c \cdot \langle \Delta I_{M} \rangle \right)$.
+For the dynamical noise scenarios, the rubella vaccination rate at birth was set to 85.38%, 73.83%, 50.88%, 27.89%, and 4.92% to produce equivalent values of $\Lambda$ (to within 2 decimal places): $\Lambda = \langle \Delta I_{R} \rangle + \text{Pois}\left( 0.15 \cdot \langle \Delta I_{R} \rangle \right)$.
 100 time series of 100 years were simulated for each scenario, before summarizing the distributions of outbreak detection methods.

### 2024-10-30 manuscript/manuscript.qmd
Add some citations for TPP and min rdt characteristics

@@
   We selected the optimal threshold as the value that maximizes accuracy.
-  We show how the optimal threshold and resulting accuracy depend on the diagnostic test, reporting rate, and the type and magnitude of the non-measles noise.
+  We show how the optimal threshold and resulting accuracy depend on the diagnostic test, testing rate, and the type and magnitude of the non-measles noise.
 
@@
 73%).
-For dynamical noise, RDT-based scenarios typically favored more sensitive alert threshold than ELISA-based scenarios (at a given reporting rate), observed with lower numbers of unavoidable cases and detection delays.
+For dynamical noise, RDT-based scenarios typically favored more sensitive alert threshold than ELISA-based scenarios (at a given testing rate), observed with lower numbers of unavoidable cases and detection delays.
 
@@
 Because measles symptoms are non-specific, it is important to account for non-measles sources of febrile rash e.g., rubella, parvovirus, varicella, etc., producing the potential for false positive results in the context of imperfect tests.
-Currently, measles outbreaks are declared on the basis of either suspected measles cases (i.e., an individual with fever and maculopapular rash [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018]) alone, cases confirmed by enzyme-linked immunosorbent assay (ELISA) to detect the presence of measles-specific IgM antibodies, or a combination of the two, depending on the elimination status of the region, with countries nearer elimination increasing the use of ELISA diagnostic confirmation for suspected cases [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018].
+Currently, measles outbreaks are declared on the basis of either suspected measles cases (i.e., an individual with fever and maculopapular rash [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018]) alone, cases confirmed by enzyme immunoassay / enzyme-linked immunosorbent assay, from here-on in referred to as ELISA, to detect the presence of measles-specific IgM antibodies, or a combination of the two, depending on the elimination status of the region, with countries nearer elimination increasing the use of ELISA diagnostic confirmation for suspected cases [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018].
 Each of these detection systems have their flaws.
@@
 In resource-poor settings these delays may be days to weeks.
-A rapid diagnostic test offers the opportunity to provide a compromise between diagnostic accuracy and timeliness, and recent developments show encouraging signs in the field as well as in theory [@warrenerEvaluationRapidDiagnostic2023; @brownRapidDiagnosticTests2020].
+A rapid diagnostic test may meet the WHO's definition and requirements for using in a surveillance setting, if not for individual patient care [@worldhealthorganizationSelectionUseEssential2021].
+Recent developments show encouraging signs in the field as well as in theory, providing a compromise between diagnostic accuracy and timeliness in most [@warrenerEvaluationRapidDiagnostic2023; @brownRapidDiagnosticTests2020], though not all [@seninMeaslesIgMRapid2024], settings.
 
@@
 That is, 85% of true measles cases will be correctly labelled as positive, and 15% of non-measles febrile rash individuals that are tested will be incorrectly labelled as positive for measles.
-This acts as a lower bound of acceptability for a new measles RDT.
-- An RDT equivalent with 90% sensitivity and specificity, and 0-day lag in result return.
-- An ELISA equivalent test with 100% sensitivity and specificity with a 0-day test result delay.
+This acts as a lower bound of acceptability for a new measles RDT [@20240613_tpp_measles_rubell_FV_EN].
+_**NOTE: TPP document used here was potentially influenced by the prelim results of this paper**_
+- An RDT equivalent with 90% sensitivity and specificity, and 0-day lag in result return [@brownRapidDiagnosticTests2020].
+- An ELISA equivalent test with 100% sensitivity and specificity with a 0-day test result delay [@hiebertEvaluationDiagnosticAccuracy2021].
 <!--- An ELISA equivalent test with 100% sensitivity and specificity with a 7-day test result delay.-->

### 2024-10-30 manuscript/manuscript.qmd
Update references in intro

@@
 Although clinical case definition is very fast and requires limited resources, it is highly sensitive, and in the face of high “background noise” from non-measles sources of febrile rash, can lead to low positive predictive value (PPV), i.e., the probability that an alert accurately captures an outbreak [@hutchinsEvaluationMeaslesClinical2004].
-And while ELISA confirmation is the standard diagnostic test for measles surveillance and has higher specificity [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018], the training and facility requirements generally mean that samples must be transported from the point of care to a separate laboratory, which incurs both costs and delays _**[ADD REFERENCE HERE]**_.
+And while ELISA confirmation is the standard diagnostic test for measles surveillance and has higher specificity [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018], the training and facility requirements generally mean that samples must be transported from the point of care to a separate laboratory, which incurs both costs and delays [@essentialprogrammeonimmunizationepiimmunizationClinicalSpecimensLaboratory2018; @worldhealthorganizationMeaslesOutbreakGuide2022; @brownRapidDiagnosticTests2020]
 In resource-poor settings these delays may be days to weeks.
 A rapid diagnostic test may meet the WHO's definition and requirements for using in a surveillance setting, if not for individual patient care [@worldhealthorganizationSelectionUseEssential2021].
-Recent developments show encouraging signs in the field as well as in theory, providing a compromise between diagnostic accuracy and timeliness in most [@warrenerEvaluationRapidDiagnostic2023; @brownRapidDiagnosticTests2020], though not all [@seninMeaslesIgMRapid2024], settings.
+Recent developments show encouraging signs in the field as well as in theory, providing a compromise between diagnostic accuracy and timeliness in most [@shonhaiInvestigationMeaslesOutbreak2015; @warrenerEvaluationRapidDiagnostic2023; @brownRapidDiagnosticTests2020], though not all [@seninMeaslesIgMRapid2024], settings.
 

### 2024-10-31 manuscript/manuscript.qmd
Update manuscript to show separate elisa and perfect test results

@@
     clinical_hline = clinical_hline,
+    nbanks = 3
 )

### 2024-10-31 OutbreakDetectionUtils/src/structs.jl

@@ -344,7 +344,7 @@ end
 function get_test_description(test_specification::IndividualTestSpecification)
     description = Match.@match test_specification begin
         IndividualTestSpecification(1.0, 0.0, 0) => "Clinical case definition"
-        [IndividualTestSpecification(0.98, 0.98, x::Int), IndividualTestSpecification(0.95, 0.98, x::Int)] => "ELISA-like ($(test_specification.sensitivity)% sens, $(test_specification.specificity)% spec, $(x) day lag)"
+        IndividualTestSpecification(0.98, 0.98, x::Int) || IndividualTestSpecification(0.95, 0.98, x::Int) => "ELISA-like ($(test_specification.sensitivity * 100)% sens, $(test_specification.specificity * 100)% spec, $(x) day lag)"
         IndividualTestSpecification(1.0, 1.0, x::Int) => "Perfect test ($(x) day lag)"
         IndividualTestSpecification(x::AbstractFloat, x::AbstractFloat, 0) where {x<1.0} => "RDT-like ($(test_specification.sensitivity * 100)% sens/spec)"
     end

### 2024-10-31 manuscript/optimal-thresholds-loading.jl

@@ -14,13 +14,28 @@ if false
 end
 
 #%%
-rdt_test_spec_vec = [IndividualTestSpecification(0.85, 0.85, 0),
-    IndividualTestSpecification(0.9, 0.9, 0)]
+rdt_test_spec_vec = [
+    IndividualTestSpecification(0.85, 0.85, 0),
+    IndividualTestSpecification(0.9, 0.9, 0),
+]
 
-elisa_test_spec_vec = [IndividualTestSpecification(1.0, 1.0, 0),
-    IndividualTestSpecification(1.0, 1.0, 14)]
+elisa_test_spec_vec = [
+    # IndividualTestSpecification(0.98, 0.98, 0),
+    IndividualTestSpecification(0.98, 0.98, 3),
+    # IndividualTestSpecification(0.98, 0.98, 7),
+    IndividualTestSpecification(0.98, 0.98, 14),
+]
 
-optimal_threshold_test_spec_vec = vcat(rdt_test_spec_vec, elisa_test_spec_vec)
+perfect_test_spec_vec = [
+    IndividualTestSpecification(1.0, 1.0, 0),
+    # IndividualTestSpecification(1.0, 1.0, 3),
+    # IndividualTestSpecification(1.0, 1.0, 7),
+    IndividualTestSpecification(1.0, 1.0, 14),
+]
+
+optimal_threshold_test_spec_vec = vcat(
+    rdt_test_spec_vec, elisa_test_spec_vec, perfect_test_spec_vec
+)
 
 optimal_threshold_alertthreshold_vec = collect(1:1:15)
 

### 2024-10-31 manuscript/optimal-thresholds-tables.jl

@@ -56,20 +56,30 @@ poisson_noise_rdt_optimal_solutions = filter(
     poisson_noise_optimal_solutions,
 );
 
-elisa_optimal_solutions = filter(
+dynamical_noise_elisa_optimal_solutions = filter(
     chars -> in(chars.individual_test_specification, elisa_test_spec_vec),
     dynamical_noise_optimal_solutions,
 );
 
+poisson_noise_elisa_optimal_solutions = filter(
+    chars -> in(chars.individual_test_specification, elisa_test_spec_vec),
+    poisson_noise_optimal_solutions,
+);
+
+perfect_test_optimal_solutions = filter(
+    chars -> in(chars.individual_test_specification, perfect_test_spec_vec),
+    poisson_noise_optimal_solutions,
+);
+
 #%%
-elisa_df = create_optimal_threshold_summary_df(
-    elisa_optimal_solutions,
+perfect_test_df = create_optimal_threshold_summary_df(
+    perfect_test_optimal_solutions,
     [:unavoidable_cases, :detectiondelays, :alert_duration_vec];
     percentiles = nothing,
     nboots = nothing,
 )
-elisa_df[!, :noise_spec] .= "Common"
-elisa_df
+perfect_test_df[!, :noise_spec] .= "All noise structures"
+perfect_test_df
 
 rdt_dynamical_df = create_optimal_threshold_summary_df(
     dynamical_noise_rdt_optimal_solutions,
@@ -87,8 +97,25 @@ rdt_poisson_df = create_optimal_threshold_summary_df(
 )
 rdt_poisson_df[!, :noise_spec] .= "Poisson noise"
 
+elisa_dynamical_df = create_optimal_threshold_summary_df(
+    dynamical_noise_elisa_optimal_solutions,
+    [:unavoidable_cases, :detectiondelays, :alert_duration_vec];
+    percentiles = nothing,
+    nboots = nothing,
+)
+elisa_dynamical_df[!, :noise_spec] .= "Dynamical noise: in-phase"
+
+elisa_poisson_df = create_optimal_threshold_summary_df(
+    poisson_noise_elisa_optimal_solutions,
+    [:unavoidable_cases, :detectiondelays, :alert_duration_vec];
+    percentiles = nothing,
+    nboots = nothing,
+)
+elisa_poisson_df[!, :noise_spec] .= "Poisson noise"
+
 thresholds_df = vcat(
[…]

### 2024-10-31 src/line_plots.jl

@@ -7,6 +7,7 @@ using StructArrays
 lineplot_colors = [
     "#56B4E9"
     "#E69F00"
+    repeat(["#483248"], 2)...
     repeat(["#000000"], 2)...
 ]
 
@@ -31,6 +32,7 @@ function line_plot(
     hidedecorations = (true, true),
     clinical_hline = true,
     hlines = nothing,
+    nbanks = 3,
     force = false,
     save_plot = true,
     kwargs...,
@@ -65,6 +67,7 @@ function line_plot(
             hidedecorations = hidedecorations,
             clinical_hline = clinical_hline,
             hlines = hlines,
+            nbanks = nbanks,
             force = force,
             save_plot = save_plot,
             kwargs...,
@@ -82,6 +85,7 @@ function line_plot(
     plotformat = "png",
     size = (2200, 1200),
     colors = lineplot_colors,
+    alpha = 0.3,
     xlabel = "Proportion Tested",
     ylabel = "Outbreak Detection\nAccuracy",
     facet_fontsize = 24,
@@ -92,6 +96,7 @@ function line_plot(
     hidedecorations = (true, true),
     clinical_hline = true,
     hlines = nothing,
+    nbanks = 3,
     force = false,
     save_plot = true,
     kwargs...,
@@ -116,9 +121,12 @@ function line_plot(
             ),
         )
         num_noise_descriptions = length(unique_noise_descriptions)
-        unique_test_specifications = unique(
-            optimal_thresholds_chars_array[1, 1].individual_test_specification
-        )
+        unique_test_specifications = sort(
+            unique(
+                optimal_thresholds_chars_array[
+                    1, 1
+                ].individual_test_specification,
+            ); by = t -> (t.sensitivity, t.specificity), rev = false)
 
         for i in axes(optimal_thresholds_chars_array, 1)
             noise_description = get_noise_description(
@@ -143,6 +151,7 @@ function line_plot(
                     outcome = outcome,
                     num_noise_descriptions = num_noise_descriptions,
                     colors = colors,
+                    alpha = alpha,
                     xlabel = xlabel,
                     ylabel = "$label_noise_description\n" * ylabel,
                     ylims = ylims,
@@ -183,19 +192,20 @@ function line_plot(
                     LineElement(; color = colors[i], linestyle = linestyle),
                 ]
             end,
-            map(
-                test_description ->
-                    replace.(
-                        test_description,
[…]

### 2024-10-31 manuscript/manuscript.qmd
Reduce alpha for manuscript line plots

--- 2024-10-30 manuscript/manuscript.qmd
@@
     ylabel = "Outbreak Detection\nAccuracy",
+    alpha = 0.1,
     facet_fontsize = 18,
@@
     ylabel = "Detection Delays\n(Days)",
+    alpha = 0.1,
     hlines = (0.0),
@@
     ylabel = "Proportion of Time Series\nIn Alert",
+    alpha = 0.1,
     hlines = (0.0),
@@
     ylabel = "Unavoidable Cases",
+    alpha = 0.1,
     facet_fontsize = 18,

### 2024-10-31 manuscript/manuscript.qmd
Clean up test and noise descriptions more and remove some extra notes

--- 2024-10-30 manuscript/manuscript.qmd
@@
 $\Lambda$ was calculated as a multiple ($c$) of the average daily measles incidence ($\langle \Delta I_{M} \rangle$): $\Lambda =   c \cdot \langle \Delta I_{M} \rangle \text{where} c \in \{1, 2, 4, 6, 8 \}$.
+Noise magnitudes will be denoted as $\Lambda \left(c\right)$ for the rest of the magnitude e.g., $\Lambda \left(8\right)$ to denote simulations where the average noise incidence is 8 times that of the average measles incidence.
 For the Poisson-noise scenarios, independent draws from a Poisson distribution with mean $c \cdot \langle \Delta I_{M} \rangle$ were simulated to produce the noise time series i.e., $\Lambda = \text{Pois}\left( c \cdot \langle \Delta I_{M} \rangle \right)$.
@@
 - An RDT equivalent with 90% sensitivity and specificity, and 0-day lag in result return [@brownRapidDiagnosticTests2020].
-- An ELISA equivalent test with 100% sensitivity and specificity with a 0-day test result delay [@hiebertEvaluationDiagnosticAccuracy2021].
-<!--- An ELISA equivalent test with 100% sensitivity and specificity with a 7-day test result delay.-->
-- An ELISA equivalent test with 100% sensitivity and specificity with a 14-day test result delay.
+- A best-case scenario ELISA equivalent with 98 % sensitivity and specificity, and 3-day lag in result return [@hiebertEvaluationDiagnosticAccuracy2021]
+- An ELISA equivalent with 98 % sensitivity and specificity, and 14-day lag in result return [@hiebertEvaluationDiagnosticAccuracy2021]
+- A perfect test with 100% sensitivity and specificity with a 0-day test result delay
+- A perfect test with 100% sensitivity and specificity with a 14-day test result delay
 
 The time series of "test-positive cases" is the daily count of those tested cases that return a positive test result.
-Thus, for each measles (@fig-outbreak-schematic b) and non-measles noise (@fig-outbreak-schematic a) time series, we have 1 time series of outbreaks (@fig-outbreak-schematic a) and 5 possible time series of test-positive cases (@fig-outbreak-schematic c) (all clinically compatible cases, plus the 4 testing scenarios), which will include false positive and negative cases resulting from imperfect diagnostic tests, that can be used to trigger outbreak alerts.
-
-## Triggering Alerts
-
-_**Need to mention that an alert is any time the threshold is exceeded, regardless of the number of consecutive days of exceedance.**_
-_**Clearly state that 1 day of exceedance is and 30 days of consecutive exceedance are both treated as a single alert i.e., the same.**_
-
-We define an "alert" as any contiguous string of 1 or more days where the 7-day moving average of the test-positive cases is greater than, or equal to, a pre-specified alert threshold, T.
-For each time series of test-positive cases, we calculate the percentage of alerts that are "correct", defined as any overlap of 1 or more days between the alert and outbreak period (@fig-outbreak-schematic c); this is analogous to the PPV of the alert system.
-Note that it is possible to have multiple alerts within a single outbreak, if the 7-day moving average of test positive cases drops below the threshold, and each would be considered correct.
-For all outbreaks that occur in the measles time series, we calculate the percentage for which at least 1 alert occurred between its start and end date (@fig-outbreak-schematic b); this is analogous to the sensitivity of the alert system.
-We also calculate the detection delay as the time from the start of an outbreak to the start of its first alert.
-If the alert starts before the outbreak and continues past the start date of the outbreak, this could be considered a correct alert with a negative delay i.e., an early warning triggered by a false positive.
+Thus, for each non-measles noise (@fig-outbreak-schematic a) and measles (@fig-outbreak-schematic b) time series, we have 1 time series of outbreaks (@fig-outbreak-schematic c) and 5 possible time series of test-positive cases (@fig-outbreak-schematic c) (all clinically compatible cases, plus the 4 testing scenarios), which will include false positive and negative cases resulting from imperfect diagnostic tests, that can be used to trigger outbreak alerts.
 
@@
 
-Finally, for each time series we calculate the number of unavoidable and avoidable outbreak cases and deaths.
+## Triggering Alerts
+
+We define an "alert" as any consecutive string of 1 or more days where the 7-day moving average of the test-positive cases is greater than, or equal to, a pre-specified alert threshold, T.
+For each time series of test-positive cases, we calculate the percentage of alerts that are "correct", defined as any overlap of 1 or more days between the alert and outbreak period (@fig-outbreak-schematic c).
+This is analogous to the PPV of the alert system, which is how it will be referred to for the rest of the manuscript.
+Note that it is possible to have multiple alerts within a single outbreak, if the 7-day moving average of test positive cases drops below the threshold, and each would be considered correct.
+For all outbreaks in the measles time series, we calculate the percentage that contain at least 1 alert within the outbreak's start and end dates (@fig-outbreak-schematic b).
+We refer to this as the sensitivity of the alert system.
+We also calculate the detection delay as the time from the start of an outbreak to the start of its first alert.
+If the alert starts before the outbreak and continues past the start date of the outbreak, this could be considered a correct alert with a negative delay i.e., an early warning triggered by a false positive.
+Finally, for each time series we calculate the number of unavoidable and avoidable outbreak cases.
 Avoidable cases are defined as those that occur within an outbreak after a correct alert is first triggered i.e., cases that could theoretically be prevented with a perfectly effective and timely response.
 Unavoidable cases are the inverse: those that occur before a correct alert, or those that occur in an undetected outbreak.
-Using a fitted age distribution of measles cases in Ghana [@mintaProgressMeaslesElimination2023] and the age-specific case fatality ratio for Ghana in 2022 [@sbarraEstimatingNationallevelMeasles2023], we scaled the values for each one-year age cohort, per annum.
 In practice, not all cases deemed avoidable are (due to imperfect and delays in responses), but to minimize the sensitivity of the results to the response implementation and operational constraints we are counting them as such.
 
-We define the accuracy of the surveillance system for a given time series as the mean of the percentage of alerts that are correct and the proportion of the outbreaks with an alert.
+We define the accuracy of the surveillance system for a given time series as the mean of the system's sensitivity and PPV.
 To examine the interaction of test and surveillance systems we varied the alert threshold, between 1 and 15 cases per day, and identified the threshold T for each combination of individual test, testing rate, noise structure and magnitude that maximizes accuracy.
-We also examined an alert method that required either the daily incidence or the 7-day moving average of daily incidence to exceed the alert threshold (see supplemental results).
 
@@
 The threshold that maximized outbreak detection accuracy depends on diagnostic test characteristics, the testing rate, and the structure of the non-measles noise (@tbl-optimal-thresholds).
-When the average noise incidence was 8 times higher than the average measles incidence, the optimal threshold ranged between 1 and 7 test-positive cases per day.
-Not surprisingly, the biggest driver of this difference was the testing rate; as a large fraction of suspected cases are tested the optimal threshold increases monotonically for all test and noise types (@tbl-optimal-thresholds).
+When the average noise incidence was 8 times higher than the average measles incidence ($\Lambda \left(8\right)$), the optimal threshold ranged between 1 and 7 test-positive cases per day.
+Not surprisingly, the biggest driver of this difference was the testing rate; as a large fraction of suspected cases are tested, the optimal threshold increases monotonically for all test and noise types (@tbl-optimal-thresholds).
+
 The maximal attainable outbreak detection accuracy at the optimal threshold depends strongly on the structure and magnitude of the background noise.
-For Poisson noise, at all magnitudes, the maximum outbreak detection accuracy increases rapidly from 65% at 10% of suspected cases tested to ~90% accuracy at ≥20% testing for all test types (@fig-accuracy).
-For dynamical SEIR noise, the ELISA-like tests perform similarly to the Poisson noise case at all magnitudes (@fig-accuracy).
-For RDT-like tests, which have lower individual sensitivity and specificity, the maximal attainable accuracy is lower than the ELISA-like test for all testing rates when the magnitude of background noise is 2-times measles incidence or higher (@fig-accuracy).
-Notably, the maximal attainable accuracy declines with increasing noise and, at all noise levels, decreases with higher testing rates as the signal becomes dominated by false positive tests (@fig-accuracy).
+For Poisson noise, at all magnitudes, the maximum outbreak detection accuracy increases rapidly from 65% at 10% of suspected cases tested to $\approx$ 90% accuracy at $\ge$ 20% testing for all test types (@fig-accuracy).
+For dynamical SEIR noise, the ELISA-like tests perform similarly to the Poisson noise case at all magnitudes (@fig-accuracy), as well as to the perfect tests with complete discrimination.
+For RDT-like tests, which have lower individual sensitivity and specificity, the maximal attainable accuracy is lower than the ELISA-like test for all testing rates at noise magnitude $\ge \Lambda \left(2\right)$ (@fig-accuracy).
+Notably, the maximal attainable accuracy declines with increasing noise and, at all noise levels, is not improved with higher testing rates as the signal becomes increasingly dominated by false positive tests (@fig-accuracy).
 
@@
 
+Optimal threshold for RDT-like, ELISA-like, and perfect tests, under dynamical and Poisson-like noise structures where the average daily noise incidence is 8 times the average daily measles incidence
+
 :::
@@
 #| fig-cap: "The accuracy of outbreak detection systems under different testing rates and noise structures.
-The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate.
+Solid lines represent tests with 0-day turnaround times, and dashed lines represent tests with result delays."
 #| label: fig-accuracy
@@
 Introducing a lag in test result reporting necessarily decreases outbreak detection accuracy because an alert can only begin once the test results are in-hand, which increases the chance that an outbreak will end before the result.
-For the conditions simulated here, introducing a 14-day lag in test reporting for an ELISA-like test reduces the outbreak detection accuracy by ###%.
+For the conditions simulated here, introducing a 14-day lag in test reporting for an ELISA-like test reduces the outbreak detection accuracy by $\approx$ 3%.
 For all simulated scenarios, this is consistent with, or higher than the accuracy achievable with an RDT-like test.
@@
 #| fig-cap: "The detection delay of outbreak detection systems under different testing rates and noise structures.
-The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate.
+Solid lines represent tests with 0-day turnaround times, and dashed lines represent tests with result delays."
 #| label: fig-delay
@@
     clinical_hline = clinical_hline,
+    nbanks = 3
 )
 ```
-
-Optimal threshold for each testing scenario, when the average noise incidence is 8 times higher than the average measles incidence.
-A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence.
-B) the noise structure is Poisson only
 
@@
 #| fig-cap: "The difference between proportion of the time series in alert for outbreak detection systems under different testing rates and noise structures.
-The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate.
+Solid lines represent tests with 0-day turnaround times, and dashed lines represent tests with result delays."
 #| label: fig-alert-proportion
@@
 #| fig-cap: "The number of unavoidable cases of outbreak detection systems under different testing rates and noise structures.
-The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate.
+Solid lines represent tests with 0-day turnaround times, and dashed lines represent tests with result delays."
 #| label: fig-unavoidable

### 2024-11-04 manuscript/manuscript.qmd
Add nalerts to supplemental appendix, refs to rubella params, and update title

@@
 ---
-title: "The Need to Develop a Holistic Infectious Disease Surveillance System"
+title: "Individual and Population Level Uncertainty Interact to Determine Performance of Outbreak Detection"
 header-title: true
@@
 For Poisson-only noise, the number of non-measles febrile rash cases each day were independent draws from a Poisson distribution with mean $\lambda$.
-For dynamical noise, we generated time series of cases from an SEIR model that matched the measles model in structure, but had $R_0 = 5$, mean latent period of 7 days, and mean infectious period of 14 days, and added some additional noise independently drawn from a Poisson distribution with mean equal to 15% of the average daily rubella incidence from the SEIR time series to account for non-rubella sources of febrile rash (@tbl-model-parameters) [_**REFERENCES**_].
+For dynamical noise, we generated time series of cases from an SEIR model that matched the measles model in structure, but had $R_0 = 5$, mean latent period of 7 days, and mean infectious period of 14 days, and added some additional noise independently drawn from a Poisson distribution with mean equal to 15% of the average daily rubella incidence from the SEIR time series to account for non-rubella sources of febrile rash (@tbl-model-parameters) [@papadopoulosEstimatesBasicReproduction2022; @RubellaCDCYellow].
 The seasonality for the dynamical noise was assumed to be in-phase with measles, anti-phase with measles (peak timing 6 months later), or non-seasonal.
@@
 That is, 85% of true measles cases will be correctly labelled as positive, and 15% of non-measles febrile rash individuals that are tested will be incorrectly labelled as positive for measles.
-This acts as a lower bound of acceptability for a new measles RDT [@20240613_tpp_measles_rubell_FV_EN].
-_**NOTE: TPP document used here was potentially influenced by the prelim results of this paper**_
+This acts as a lower bound of acceptability for a new measles RDT [@20240613_tpp_measles_rubell_FV_EN] _**NOTE: TPP document used here was potentially influenced by the prelim results of this paper**_
 - An RDT equivalent with 90% sensitivity and specificity, and 0-day lag in result return [@brownRapidDiagnosticTests2020].
-- A best-case scenario ELISA equivalent with 98 % sensitivity and specificity, and 3-day lag in result return [@hiebertEvaluationDiagnosticAccuracy2021]
-- An ELISA equivalent with 98 % sensitivity and specificity, and 14-day lag in result return [@hiebertEvaluationDiagnosticAccuracy2021]
-- A perfect test with 100% sensitivity and specificity with a 0-day test result delay
-- A perfect test with 100% sensitivity and specificity with a 14-day test result delay
+<!--- A best-case scenario ELISA equivalent with 98 % sensitivity and specificity, and 3-day lag in result return [@hiebertEvaluationDiagnosticAccuracy2021]-->
+<!--- An ELISA equivalent with 98 % sensitivity and specificity, and 14-day lag in result return [@hiebertEvaluationDiagnosticAccuracy2021]-->
+- An ELISA-like perfect test with 100% sensitivity and specificity, and a 0-day test result delay.
+This is more accurate than is observed for current ELISA tests [@hiebertEvaluationDiagnosticAccuracy2021], but it used to evaluate the theoretical best-case scenario
+- An ELISA-like perfect test with 100% sensitivity and specificity, and a 14-day test result delay
 
@@
 using OutbreakDetection: line_plot
+alpha = 0.2
+nbanks = 1
+legend_rowsize = Makie.Relative(0.05);
 ```
@@
     ylabel = "Outbreak Detection\nAccuracy",
-    alpha = 0.1,
+    alpha = alpha,
     facet_fontsize = 18,
@@
     clinical_hline = clinical_hline,
+    nbanks = nbanks,
+    legend_rowsize = legend_rowsize,
 )
@@
     ylabel = "Detection Delays\n(Days)",
-    alpha = 0.1,
+    alpha = alpha,
     hlines = (0.0),
@@
     clinical_hline = clinical_hline,
-    nbanks = 3
+    nbanks = nbanks,
+    legend_rowsize = legend_rowsize,
 )
@@
     ylabel = "Proportion of Time Series\nIn Alert",
-    alpha = 0.1,
+    alpha = alpha,
     hlines = (0.0),
@@
 Notably, for all noise and testing combinations, the detection delay and number of unavoidable cases is lowest at low testing rates.
-This occurs because the system performs best with a highly sensitive threshold (1 test positive case) that means that the system in “alert” 15-30% of the time (@fig-alert-proportion, Supplemental Figure 2).
+This occurs because the system performs best with a highly sensitive threshold (1 test positive case), which means that the system in “alert” 15-30% of the time as a results of longer and more frequent alerts (@fig-alert-proportion, Supplemental Figures 2 & 3).
 Higher testing rates give rise to more a more specific outbreak detection system and a decrease in the proportion of the time series that is designated as alert.
@@
     ylabel = "Unavoidable Cases",
-    alpha = 0.1,
+    alpha = alpha,
     facet_fontsize = 18,
@@
     clinical_hline = clinical_hline,
-    cases_scaling = gha_2022_scale_population_per_annum
+    cases_scaling = gha_2022_scale_population_per_annum,
+    nbanks = nbanks,
+    legend_rowsize = legend_rowsize,
 )

### 2024-11-04 manuscript/optimal-thresholds-loading.jl

@@ -20,10 +20,10 @@ rdt_test_spec_vec = [
 ]
 
 elisa_test_spec_vec = [
-    # IndividualTestSpecification(0.98, 0.98, 0),
-    IndividualTestSpecification(0.98, 0.98, 3),
-    # IndividualTestSpecification(0.98, 0.98, 7),
-    IndividualTestSpecification(0.98, 0.98, 14),
+#     # IndividualTestSpecification(0.98, 0.98, 0),
+#     IndividualTestSpecification(0.98, 0.98, 3),
+#     # IndividualTestSpecification(0.98, 0.98, 7),
+#     IndividualTestSpecification(0.98, 0.98, 14),
 ]
 
 perfect_test_spec_vec = [

### 2024-11-04 manuscript/optimal-thresholds-tables.jl

@@ -56,15 +56,15 @@ poisson_noise_rdt_optimal_solutions = filter(
     poisson_noise_optimal_solutions,
 );
 
-dynamical_noise_elisa_optimal_solutions = filter(
-    chars -> in(chars.individual_test_specification, elisa_test_spec_vec),
-    dynamical_noise_optimal_solutions,
-);
-
-poisson_noise_elisa_optimal_solutions = filter(
-    chars -> in(chars.individual_test_specification, elisa_test_spec_vec),
-    poisson_noise_optimal_solutions,
-);
+# dynamical_noise_elisa_optimal_solutions = filter(
+#     chars -> in(chars.individual_test_specification, elisa_test_spec_vec),
+#     dynamical_noise_optimal_solutions,
+# );
+#
+# poisson_noise_elisa_optimal_solutions = filter(
+#     chars -> in(chars.individual_test_specification, elisa_test_spec_vec),
+#     poisson_noise_optimal_solutions,
+# );
 
 perfect_test_optimal_solutions = filter(
     chars -> in(chars.individual_test_specification, perfect_test_spec_vec),
@@ -97,24 +97,27 @@ rdt_poisson_df = create_optimal_threshold_summary_df(
 )
 rdt_poisson_df[!, :noise_spec] .= "Poisson noise"
 
-elisa_dynamical_df = create_optimal_threshold_summary_df(
-    dynamical_noise_elisa_optimal_solutions,
-    [:unavoidable_cases, :detectiondelays, :alert_duration_vec];
-    percentiles = nothing,
-    nboots = nothing,
-)
-elisa_dynamical_df[!, :noise_spec] .= "Dynamical noise: in-phase"
-
-elisa_poisson_df = create_optimal_threshold_summary_df(
-    poisson_noise_elisa_optimal_solutions,
-    [:unavoidable_cases, :detectiondelays, :alert_duration_vec];
-    percentiles = nothing,
-    nboots = nothing,
-)
-elisa_poisson_df[!, :noise_spec] .= "Poisson noise"
+# elisa_dynamical_df = create_optimal_threshold_summary_df(
+#     dynamical_noise_elisa_optimal_solutions,
+#     [:unavoidable_cases, :detectiondelays, :alert_duration_vec];
+#     percentiles = nothing,
+#     nboots = nothing,
+# )
[…]

### 2024-11-04 src/line_plots.jl

@@ -32,7 +32,8 @@ function line_plot(
     hidedecorations = (true, true),
     clinical_hline = true,
     hlines = nothing,
-    nbanks = 3,
+    nbanks = 1,
+    legend_rowsize = Makie.Relative(0.05),
     force = false,
     save_plot = true,
     kwargs...,
@@ -68,6 +69,7 @@ function line_plot(
             clinical_hline = clinical_hline,
             hlines = hlines,
             nbanks = nbanks,
+            legend_rowsize = legend_rowsize,
             force = force,
             save_plot = save_plot,
             kwargs...,
@@ -96,7 +98,8 @@ function line_plot(
     hidedecorations = (true, true),
     clinical_hline = true,
     hlines = nothing,
-    nbanks = 3,
+    nbanks = 1,
+    legend_rowsize = Makie.Relative(0.05),
     force = false,
     save_plot = true,
     kwargs...,
@@ -205,7 +208,7 @@ function line_plot(
             orientation = :horizontal,
             nbanks = nbanks,
         )
-        rowsize!(fig.layout, 0, Relative(0.1))
+        rowsize!(fig.layout, 0, legend_rowsize)
 
         if save_plot
             Makie.save(plotpath, fig; size = size)

### 2024-11-06 manuscript/ARCHIVE/manuscript.qmd
Delete old manuscript files, archive qmd, and use CairoMakie to produce SVG plots in particular directory
Going to use Typst with Make to build the manuscript files, so makes
sense to switch to CairoMakie at the same time

--- 2024-11-04 manuscript/manuscript.qmd
@@
   - "Outbreak Detection"
-word-count: true
+word-count: false
 line-numbers: true

### 2024-11-06 manuscript/ARCHIVE/supplemental-appendix.qmd
Delete old manuscript files, archive qmd, and use CairoMakie to produce SVG plots in particular directory
Going to use Typst with Make to build the manuscript files, so makes
sense to switch to CairoMakie at the same time

--- 2024-10-24 manuscript/supplemental-appendix.qmd
@@
 #| echo: false
+include("optimal-thresholds-loading.jl");
 include(srcdir("makie-plotting-setup.jl"));
 using OutbreakDetection: line_plot
+alpha = 0.2
+nbanks = 1
+legend_rowsize = Makie.Relative(0.05);
+clinical_hline = false
 ```
@@
 ```{julia}
-#| fig-cap: "The difference between proportion of the time series in outbreak for outbreak detection systems under different testing rates and noise structures.
-The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+#| fig-cap: "The difference between the proportion of the time series in outbreak for outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate.
+Solid lines represent tests with 0-day turnaround times, and dashed lines represent tests with result delays."
 #| label: fig-outbreak-proportion
@@
     ylabel = "Proportion of Time\nSeries In Outbreak",
+    alpha = alpha,
     facet_fontsize = 18,
@@
     show_y_facet_label = false,
-    ylims = (0.0, 0.35),
+    ylims = (0.5, 0.25),
     force = true,
@@
 #| fig-cap: "The alert durations of outbreak detection systems under different testing rates and noise structures.
-The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate."
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate.
+Solid lines represent tests with 0-day turnaround times, and dashed lines represent tests with result delays."
 #| label: fig-alert-duration
@@
     ylabel = "Alert Duration\n(Days)",
+    alpha = alpha,
     facet_fontsize = 18,
@@
     clinical_hline = clinical_hline,
+    nbanks = nbanks,
+    legend_rowsize = legend_rowsize,
 )
 ```
+
+```{julia}
+#| fig-cap: "The number of alerts of outbreak detection systems under different testing rates and noise structures.
+The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate.
+Solid lines represent tests with 0-day turnaround times, and dashed lines represent tests with result delays."
+#| label: fig-num-alerts
+line_plot(
+    optimal_threshold_characteristics;
+    outcome = :nalerts,
+    ylabel = "Number of Alerts",
+    alpha = alpha,
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (0, 350),
+    force = true,
+    save_plot = false,
+    clinical_hline = clinical_hline,
+    nbanks = nbanks,
+    legend_rowsize = legend_rowsize,
+)
+```

### 2024-11-06 manuscript/line-plots-values-check.jl

@@ -1,98 +1,0 @@
-#%%
-using DrWatson
-@quickactivate "OutbreakDetection"
-
-using OutbreakDetectionUtils:
-    IndividualTestSpecification, DynamicalNoiseSpecification
-using StatsBase: mean
-
-
-include(projectdir("manuscript", "optimal-thresholds-loading.jl"));
-
-if false
-    include("optimal-thresholds-loading.jl")
-end
-
-#%%
-mean_elisa_0d_delays = map(
-    test_chars -> mean(vcat(getproperty(test_chars, :detectiondelays)...)),
-    filter(
-        chars ->
-            chars.individual_test_specification ==
-            IndividualTestSpecification(1.0, 1.0, 0),
-        dynamical_noise_optimal_solutions,
-    ).outbreak_threshold_chars,
-)
-
-mean_rdt_90_8x_dynamical_delays = map(
-    test_chars -> mean(vcat(getproperty(test_chars, :detectiondelays)...)),
-    filter(
-        chars ->
-            chars.individual_test_specification ==
-            IndividualTestSpecification(0.9, 0.9, 0),
-        dynamical_noise_optimal_solutions,
-    ).outbreak_threshold_chars,
-)
-
-mean_rdt_85_8x_dynamical_delays = map(
-    test_chars -> mean(vcat(getproperty(test_chars, :detectiondelays)...)),
-    filter(
-        chars ->
-            chars.individual_test_specification ==
-            IndividualTestSpecification(0.85, 0.85, 0),
-        dynamical_noise_optimal_solutions,
-    ).outbreak_threshold_chars,
-)
-
-mapreduce(
-    vcat,
-    (
-        ("ELISA", mean_elisa_0d_delays),
-        ("90%", mean_rdt_90_8x_dynamical_delays),
-        ("85%", mean_rdt_85_8x_dynamical_delays),
-    ),
-) do (label, mean_delays_vec)
-    Dict(label => round.(extrema(mean_delays_vec); digits = 1))
-end
-
-#%%
-mean_elisa_0d_alert_duration = map(
-    test_chars -> mean(vcat(getproperty(test_chars, :alert_duration_vec)...)),
-    filter(
-        chars ->
-            chars.individual_test_specification ==
-            IndividualTestSpecification(1.0, 1.0, 0),
-        dynamical_noise_optimal_solutions,
-    ).outbreak_threshold_chars,
-)
-
-mean_rdt_90_8x_dynamical_alert_duration = map(
[…]

### 2024-11-06 manuscript/optimal-thresholds-loading.jl

@@ -1,126 +1,0 @@
-#%%
-using DrWatson
-@quickactivate "OutbreakDetection"
-
-using OutbreakDetectionUtils:
-    IndividualTestSpecification, create_combinations_vec,
-    DynamicalNoiseSpecification, PoissonNoiseSpecification
-using OutbreakDetection: collect_OptimalThresholdCharacteristics
-
-include(srcdir("ensemble-parameters.jl"))
-
-if false
-    include("../src/ensemble-parameters.jl")
-end
-
-#%%
-rdt_test_spec_vec = [
-    IndividualTestSpecification(0.85, 0.85, 0),
-    IndividualTestSpecification(0.9, 0.9, 0),
-]
-
-elisa_test_spec_vec = [
-#     # IndividualTestSpecification(0.98, 0.98, 0),
-#     IndividualTestSpecification(0.98, 0.98, 3),
-#     # IndividualTestSpecification(0.98, 0.98, 7),
-#     IndividualTestSpecification(0.98, 0.98, 14),
-]
-
-perfect_test_spec_vec = [
-    IndividualTestSpecification(1.0, 1.0, 0),
-    # IndividualTestSpecification(1.0, 1.0, 3),
-    # IndividualTestSpecification(1.0, 1.0, 7),
-    IndividualTestSpecification(1.0, 1.0, 14),
-]
-
-optimal_threshold_test_spec_vec = vcat(
-    rdt_test_spec_vec, elisa_test_spec_vec, perfect_test_spec_vec
-)
-
-optimal_threshold_alertthreshold_vec = collect(1:1:15)
-
-R_0_vec = [16.0]
-
-ensemble_dynamics_spec_vec = create_combinations_vec(
-    DynamicsParameters,
-    (
-        [ensemble_state_specification.init_states.N],
-        [27],
-        [0.2],
-        [SIGMA],
-        [GAMMA],
-        R_0_vec,
-        [0.8],
-    ),
-)
-
-ensemble_spec_vec = create_combinations_vec(
-    EnsembleSpecification,
-    (
-        [ensemble_model_type],
-        [ensemble_state_specification],
-        ensemble_dynamics_spec_vec,
-        [ensemble_time_specification],
-        [ensemble_nsims],
-    ),
-)
-
-alert_method_vec = ["movingavg"]
-
-#%%
-ensemble_noise_specification =
-    filter(ensemble_noise_specification_vec) do noise_spec
[…]

### 2024-11-06 manuscript/optimal-thresholds-tables.jl

@@ -1,243 +1,0 @@
-#%%
-using DrWatson
-@quickactivate "OutbreakDetection"
-
-using DataFrames
-using CategoricalArrays
-using Match: Match
-using CSV: CSV
-using Chain: Chain
-
-using OutbreakDetectionUtils:
-    create_optimal_thresholds_df, create_optimal_threshold_summary_df
-
-import OutbreakDetectionUtils: create_optimal_threshold_summary_df
-
-include(projectdir("manuscript", "optimal-thresholds-loading.jl"));
-
-if false
-    include("optimal-thresholds-loading.jl")
-end
-
-#%%
-cfr_df = CSV.read(
-    datadir("CFR_2022.csv"),
-    DataFrame; delim = ',',
-    header = true,
-    types = [String, Int64, Float64],
-    silencewarnings = true,
-)
-dropmissing!(cfr_df)
-
-gha_cfr = only(cfr_df[cfr_df.country .== "GHA", :CFR])
-
-population_df = CSV.read(
-    datadir("input-populations.csv"),
-    DataFrame; delim = ',',
-    header = true,
-)
-
-gha_2022_pop = only(
-    population_df[population_df.ISO3_code .== "GHA", "2022"]
-)
-gha_2022_scale_population =
-    gha_2022_pop / ensemble_state_specification.init_states.N
-
-gha_2022_scale_population_per_annum = gha_2022_scale_population / 100
-
-#%%
-dynamical_noise_rdt_optimal_solutions = filter(
-    chars -> in(chars.individual_test_specification, rdt_test_spec_vec),
-    dynamical_noise_optimal_solutions,
-);
-
-poisson_noise_rdt_optimal_solutions = filter(
-    chars -> in(chars.individual_test_specification, rdt_test_spec_vec),
-    poisson_noise_optimal_solutions,
-);
-
-# dynamical_noise_elisa_optimal_solutions = filter(
-#     chars -> in(chars.individual_test_specification, elisa_test_spec_vec),
-#     dynamical_noise_optimal_solutions,
-# );
-#
-# poisson_noise_elisa_optimal_solutions = filter(
-#     chars -> in(chars.individual_test_specification, elisa_test_spec_vec),
-#     poisson_noise_optimal_solutions,
-# );
-
-perfect_test_optimal_solutions = filter(
-    chars -> in(chars.individual_test_specification, perfect_test_spec_vec),
-    poisson_noise_optimal_solutions,
-);
-
-#%%
[…]

### 2024-11-06 manuscript/optimal-thresholds.jl

@@ -1,0 +1,33 @@
+#%%
+using DrWatson
+@quickactivate "OutbreakDetection"
+
+using DataFrames
+using CategoricalArrays
+using Match: Match
+using CSV: CSV
+using Chain: Chain
+
+using OutbreakDetectionUtils:
+    create_optimal_thresholds_df, create_optimal_threshold_summary_df
+
+import OutbreakDetectionUtils: create_optimal_threshold_summary_df
+
+#%%
+manuscriptdir(args...) = DrWatson.projectdir("manuscript", args...)
+manuscript_files(args...) = manuscriptdir("manuscript_files", args...)
+manuscript_plotdir(args...) = manuscript_files("plots", args...)
+manuscript_tabledir(args...) = manuscript_files("tables", args...)
+appendix_files(args...) = manuscriptdir("supplemental-appendix_files", args...)
+
+#%%
+include(manuscriptdir("optimal-thresholds_loading.jl"));
+
+#%%
+include(manuscriptdir("optimal-thresholds_tables.jl"));
+
+#%%
+include(manuscriptdir("optimal-thresholds_plots.jl"));
+
+#%%
+# include(manuscriptdir("optimal-thresholds_checks.jl"))

### 2024-11-06 manuscript/optimal-thresholds_checks.jl

@@ -1,0 +1,83 @@
+#%%
+mean_elisa_0d_delays = map(
+    test_chars -> mean(vcat(getproperty(test_chars, :detectiondelays)...)),
+    filter(
+        chars ->
+            chars.individual_test_specification ==
+            IndividualTestSpecification(1.0, 1.0, 0),
+        dynamical_noise_optimal_solutions,
+    ).outbreak_threshold_chars,
+)
+
+mean_rdt_90_8x_dynamical_delays = map(
+    test_chars -> mean(vcat(getproperty(test_chars, :detectiondelays)...)),
+    filter(
+        chars ->
+            chars.individual_test_specification ==
+            IndividualTestSpecification(0.9, 0.9, 0),
+        dynamical_noise_optimal_solutions,
+    ).outbreak_threshold_chars,
+)
+
+mean_rdt_85_8x_dynamical_delays = map(
+    test_chars -> mean(vcat(getproperty(test_chars, :detectiondelays)...)),
+    filter(
+        chars ->
+            chars.individual_test_specification ==
+            IndividualTestSpecification(0.85, 0.85, 0),
+        dynamical_noise_optimal_solutions,
+    ).outbreak_threshold_chars,
+)
+
+mapreduce(
+    vcat,
+    (
+        ("ELISA", mean_elisa_0d_delays),
+        ("90%", mean_rdt_90_8x_dynamical_delays),
+        ("85%", mean_rdt_85_8x_dynamical_delays),
+    ),
+) do (label, mean_delays_vec)
+    Dict(label => round.(extrema(mean_delays_vec); digits = 1))
+end
+
+#%%
+mean_elisa_0d_alert_duration = map(
+    test_chars -> mean(vcat(getproperty(test_chars, :alert_duration_vec)...)),
+    filter(
+        chars ->
+            chars.individual_test_specification ==
+            IndividualTestSpecification(1.0, 1.0, 0),
+        dynamical_noise_optimal_solutions,
+    ).outbreak_threshold_chars,
+)
+
+mean_rdt_90_8x_dynamical_alert_duration = map(
+    test_chars -> mean(vcat(getproperty(test_chars, :alert_duration_vec)...)),
+    filter(
+        chars ->
+            chars.individual_test_specification ==
+            IndividualTestSpecification(0.9, 0.9, 0),
+        dynamical_noise_optimal_solutions,
+    ).outbreak_threshold_chars,
+)
+
[…]

### 2024-11-06 manuscript/optimal-thresholds_loading.jl

@@ -1,0 +1,126 @@
+#%%
+using DrWatson
+@quickactivate "OutbreakDetection"
+
+using OutbreakDetectionUtils:
+    IndividualTestSpecification, create_combinations_vec,
+    DynamicalNoiseSpecification, PoissonNoiseSpecification
+using OutbreakDetection: collect_OptimalThresholdCharacteristics
+
+include(srcdir("ensemble-parameters.jl"))
+
+if false
+    include("../src/ensemble-parameters.jl")
+end
+
+#%%
+rdt_test_spec_vec = [
+    IndividualTestSpecification(0.85, 0.85, 0),
+    IndividualTestSpecification(0.9, 0.9, 0),
+]
+
+elisa_test_spec_vec = [
+#     # IndividualTestSpecification(0.98, 0.98, 0),
+#     IndividualTestSpecification(0.98, 0.98, 3),
+#     # IndividualTestSpecification(0.98, 0.98, 7),
+#     IndividualTestSpecification(0.98, 0.98, 14),
+]
+
+perfect_test_spec_vec = [
+    IndividualTestSpecification(1.0, 1.0, 0),
+    # IndividualTestSpecification(1.0, 1.0, 3),
+    # IndividualTestSpecification(1.0, 1.0, 7),
+    IndividualTestSpecification(1.0, 1.0, 14),
+]
+
+optimal_threshold_test_spec_vec = vcat(
+    rdt_test_spec_vec, elisa_test_spec_vec, perfect_test_spec_vec
+)
+
+optimal_threshold_alertthreshold_vec = collect(1:1:15)
+
+R_0_vec = [16.0]
+
+ensemble_dynamics_spec_vec = create_combinations_vec(
+    DynamicsParameters,
+    (
+        [ensemble_state_specification.init_states.N],
+        [27],
+        [0.2],
+        [SIGMA],
+        [GAMMA],
+        R_0_vec,
+        [0.8],
+    ),
+)
+
+ensemble_spec_vec = create_combinations_vec(
+    EnsembleSpecification,
+    (
+        [ensemble_model_type],
+        [ensemble_state_specification],
+        ensemble_dynamics_spec_vec,
+        [ensemble_time_specification],
+        [ensemble_nsims],
+    ),
+)
+
+alert_method_vec = ["movingavg"]
+
+#%%
+ensemble_noise_specification =
+    filter(ensemble_noise_specification_vec) do noise_spec
[…]

### 2024-11-06 manuscript/optimal-thresholds_plots.jl

@@ -1,0 +1,111 @@
+#%%
+manuscriptdir("plotting-setup.jl")
+
+#%%
+using OutbreakDetection: line_plot
+alpha = 0.2
+nbanks = 1
+legend_rowsize = Makie.Relative(0.05);
+
+#%%
+accuracy_plot = line_plot(
+    optimal_threshold_characteristics;
+    outcome = :accuracy,
+    ylabel = "Outbreak Detection\nAccuracy",
+    alpha = alpha,
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (0.5, 1.0),
+    force = true,
+    plotdirpath = DrWatson.projectdir("manuscript"),
+    plotname = "line_accuracy_plot",
+    save_plot = true,
+    clinical_hline = clinical_hline,
+    nbanks = nbanks,
+    legend_rowsize = legend_rowsize,
+)
+
+save(
+    manuscript_plotdir("optimal-thresholds_accuracy-plot.svg"),
+    accuracy_plot,
+)
+
+#%%
+delays_plot = line_plot(
+    optimal_threshold_characteristics;
+    outcome = :detectiondelays,
+    ylabel = "Detection Delays\n(Days)",
+    alpha = alpha,
+    hlines = (0.0),
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (-100, 100),
+    force = true,
+    plotdirpath = DrWatson.projectdir("manuscript"),
+    plotname = "line_delays_plot",
+    save_plot = true,
+    clinical_hline = clinical_hline,
+    nbanks = nbanks,
+    legend_rowsize = legend_rowsize,
+)
+
+save(
+    manuscript_plotdir("optimal-thresholds_delays-plot.svg"),
+    delays_plot,
+)
+
+#%%
+prop_alert_plot = line_plot(
+    optimal_threshold_characteristics;
+    outcome = :proportion_timeseries_in_alert,
+    ylabel = "Proportion of Time Series\nIn Alert",
+    alpha = alpha,
+    hlines = (0.0),
+    facet_fontsize = 18,
+    labelsize = 20,
+    show_x_facet_label = true,
+    show_y_facet_label = false,
+    ylims = (0, 0.35),
+    force = true,
+    plotdirpath = DrWatson.projectdir("manuscript"),
[…]

### 2024-11-06 manuscript/optimal-thresholds_tables.jl

@@ -1,0 +1,225 @@
+#%%
+cfr_df = CSV.read(
+    datadir("CFR_2022.csv"),
+    DataFrame; delim = ',',
+    header = true,
+    types = [String, Int64, Float64],
+    silencewarnings = true,
+)
+dropmissing!(cfr_df)
+
+gha_cfr = only(cfr_df[cfr_df.country .== "GHA", :CFR])
+
+population_df = CSV.read(
+    datadir("input-populations.csv"),
+    DataFrame; delim = ',',
+    header = true,
+)
+
+gha_2022_pop = only(
+    population_df[population_df.ISO3_code .== "GHA", "2022"]
+)
+gha_2022_scale_population =
+    gha_2022_pop / ensemble_state_specification.init_states.N
+
+gha_2022_scale_population_per_annum = gha_2022_scale_population / 100
+
+#%%
+dynamical_noise_rdt_optimal_solutions = filter(
+    chars -> in(chars.individual_test_specification, rdt_test_spec_vec),
+    dynamical_noise_optimal_solutions,
+);
+
+poisson_noise_rdt_optimal_solutions = filter(
+    chars -> in(chars.individual_test_specification, rdt_test_spec_vec),
+    poisson_noise_optimal_solutions,
+);
+
+# dynamical_noise_elisa_optimal_solutions = filter(
+#     chars -> in(chars.individual_test_specification, elisa_test_spec_vec),
+#     dynamical_noise_optimal_solutions,
+# );
+#
+# poisson_noise_elisa_optimal_solutions = filter(
+#     chars -> in(chars.individual_test_specification, elisa_test_spec_vec),
+#     poisson_noise_optimal_solutions,
+# );
+
+perfect_test_optimal_solutions = filter(
+    chars -> in(chars.individual_test_specification, perfect_test_spec_vec),
+    poisson_noise_optimal_solutions,
+);
+
+#%%
+perfect_test_df = create_optimal_threshold_summary_df(
+    perfect_test_optimal_solutions,
+    [:unavoidable_cases, :detectiondelays, :alert_duration_vec];
+    percentiles = nothing,
+    nboots = nothing,
+)
+perfect_test_df[!, :noise_spec] .= "All noise structures"
+perfect_test_df
+
+rdt_dynamical_df = create_optimal_threshold_summary_df(
+    dynamical_noise_rdt_optimal_solutions,
+    [:unavoidable_cases, :detectiondelays, :alert_duration_vec];
[…]

### 2024-11-06 manuscript/plotting-setup.jl

@@ -1,0 +1,27 @@
+using CairoMakie
+
+function theme_adjustments()
+    return Theme(;
+        fontsize = 16,
+        Axis = (;
+            xlabelsize = 20,
+            ylabelsize = 20,
+            xlabelfont = :bold,
+            ylabelfont = :bold,
+        ),
+        Colorbar = (;
+            labelsize = 20,
+            labelfont = :bold,
+        ),
+    )
+end
+custom_theme = merge(theme_adjustments(), theme_minimal())
+
+set_theme!(
+    custom_theme;
+    fontsize = 16,
+    linewidth = 4,
+)
+
+update_theme!(; size = (1300, 800))
+CairoMakie.activate!(; type = "svg", px_per_unit = 1.0)

### 2024-11-06 manuscript/schematic-plot.jl

@@ -1,14 +1,13 @@
 #%%
-using DrWatson
-@quickactivate "OutbreakDetection"
+manuscriptdir("plotting-setup.jl")
 
-using GLMakie
+#%%
 using StatsBase
 
-using OutbreakDetection
-using OutbreakDetectionUtils
+using OutbreakDetection: N_MISSED_OUTBREAKS_COLOR,
+    PERC_OUTBREAKS_DETECTED_COLOR, N_MISSED_OUTBREAKS_COLOR, N_ALERTS_COLOR
 
-include(srcdir("makie-plotting-setup.jl"))
+using OutbreakDetectionUtils
 
 #%%
 states_p = StateParameters(;
@@ -388,4 +387,9 @@ schematic_with_shade_fig = plot_schematic(
     outbreak_detection_specification.alert_threshold; time_p = time_p,
     shade_alert_outbreak_overlap = true,
     xlims = (5, 13),
+)
+
+save(
+    manuscript_plotdir("schematic-plot.svg"),
+    schematic_with_shade_fig,
 )

### 2024-11-06 src/line_plots.jl

@@ -458,8 +458,7 @@ function _line_plot_facet(
             subsetted_df.percent_clinic_tested,
             subsetted_df[!, outcome_10th],
             subsetted_df[!, outcome_90th];
-            color = colors[i],
-            alpha = alpha,
+            color = (colors[i], alpha),
         )
 
         ylims!(ax, ylims)
